### 2020-01-20 ccsn.tex
add files

%\documentclass[aps, twocolumn, superscriptaddress, showpacs, linenumbers, nofootinbib]{revtex4-1}
\documentclass[aps, twocolumn, superscriptaddress, showpacs, nofootinbib, longbibliography]{revtex4-1}

% \documentclass[aps,prl,reprint,showpacs]{revtex4-1}

%\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl} 
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{wasysym}
\usepackage{psfrag}
%\usepackage{comment}
\usepackage{ifthen}
\usepackage[citecolor=blue,colorlinks=true]{hyperref}
\usepackage{longtable}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
%\usepackage{ulem}
%\usepackage{subfigure}
\newcommand{\note}[1]{{\color{red}#1}}
\newcommand{\second}{\ensuremath{\mathrm{\,s}}}

% \usepackage{svn-multi}
%\restylefloat{table}

\begin{document}
\linenumbers

\title{Identification of protoneutron star g-modes in gravitational-wave data}

\begin{abstract}

\end{abstract}

\maketitle

\input{introduction}
\input{methods}
\input{noisestudy}
\input{discussion}

\bigskip\noindent\textit{Acknowledgments} ---

%\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}

### 2020-01-20 results.tex
add files

\section{Detection sensitivity with Advanced gravitational wave detectors}

### 2020-03-04 methods.tex
Edits in methods, biblio,

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
the CCSN gravitational wave observations.
An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
on 25 1D simulations with the
{\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code. Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polynomial regression with
heteroscedastic errors
\begin{equation}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with $f_i$. The model for frequency-dependent variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
{\textcolor{red}{add citation} that implements a maximum likelihood approach was used to fit the model. The
best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
\begin{equation}\label{eq:universal}
r_i=f_i + f_i^3 + \epsilon_i
\end{equation}
 The best-fitting model achieves a coefficient of determination of $R^2=$  \textcolor{red}{insert Rsquared}.
Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.
The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
\begin{figure}
 \centering
 	\includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
\end{figure}

Here we analyse the gravitational wave signal {\tt s20-gw-10kpc} \textcolor{red}{insert more detailed description}, originally sampled at 10 kHz but resampled to the LIGO sampling rate of 16384 Hz.
A spectrogram of this signal is shown in Figure \textcolor{red}{insert spectrogram of signal} based on autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.The dominant emission mode correspondes to the $\mbox{ }^2 g_1$-mode \cite{} and we have developed a time-frequency method to track the ridge $m(t)$ in the spectrogram, taking into account that it is monotonically increasing. Starting from either the left- or right-most column of the time-frequency  matrix we identify and trace the sequence of amplitude peaks within a certain frequency band given the monotonicity constraint. \textcolor{red}{Are more details regarding the ridge tracking required here?}

We identify the instantaneous frequency $f(t_i)$ corresponding the  ridge $m(t_i)$ for the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. Now we can use the universal relationship in (\ref{eq:universal}) to obtain estimates of the time evolution of the
ratios together with 95\% confidence intervals. These are given in Figure \ref{fig:xx} \textcolor{red}{insert figure} where the black points are the true ratio values, the red points the
estimates and the grey bands represent 95\% confidence bands. In this case without any noise, the coverage of our 95\% confidence band is xx\%.

In the following simulation study we explore how accurately we can estimate the parameters when the gravitational wave signal is embedded in noise.
For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise using the noise power spectral density \textcolor{red}{insert formula} for varying
SNRs, respectively distances to the source. We estimate the coverage probability of the 95\% confidence band by calculating the proportion of times that the true ratio lies outside one of the pointwise 95\% confidence intervals.
These coverage probabilities together for varying SNRs are given in Table \textcolor{red}{insert Table} and displayed in the form of boxplots in Figure \textcolor{red}{insert Figure}

### 2020-03-18 introduction.tex
First commit with the introduction

\section{Introduction}

\textbf{CCSN Simulation and GW emission}\\
The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova. Core-collapse supernova (CCSN) explosions are one most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories. This is because even the most common type of CCSN, the neutrino-driven explosion, have a rate about three per century \cite{Gossan:2016} within our galaxy. The other main type of explosion, those produced by the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016}. However, the rate of events of this kind is much lower than the one for the neutrino driven mechanism $\sim 10^{-4} \rm{yr}^{-1}$, which represents less than $1 \%$ of all CCSNe.
Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the explosion.

\textbf{Mode analysis and relations}\\
\textbf{Martin s20 simulation and code description}\\
\textbf{This paper method description and previous works}\\
\textbf{Previous work in identifying modes from spectrogram}\\
\textbf{Paper organization}\\

### 2020-05-28 methods.tex
method section update
Structure of method section.
g-mode reconstruction

@@
 These coverage probabilities together for varying SNRs are given in Table \textcolor{red}{insert Table} and displayed in the form of boxplots in Figure \textcolor{red}{insert Figure}
+
+\subsection{G-mode reconstruction}
+
+Given the spectrogram and an specified time interval for the g-mode reconstruction, our proposal method works as follows.
+The starting point must be specified.
+It can be either at the beginning or at the end of the signal.
+Then, in one of these extremes, the maximum energy value is identified, registering its frequency.
+This is done independently for a number of consecutive time intervals.
+Then we calculate the median of these frequency values, providing a robust starting value for the g-mode reconstruction.
+
+The starting frequency value is the first g-mode estimate for the first or the last time interval, depending on the specified starting location.
+If the reconstruction is set to start at the beginning of the signal, the reconstruction will be done progressively over the time intervals, where each maximum frequency value will be calculated within a frequency range specified by the previous g-mode estimate.
+Given the non-decreasing behaviour of the true g-mode values, the g-mode estimates will be forced to be greater or equal than the one estimated for its previous time interval, and lower than a specified upper limit.
+As a result, the g-modes estimates will be a non-decreasing sequence of frequency values.
+
+If the reconstruction is set to start at the end of the signal, the g-modes will be estimated backward in time.
+Each maximum frequency is calculated within a range determined by its successor (in time) g-mode estimate.
+These estimates are forced to be lower or equal than its successor (in time) estimate, but greater than a specified lower limit.
+Thus, a non-decreasing sequence of g-mode estimates is guaranteed.
+
+This g-mode reconstruction method works if and only if the signal is strong enough to provide information about the g-mode, which is reflected in the spectrogram.
+
+\subsection{ Confidence bands}
+A description of how they are built.  This is important because these bands are used to calculate the coverage probabilities in our simulation studies.

### 2020-06-03 introduction.tex
A bit of work in the introduction

@@
 
+The computational modelling of the core-collapse escenario is challenging due the large number of physical processes involved whose role in the escenario is not completely understood.
+For instance, there are uncertainties in the stellar evolution of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) or the neutrino interactions.
+These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
+One simulation of one single progenitor in 3D with accurate neutrino transport and realistic equation of state (EoS) can take several months of intense calculations on a scientific supercomputer facility.
+
 \textbf{Mode analysis and relations}\\

### 2020-06-04 methods.tex
confidence band section
Confidence band section

@@
 \subsection{ Confidence bands}
-A description of how they are built.  This is important because these bands are used to calculate the coverage probabilities in our simulation studies.
+Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:universal}.
+The g-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios.
+Since the g-mode estimates are indexed by time, the confidence intervals for the ratios are too.
+Thus, we generate the confidence band by interpolating the lower and upper limits of the collection of consecutive confidence intervals, which will be valid for the time range of the g-mode estimates.
+This confidence band is used to estimate the coverage probabilities in our simulation studies presented below.

### 2020-06-17 simulations.tex
minor changes

\section{Core collapse supernova simulations}

\textbf{Martin's simulations and code description}\\
\textbf{1D simlation data to fit the ratio vs frequency model - AA and CoConut outputs}

### 2020-06-17 introduction.tex
minor changes

@@
 \textbf{CCSN Simulation and GW emission}\\
-The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova.
-Core-collapse supernova (CCSN) explosions are one most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories.
-This is because even the most common type of CCSN, the neutrino-driven explosion, have a rate about three per century \cite{Gossan:2016} within our galaxy.
+The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading to the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova.
+Core-collapse supernova (CCSN) explosions are one of the most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories.
+This is because even the most common type of CCSN, the neutrino-driven explosion supernova, have a rate of about three per century \cite{Gossan:2016} within our galaxy.
 The other main type of explosion, those produced by the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016}.
@@
 
-The computational modelling of the core-collapse escenario is challenging due the large number of physical processes involved whose role in the escenario is not completely understood.
-For instance, there are uncertainties in the stellar evolution of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) or the neutrino interactions.
+The computational modelling of the core-collapse phenomena is challenging due the large number of physical processes involved whose role is not completely understood.
+There are uncertainties in the stellar evolution of massive stars or in the nuclear and weak force interactions necessary for the equation of state (EoS) or the neutrino interactions.
 These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
@@
 
-\textbf{Mode analysis and relations}\\
-\textbf{Martin s20 simulation and code description}\\
-\textbf{This paper method description and previous works}\\
+\textbf{PNS oscillation modes}\\
+\textbf{Asteroseismology and universal relationships}\\
+\textbf{Introduce the topic of the paper: method to extract from the GW data the Mass and the radius of the PNS as function of time using the universal relations.}
 \textbf{Previous work in identifying modes from spectrogram}\\

### 2020-06-17 methods.tex
minor changes

@@
 In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
-the CCSN gravitational wave observations.
+the gravitational wave observations.
 An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
@@
 with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
-{\textcolor{red}{add citation} that implements a maximum likelihood approach was used to fit the model. The
+\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The
 best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
@@
 \end{equation}
- The best-fitting model achieves a coefficient of determination of $R^2=$  \textcolor{red}{insert Rsquared}.
+ The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
 Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.

### 2020-06-22 methods.tex
work on section methods

@@
 \section{Methods description}
-In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
-the gravitational wave observations.
-An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
-on 25 1D simulations with the
-{\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code.
-Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polynomial regression with
-heteroscedastic errors
+In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius (in units of solar mass and km) from the gravitational wave observations.
+An integral part of this strategy is the universal relationship between the characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes and a combination of the mass, the radius of the PNS, the shock radius and the total mass inside the shock as demonstrated by \cite{Torres:2019}.
+In this paper we focus on the $^2g_2$ mode.
+
+Using 25 1D simulations obtained with the {\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code, we parametrize the ratio with a cubic polynomial regression with heteroscedastic errors
+%Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn
 \begin{equation}
@@
 \end{equation}
-where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with $f_i$. The model for frequency-dependent variances is
+where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent variances is
 \begin{equation}
@@
 with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
-\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The
-best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
+\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The best
+fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the Aikaike information
+criterion with coefficients given in Table \ref{tab:model},
+
 \begin{equation}\label{eq:universal}
-r_i=f_i + f_i^3 + \epsilon_i
+r_i = \beta_1 f_i + \beta_3 f_i^3 + \epsilon_i
 \end{equation}
+
 The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
-Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.
 The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
+
+\begin{table}[h]
+  \begin{tabular}{lll}
+    \hline
+    Coefficient & Estimate & standard error \\
+    \hline
+    $\beta_1$   & $6.09 \times 10^{-7}$ & $1.75 \times 10^{-8}$ \\
+    $\beta_3$   & $6.24 \times 10^{-13}$ & $8.79 \times 10^{-15}$ \\
+    \hline
+  \end{tabular}
+\label{tab:model}
+\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $^2g_2$ mode.}
+\end{table}
+
 \begin{figure}
@@
 
-Here we analyse the gravitational wave signal {\tt s20-gw-10kpc} \textcolor{red}{insert more detailed description}, originally sampled at 10 kHz but resampled to the LIGO sampling rate of 16384 Hz.
-A spectrogram of this signal is shown in Figure \textcolor{red}{insert spectrogram of signal} based on autoregressive estimates of the local spectra for successive time intervals of 
+We consider the gravitational wave signal {\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally sampled at 20 kHz but resampled to the LIGO sampling rate of 16384 Hz.
+A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on autoregressive estimates of the local spectra for successive time intervals of 
 \textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.The dominant emission mode correspondes to the $\mbox{ }^2 g_1$-mode \cite{} and we have developed a time-frequency method to track the ridge $m(t)$ in the spectrogram, taking into account that it is monotonically increasing.
@@
 \textcolor{red}{Are more details regarding the ridge tracking required here?}
+
+\begin{figure}
+ \centering
+ \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram.pdf}
+ \caption{Spectrogram} \label{fig:spec}
+\end{figure}
 

### 2020-06-24 methods.tex
add plots

@@
 \section{Methods description}
-In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius (in units of solar mass and km) from the gravitational wave observations.
-An integral part of this strategy is the universal relationship between the characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes and a combination of the mass, the radius of the PNS, the shock radius and the total mass inside the shock as demonstrated by \cite{Torres:2019}.
-In this paper we focus on the $^2g_2$ mode.
+In this section, we outline a strategy for estimating the time evolution of the
+ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
+(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
+oscillation mode in the gravitational wave detector data.
+An integral part of this strategy is the universal relationships that relate the
+characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
+and the radius of the PNS, the shock radius and the total mass inside the shock as
+demonstrated in \cite{Torres:2019}.
 
-Using 25 1D simulations obtained with the {\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code, we parametrize the ratio with a cubic polynomial regression with heteroscedastic errors
+Using 25 1D simulations obtained with the {\sc AENUS-ALCAR }code \cite{} and the
+{\sc CoCoNuT} \cite{} code, we parametrize the ratio with a cubic polynomial
+regression with heteroscedastic errors
 %Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn
+
 \begin{equation}
@@
 \end{equation}
-where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent variances is
+where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with
+variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent
+variances is
 \begin{equation}
@@
 with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
-\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The best
-fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the Aikaike information
-criterion with coefficients given in Table \ref{tab:model},
+\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.
+The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to
+the Aikaike information criterion with coefficients given in Table \ref{tab:model},
 
@@
 The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
-The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
+The data and fit of the model including 95\% confidence bands are displayed in
+Figure \ref{fig:LMVAR}.
 
@@
   \end{tabular}
-\label{tab:model}
-\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $^2g_2$ mode.}
+\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.}\label{tab:model}
 \end{table}
@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
- \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
+ \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/model}
+ \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{fig:LMVAR}
 \end{figure}
 
-We consider the gravitational wave signal {\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally sampled at 20 kHz but resampled to the LIGO sampling rate of 16384 Hz.
-A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on autoregressive estimates of the local spectra for successive time intervals of 
-\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.The dominant emission mode correspondes to the $\mbox{ }^2 g_1$-mode \cite{} and we have developed a time-frequency method to track the ridge $m(t)$ in the spectrogram, taking into account that it is monotonically increasing.
-Starting from either the left- or right-most column of the time-frequency  matrix we identify and trace the sequence of amplitude peaks within a certain frequency band given the monotonicity constraint.
-\textcolor{red}{Are more details regarding the ridge tracking required here?}
+To develop the method we considered the gravitational wave signal
+{\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally
+sampled at 16384 Hz but resampled at 4096 Hz.
+A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
+autoregressive estimates of the local spectra for successive time intervals of 
+\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
+The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode and we have
+developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
+taking into account that it is monotonically increasing as time goes.
+Starting from either the left- or right-most column of the time-frequency matrix
+we identify and trace the sequence of amplitude peaks within a certain frequency
+band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
+details on the reconstruction of the $g$ mode ridge.
 
@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram.pdf}
- \caption{Spectrogram} \label{fig:spec}
+ \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
+ \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
+   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
+   with each other.} \label{fig:spectrogram}
 \end{figure}
 
-We identify the instantaneous frequency $f(t_i)$ corresponding the  ridge $m(t_i)$ for the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. Now we can use the universal relationship in (\ref{eq:universal}) to obtain estimates of the time evolution of the
-ratios together with 95\% confidence intervals. These are given in Figure \ref{fig:xx} \textcolor{red}{insert figure} where the black points are the true ratio values, the red points the
-estimates and the grey bands represent 95\% confidence bands. In this case without any noise, the coverage of our 95\% confidence band is xx\%.
+We identify the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
+the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
+for values in between the $t_i$. We then use the equation (\ref{eq:universal}) to obtain
+estimates of the time evolution of the ratio together with 95\% confidence intervals.
+An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
+the grey bands represent 95\% confidence bands. The black points are the true ratio values
+computed using the mass and radius values obtained from the simulation code.
 
-In the following simulation study we explore how accurately we can estimate the parameters when the gravitational wave signal is embedded in noise.
-For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise using the noise power spectral density \textcolor{red}{insert formula} for varying
-SNRs, respectively distances to the source.
-We estimate the coverage probability of the 95\% confidence band by calculating the proportion of times that the true ratio lies outside one of the pointwise 95\% confidence intervals.
-These coverage probabilities together for varying SNRs are given in Table \textcolor{red}{insert Table} and displayed in the form of boxplots in Figure \textcolor{red}{insert Figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
+ \caption{} \label{fig:ratio}
+\end{figure}
 
-\subsection{G-mode reconstruction}
+%In this case without any noise, the coverage of our 95\% confidence band is xx\%.
 
-Given the spectrogram and an specified time interval for the g-mode reconstruction, our proposal method works as follows.
-The starting point must be specified.
-It can be either at the beginning or at the end of the signal.
-Then, in one of these extremes, the maximum energy value is identified, registering its frequency.
-This is done independently for a number of consecutive time intervals.
-Then we calculate the median of these frequency values, providing a robust starting value for the g-mode reconstruction.
-
-The starting frequency value is the first g-mode estimate for the first or the last time interval, depending on the specified starting location.
-If the reconstruction is set to start at the beginning of the signal, the reconstruction will be done progressively over the time intervals, where each maximum frequency value will be calculated within a frequency range specified by the previous g-mode estimate.
-Given the non-decreasing behaviour of the true g-mode values, the g-mode estimates will be forced to be greater or equal than the one estimated for its previous time interval, and lower than a specified upper limit.
-As a result, the g-modes estimates will be a non-decreasing sequence of frequency values.
-
-If the reconstruction is set to start at the end of the signal, the g-modes will be estimated backward in time.
-Each maximum frequency is calculated within a range determined by its successor (in time) g-mode estimate.
-These estimates are forced to be lower or equal than its successor (in time) estimate, but greater than a specified lower limit.
-Thus, a non-decreasing sequence of g-mode estimates is guaranteed.
-
-This g-mode reconstruction method works if and only if the signal is strong enough to provide information about the g-mode, which is reflected in the spectrogram.
-
-\subsection{ Confidence bands}
-Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:universal}.
-The g-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios.
-Since the g-mode estimates are indexed by time, the confidence intervals for the ratios are too.
-Thus, we generate the confidence band by interpolating the lower and upper limits of the collection of consecutive confidence intervals, which will be valid for the time range of the g-mode estimates.
-This confidence band is used to estimate the coverage probabilities in our simulation studies presented below.
+In the following simulation study we explore how accurately we can estimate the parameters
+when the gravitational wave signal is embedded in noise.
+For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise
+using the noise power spectral density \textcolor{red}{insert formula} for varying
+SNRs, respectively distances to the source. We estimate the coverage probability of the 95\%
+confidence band by calculating the proportion of times that the true ratio lies outside one
+of the pointwise 95\% confidence intervals.
+These coverage probabilities together for varying SNRs are given in Table
+\textcolor{red}{insert Table} and displayed in the form of boxplots in Figure
+\textcolor{red}{insert Figure}

### 2020-06-30 introduction.tex
more work in the introduction

--- 2020-03-18 introduction.tex
@@
 
-\textbf{CCSN Simulation and GW emission}\\
 The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova.
@@
 
+In the past years an impressive progress has been made in the development of numerical codes, which allows to obtain more accurate CCSN simulations.
+The waveforms produced by the magneto-rotational mechanism in particular is well understood.
+The core-bounce signal can be directly related with the rotational properties of the core \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
+However, the low rate of this kind of events and is low amplitude and high frequency of the bounce signal in the slow-rotation case will probably impede its detection.
+
+In the case of most common neutrino-driven mechanism, the GW emission is manly produced during the hydrodynamical bounce and the unstable evolution of the fluid inside the region formed by the recently formed proto-neutron star (PNS) and the accretion shock.
+The dynamics excite the different modes of oscillation of the PNS \cite{kokkotas, Friedman:2013}.
+Unluckily, in this case is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
+There are many reasons that explain this issue.
+The large number of physical processes involved whose role in the escenario is not completely understood.
+For instance, exist uncertainties in the stellar evolution of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) or the neutrino interactions.
+Furthermore, the stochastic and chaotic nature of the instabilities is transferred to the GW emission, resulting in the same progenitor leading to a significantly different waveform.
+These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
+One simulation of one single progenitor in 3D with accurate neutrino transport and realistic equation of state (EoS) can take several months of intense calculations on a scientific supercomputer facility, which complicates the systematic exploration of the progenitor parameters.
+
+Common features in the GW signal, that have been interpreted as g modes of the PNS, have been reported in many works \cite{murphy:09, Cerda:2013, mueller:13gw, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+Typically, the frequencies associated with the modes rise monotonically with time during the contraction of the PNS.
+The characteristic frequencies of the modes associated to the PNS make them promising candidates for detection in ground-based interferometers.
+
 \textbf{Mode analysis and relations}\\
-\textbf{Martin s20 simulation and code description}\\
+
 \textbf{This paper method description and previous works}\\

### 2020-06-30 appendixA.tex

\section{G-mode reconstruction}

Given the spectrogram and an specified time interval for the g-mode reconstruction, our proposal method works as follows.  The starting point must be specified.  It can be either at the beginning or at the end of the signal.  Then, in one of these extremes, the maximum energy value is identified, registering its frequency.  This is done independently for a number of consecutive time intervals.  Then we calculate the median of these frequency values, providing a robust starting value for the g-mode reconstruction.

The starting frequency value is the first g-mode estimate for the first or the last time interval, depending on the specified starting location.  If the reconstruction is set to start at the beginning of the signal, the reconstruction will be done progressively over the time intervals, where each maximum frequency value will be calculated within a frequency range specified by the previous g-mode estimate.  Given the non-decreasing behaviour of the true g-mode values, the g-mode estimates will be forced to be greater or equal than the one estimated for its previous time interval, and lower than a specified upper limit.  As a result, the g-modes estimates will be a non-decreasing sequence of frequency values.

If the reconstruction is set to start at the end of the signal, the g-modes will be estimated backward in time.  Each maximum frequency is calculated within a range determined by its successor (in time) g-mode estimate.  These estimates are forced to be lower or equal than its successor (in time) estimate, but greater than a specified lower limit. Thus, a non-decreasing sequence of g-mode estimates is guaranteed.

This g-mode reconstruction method works if and only if the signal is strong enough to provide information about the g-mode, which is reflected in the spectrogram.

Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:universal}. The g-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios. Since the g-mode estimates are indexed by time, the confidence intervals for the ratios are too.  Thus, we generate the confidence band by interpolating the lower and upper limits of the collection of consecutive confidence intervals, which will be valid for the time range of the g-mode estimates.  This confidence band is used to estimate the coverage probabilities in our simulation studies presented below.

### 2020-07-01 methods.tex
more updates

@@
 \section{Methods description}
+\label{methods}
+
 In this section, we outline a strategy for estimating the time evolution of the
@@
 oscillation mode in the gravitational wave detector data.
-An integral part of this strategy is the universal relationships that relate the
+An integral part of this strategy is the universal relations that relate the
 characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
@@
 To develop the method we considered the gravitational wave signal
-{\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally
+{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
 sampled at 16384 Hz but resampled at 4096 Hz.
@@
 
-We identify the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
+We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
 the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use the equation (\ref{eq:universal}) to obtain
+for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
@@
 
-%In this case without any noise, the coverage of our 95\% confidence band is xx\%.
-
-In the following simulation study we explore how accurately we can estimate the parameters
-when the gravitational wave signal is embedded in noise.
-For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise
-using the noise power spectral density \textcolor{red}{insert formula} for varying
-SNRs, respectively distances to the source. We estimate the coverage probability of the 95\%
-confidence band by calculating the proportion of times that the true ratio lies outside one
-of the pointwise 95\% confidence intervals.
-These coverage probabilities together for varying SNRs are given in Table
-\textcolor{red}{insert Table} and displayed in the form of boxplots in Figure
-\textcolor{red}{insert Figure}
+In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\%.
+In the next section we explore how accurately we can estimate $r(t)$ when the gravitational wave
+signal is embedded in noise.

### 2020-07-01 results.tex
more updates

@@
 \section{Detection sensitivity with Advanced gravitational wave detectors}
+\label{sec:results}
+
+To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
+gravitational wave detector data, we have added the {\tt s20-gw-10kpc} GW signal to hundred
+Gaussian noise realisations whose power spectral density follows advanced LIGO
+spectrum~\cite{aLIGOspectrum}. We have varied the distances to the source, covering a large
+range of distances for which a detection is feasible. The source is optimally oriented with
+respect to the gravitational wave detector. We are assuming a GW signal from a core collapse
+phenomena has been identified in the data and that the beginning of the GW signal is known.
+For each of the simulations, we reconstruct the ratio and compute few quantities that measure
+the reconstruction accuracy. The first quantity is the coverage probability which is the
+fraction of the ratio that falls within the 95\% confiddence interval.
+
+For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise
+using the noise power spectral density \textcolor{red}{insert formula} for varying
+SNRs, respectively distances to the source. We estimate the coverage probability of the 95\%
+confidence band by calculating the proportion of times that the true ratio lies outside one
+of the pointwise 95\% confidence intervals.
+These coverage probabilities together for varying SNRs are given in Table
+\textcolor{red}{insert Table} and displayed in the form of boxplots in Figure
+\textcolor{red}{insert Figure}

### 2020-07-06 methods.tex
Update for section results

@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/model}
- \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{fig:LMVAR}
+ \includegraphics[width=0.5\textwidth]{plots/model}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the {\sc AENUS-ALCAR} data points.} \label{fig:LMVAR}
 \end{figure}
@@
 \textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
-The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode and we have
+The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode. We have
 developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
-taking into account that it is monotonically increasing as time goes.
+taking into account that it is monotonically increasing as time goes,
+a property of the $\mbox{}^2 g_2$-mode.
 Starting from either the left- or right-most column of the time-frequency matrix
@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
+ \includegraphics[width=0.5\textwidth]{plots/spectrogram}
  \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
@@
 An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
-the grey bands represent 95\% confidence bands. The black points are the true ratio values
+the grey bands represent 95\% confidence bands. The black points are the ratio values
 computed using the mass and radius values obtained from the simulation code.
@@
  \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
- \caption{} \label{fig:ratio}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\tt s20-gw-10kpc} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
 \end{figure}

### 2020-07-06 results.tex
Update for section results

--- 2020-01-20 results.tex
@@
 \section{Detection sensitivity with Advanced gravitational wave detectors}
+\label{sec:results}
+
+To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
+gravitational wave detector data, we have added {\tt s20-gw-10kpc} GW signal to 
+Gaussian noise realisations whose power spectral density follows advanced LIGO
+spectrum~\cite{aLIGOsens:2018}. 
+We have varied the distances to the source, covering a large
+range of distances for which a detection in second generation of gravitational wave detectors
+is feasible. The source is optimally oriented with
+respect to the gravitational wave detector. We are assuming a GW signal from a core collapse
+phenomena has been identified in the data and that the beginning of the GW signal is known.
+The data (signal embedded in noise) are whitened using the function {\it prewhiten} of the R-package TSA.
+An auto-regressive model with maximal \textcolor{red}{100} coefficients has been used.
+
+For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
+of length $N$ and compute two quantities that compare {$r_i$} to ratio {$r_i^0$} derived from
+the PNS mass and radius generated by the simulation code that produces {\tt s20-gw-10kpc}.
+measure the estimation accuracy.
+The first quantity is the coverage probability ($covpbb$) which is the fraction of the
+ratio {$r_i^0$} values that fall within the 95\% confidence interval of {$r_i$}.
+We also compute the $precision$ value given by
+\begin{equation}
+precision=\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
+\end{equation}
+
+Figure \ref{fig:s20results} is showing the median of $covpbb$ and $precision$ as
+function of the distance of the source as well as the confidence belt corresponding
+to the median absolute deviation.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/s20-gw_10kpc16384_covpbb_prec}
+ \caption{$covpbb$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth.
+The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
+\end{figure}

### 2020-07-07 simulations.tex
more updates
Added a short description of the simulations in section II with a table of the 2d models and a few new references.

@@
 \section{Core collapse supernova simulations}
+\label{sec:simulations}
 
@@
 \textbf{1D simlation data to fit the ratio vs frequency model - AA and CoConut outputs}
+
+\maob{%%
+  We apply our analysis to a set of 25 spherically symmetric
+  \citep{Torres:2019} and 7 two-dimensional axisymmetric models of
+  stellar core collapse simulated with two codes, CoCoNuT
+  (one-dimensional models) and AENUS-ALCAR
+  \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
+  (one- and two-dimensional models).  CoCoNuT
+  \citep{Cerda-Duran__2008__AA__GRMHD-code} is a code for general
+  relativistic hydrodynamics coupled to the Fast Multigroup Transport
+  scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
+  description of the emission and transport of neutrinos.  AENUS-ALCAR
+  \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
+  combines special relativistic (magneto-)hydrodynamics, a modified
+  Newtonian gravitational potential approximating the effects of
+  general relativity \citep{Marek_etal__2006__AA__TOV-potential}, and
+  a spectral two-moment neutrino transport solver
+  \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
+  We included the relevant reactions between matter and neutrinos of
+  all flavours, i.e., emission and absorption by nucleons and nuclei,
+  electron-positron pair annihilation, nucleonic bremsstrahlung, and
+  scattering off nucleons, nuclei, and electrons.
+  %%
+}
+
+\maob{%%
+  Our spherically symmetric simulations cover the range of initial
+  stellar masses $M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$.  They were
+  carried out using six equations of state (EOSs).  They are
+  complemented by a set of two-dimensional axisymmetric models
+  consisting of stellar core collapse of five stars with zero-age
+  main-sequence masses of $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
+  evolved through the hydrostatic phases by
+  \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
+  We performed one simulation of each stellar model using the equation
+  of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS}
+  with an incompressibility of $K = 220 \, \mathrm{MeV}$ and added
+  comparison models with the SFHo EOS
+  \cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
+  and the EOS of
+  \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
+  for the one with $M_{\mathrm{ZAMS}} = 15 \, \Msol$ (see Table
+  \ref{Tab:2dSimList} for a list of models).
+  %%
+}
+
+\maob{%%
+  We mapped the pre-collapse state of the stars to a spherical
+  coordinate system with $n_r = 400$ zones in radial direction
+  distributed logarithmically with a minimum grid width of $(\Delta
+  r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
+  $r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and 
+  $n_{\theta} = 128$ equidistant cells in angular direction.  For the
+  neutrino energies, we used a logarithmic grid with $n_e = 10$
+  bins up to $240 \, \mathrm{MeV}$.
+  %%
+}
+
+\maob{%%
+  All spherical and most axisymmetric models fail to achieve shock
+  revival during the time of our simulations.  Only the two stars with
+  the highest masses, \texttt{s25} and \texttt{s40}, develop
+  relatively late explosions in axisymmetry.  Consequently, mass
+  accretion onto the PNSs proceeds at high rates for a long time in
+  all cases and causes them to oscillate with their characteristic
+  frequencies.  The final masses of the PNSs are in the range of
+  $M_{\mathrm{PNS}} = 1.47 - 2.33 \, \Msol$, i.e., insufficient for
+  producing a black hole.
+  %%
+}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|cc|ccc}
+    \hline
+    Simulation & $M_\mathrm{ZAMS} [\Msol]$ & EOS & $t_{\mathrm{f}} [\mathrm{s}]$&
+    explosion & $M_{\mathrm{PNS}} [\Msol]$
+    \\ 
+    \hline
+    \texttt{s11} & 11.2 & LS220 & 1.86 & $\times$ &  1.47
+    \\ 
+    \texttt{s15} & 15.0 & LS220 & 1.66 & $\times$ &  2.00
+    \\ 
+    \texttt{s15S} & 15.0 & SFHo & 1.75 & $\times$ &  2.02
+    \\ 
+    \texttt{s15G} & 15.0 & GShen & 0.97 & $\times$ &  1.86
+    \\ 
+    \texttt{s20} & 20.0 & LS220 & 1.53 & $\times$ & 1.75
+    \\ 
+    \texttt{s25} & 25.0 & LS220 & 1.60 & $0.91$ &  2.33
+    \\ 
+    \texttt{s40} & 40.0 & LS220 & 1.70 & $1.52$ &  2.23
+    \\ \hline
+  \end{tabular}
+  \caption{%%
+    List of axisymmetric simulations.  We present the name of the models, the initial
+    mass of the progenitors, and the EOS used, and the  final
+    post-bounce time of the simulations.  For models which explode,
+    we list the time at which the shock starts to expand in column
+    ``explosion''; otherwise, a $\times$ sign is displayed.  The final
+    column indicates the mass of the PNS at the end of the
+    simulation.  
+    %%
+  }
+  \label{Tab:2dSimList}
+\end{table}
+
+%%% Local Variables:
+%%% TeX-master: "ccsn"
+%%% End:

### 2020-07-14 introduction.tex
More work in the introduction

@@
 The characteristic frequencies of the modes associated to the PNS make them promising candidates for detection in ground-based interferometers.
+The presence of g-methods in hot PNS has been studied since the end of last century.
+The oscillation modes related with the surface of hot PNS was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
+Additionally, the stratified structure of the PNS allow the presence of different kinds of g-modes related with the core \cite{Reisenegger:1992}.
+Many posterior works used simplified background neutron star models (equilibrium configurations), to study different variations of the equilibrium system, like rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
+Sotani \& Takiwaki \cite{Sotani:2016} studied the oscillation modes before the explosion using a simplified fits to numerical simulations.
 
-\textbf{Mode analysis and relations}\\
+In previous works \cite{Torres:2018, Torres:2019a}, we explore the eigenmode spectrum using as background results of CCSN numerical simulations and the cavity form by the PNS and the shock.
+We showed that the GW time-frequency distribution correspond with the frequencies of oscillation of different families of p- and g-modes.
+These works reveal that is posible to perform CCSN asteroseismology and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
+In this line of research, in \cite{Torres:2019b} we derived the relations between the different types of modes with some with the evolution of mass and radius of the PNS.
+These relations are universal in the sense that they not depend of the EOS, the mass of the progenitor or the code used to perform the simulation.
 
-\textbf{This paper method description and previous works}\\
-\textbf{Previous work in identifying modes from spectrogram}\\
-\textbf{Paper organization}\\
+In this paper, we present a method to extract from the GW data the mass and the radius of the PNS as function of time using the universal relations.
+We inject a GW signal, obtained from a 2D CCSN simulation, into simulated non-white Gaussian noise which spectrum follows the AdLIGO sensitivity curve.
+We show how the algorithm is able to extract the time-frequency evolution of the main arc of GW emission, which corresponds to the $^2\rm{g}_2$, according to the nomenclature used in \cite{Torres:2019b}.
+Then the universal relation for this mode is inverted to obtain the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$.
+The evolution of the estimated ratio and the true ratio obtained from the simulation is remarkably similar.
+
+\textbf{Paper organisation}\\

### 2020-07-14 methods.tex
add refs in methods.tex

@@
 
-Using 25 1D simulations obtained with the {\sc AENUS-ALCAR }code \cite{} and the
-{\sc CoCoNuT} \cite{} code, we parametrize the ratio with a cubic polynomial
+Using 25 spherically symetric (1D) simulations obtained with the {\sc AENUS-ALCAR }code \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics} and the
+{\sc CoCoNuT} \citep{Cerda-Duran__2008__AA__GRMHD-code} code, we parametrize the ratio with a cubic polynomial
 regression with heteroscedastic errors
@@
 \textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
-The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode. We have
+The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
 developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
@@
 An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
-the grey bands represent 95\% confidence bands. The black points are the ratio values
-computed using the mass and radius values obtained from the simulation code.
+the grey bands represent 95\% confidence bands. Ratio values
+computed using the mass and radius values obtained from the simulation code are shown in black.
 
@@
 In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\%.
-In the next section we explore how accurately we can estimate $r(t)$ when the gravitational wave
+In the next section we investigate the performance of reconstruction of $r(t)$ when the gravitational wave
 signal is embedded in noise.

### 2020-08-06 methods.tex
Method section update
Table with estimates

@@
 \begin{equation}
+\label{model1}
 r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
@@
 \end{equation}
-with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
+with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}
 \cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.
+
 The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to
-the Aikaike information criterion with coefficients given in Table \ref{tab:model},
+the Aikaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{model1}.
+The data and fit of the model including 95\% confidence bands are displayed in
+Figure~\ref{fig:LMVAR}.
 
-\begin{equation}\label{eq:universal}
-r_i = \beta_1 f_i + \beta_3 f_i^3 + \epsilon_i
-\end{equation}
+%\begin{equation}\label{eq:universal}
+%r_i = \beta_1 f_i + \beta_3 f_i^3 + \epsilon_i
+%\end{equation}
 
-The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
-The data and fit of the model including 95\% confidence bands are displayed in
-Figure \ref{fig:LMVAR}.
+%The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
+%The data and fit of the model including 95\% confidence bands are displayed in
+%Figure~\ref{fig:LMVAR}.
 
 \begin{table}[h]
-  \begin{tabular}{lll}
+%  \begin{tabular}{lll}
+%    \hline
+%    Coefficient & Estimate & standard error \\
+%    \hline
+%    $\beta_1$   & $6.09 \times 10^{-7}$ & $1.75 \times 10^{-8}$ \\
+%    $\beta_3$   & $6.24 \times 10^{-13}$ & $8.79 \times 10^{-15}$ \\
+%    \hline
+%  \end{tabular}
+
+  \begin{tabular}{crr}
     \hline
-    Coefficient & Estimate & standard error \\
+    Coefficient & \multicolumn{1}{c}{Estimate} & Standard error \\
     \hline
-    $\beta_1$   & $6.09 \times 10^{-7}$ & $1.75 \times 10^{-8}$ \\
-    $\beta_3$   & $6.24 \times 10^{-13}$ & $8.79 \times 10^{-15}$ \\
+   $\beta_1$  &  $ 1.00 \times 10^{-06}$ & $2.12 \times 10^{-08}$ \\   
+   $\beta_2$  &  $-8.22 \times 10^{-10}$ & $5.00 \times 10^{-11}$ \\
+   $\beta_3$  &  $ 1.01 \times 10^{-12}$ & $2.70 \times 10^{-14}$ \\
+   $\alpha_0$ &  $-1.02 \times 10^{+01}$ & $6.80 \times 10^{-02}$ \\
+   $\alpha_1$ &  $ 7.24 \times 10^{-04}$ & $1.56 \times 10^{-04}$ \\
+   $\alpha_2$ &  $ 6.23 \times 10^{-07}$ & $8.15 \times 10^{-08}$ \\   
     \hline

### 2020-08-12 results.tex
Add results and plots

--- 2020-01-20 results.tex
@@
 \section{Detection sensitivity with Advanced gravitational wave detectors}
+\label{sec:results}
+
+To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
+gravitational wave detector data, we have added {\tt s20-gw-10kpc} GW signal to 
+Gaussian noise realisations whose power spectral density follows advanced LIGO
+spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}. 
+We have varied the distance to the source, covering a large
+range of distances for which a detection in second generation of gravitational wave detectors
+is feasible. The source is optimally oriented with
+respect to the gravitational wave detector. We are assuming a GW signal from a core collapse
+phenomena has been identified in the data and that the beginning of the GW signal is known within $O(10~ms)$.
+The data (signal embedded in noise) are whitened using the function {\it prewhiten} of the R-package TSA.
+An auto-regressive model with maximal \textcolor{red}{100} coefficients has been used.
+
+For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
+of length $N$ and compute two quantities that compare {$r_i$} to ratio {$r_i^0$} derived from
+the PNS mass and radius generated by the simulation code that produces {\tt s20-gw-10kpc}.
+
+The first quantity, $coverage$, is the fraction of the
+ratio {$r_i^0$} values that fall within the 95\% confidence interval of {$r_i$}.
+We also compute the $precision$ value given by
+\begin{equation}
+precision=\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
+\end{equation}
+
+Figure \ref{fig:s20results} is showing the median of $coverage$ and $precision$ as
+function of the distance of the source as well as the confidence belt corresponding
+to the median absolute deviation. As expected, the estimation of $r$ is maximal when
+the source is nearby and decreases with the distance. At small distance the $precision$
+value is small but not null. This reflects the approximtaion of the model used for $r$.
+It is nevertheless remarkable that one can reconstruct the ratio time series with a good
+precision at distance up to $\sim$ 10 kpc for this particular waveform. We have tested
+that the method does not depend on features of {\tt s20-gw-10kpc} using 7 other waveforms
+described in section \ref{sec:simulation} covering a large range of progenitor masses.
+Figure \ref{fig:aLIGOall} shows that apart {\tt s11.2--LS220}, the ratio is well
+reconstructed for all waveforms up to $\sim$ 10kpc.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/s20-gw_covpbb_prec_aLIGO}
+ \caption{$covpbb$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth.
+The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
+\end{figure}
+
+\begin{figure}
+ \centering
+ \includegraphics[width=0.5\textwidth]{plots/spectrum}
+ \caption{} \label{fig:spectrum}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/covppb_all_aLIGO}
+ \caption{$covpbb$ for 7 waveforms embedded in aLIGO noise at different distance from the Earth. } \label{fig:aLIGOall}
+\end{figure}

### 2020-08-12 simulations.tex

@@
 \maob{%%
-  We apply our analysis to a set of 25 spherically symmetric
-  \citep{Torres:2019} and 7 two-dimensional axisymmetric models of
+  We apply our analysis to spherically symmetric
+  \citep{Torres:2019} and two-dimensional axisymmetric models of
   stellar core collapse simulated with two codes, CoCoNuT
@@
 \maob{%%
-  Our spherically symmetric simulations cover the range of initial
-  stellar masses $M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$.  They were
-  carried out using six equations of state (EOSs).  They are
-  complemented by a set of two-dimensional axisymmetric models
-  consisting of stellar core collapse of five stars with zero-age
-  main-sequence masses of $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
-  evolved through the hydrostatic phases by
+  We use two sets of 25 models in the range of initial stellar masses
+  $M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two
+  codes.  They were carried out using six equations of state (EOSs).
+  They are complemented by a set of 7 two-dimensional axisymmetric
+  models consisting of stellar core collapse of five stars with
+  zero-age main-sequence masses of
+  $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$ evolved through the
+  hydrostatic phases by
   \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.

### 2020-08-12 results.tex
changes

@@
   \includegraphics[width=0.5\textwidth]{plots/s20-gw_covpbb_prec_aLIGO}
- \caption{$covpbb$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth.
+ \caption{$coverage$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth.
 The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
@@
   \includegraphics[width=0.5\textwidth]{plots/covppb_all_aLIGO}
- \caption{$covpbb$ for 7 waveforms embedded in aLIGO noise at different distance from the Earth. } \label{fig:aLIGOall}
+ \caption{$coverage$ for 7 waveforms embedded in aLIGO noise at different distance from the Earth. } \label{fig:aLIGOall}
 \end{figure}

### 2020-08-31 methods.tex
minor fixes

@@
 and the radius of the PNS, the shock radius and the total mass inside the shock as
-demonstrated in \cite{Torres:2019}.
+demonstrated in \cite{Torres:2019b}.
 
@@
 To develop the method we considered the gravitational wave signal
-{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
+{\texttt s20S} described in Section \ref{sec:simulations}, originally
 sampled at 16384 Hz but resampled at 4096 Hz.
@@
  \includegraphics[width=0.5\textwidth]{plots/spectrogram}
- \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
+ \caption{Spectrogram of the gravitational wave signal {\texttt s20S} sampled at \unit[4096]{Hz}.
    The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
@@
 the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
+for values in between the $t_i$. We then use equation (\eqref{model1}) to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
@@
  \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\tt s20-gw-10kpc} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
 \end{figure}

### 2020-08-31 results.tex
add SNR table
minor fixes

@@
 To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
-gravitational wave detector data, we have added {\tt s20-gw-10kpc} GW signal to 
+gravitational wave detector data, we have added {\texttt s20S} GW signal to 
 Gaussian noise realisations whose power spectral density follows advanced LIGO
@@
 of length $N$ and compute two quantities that compare {$r_i$} to ratio {$r_i^0$} derived from
-the PNS mass and radius generated by the simulation code that produces {\tt s20-gw-10kpc}.
+the PNS mass and radius generated by the simulation code that produces {\texttt s20S}.
 
@@
 precision at distance up to $\sim$ 10 kpc for this particular waveform. We have tested
-that the method does not depend on features of {\tt s20-gw-10kpc} using 7 other waveforms
+that the method does not depend on features of {\texttt s20S} using 7 other waveforms
 described in section \ref{sec:simulation} covering a large range of progenitor masses.
@@
   \includegraphics[width=0.5\textwidth]{plots/s20-gw_covpbb_prec_aLIGO}
- \caption{$coverage$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth.
+ \caption{$coverage$ and $precision$ for {\texttt s20S} signal embedded in aLIGO noise at different distance from the Earth.
 The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
@@
 \end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|cccccccc}
+    \hline
+
+    Simulation & \texttt{s11} & \texttt{s15} & \texttt{s15S} & \texttt{s15G} & \texttt{s20} & \texttt{s20S} & \texttt{s25}  & \texttt{s40}
+    \\   
+    \hline
+    aLIGO      & 19.5         & 55.4         & 59.0          &   60.0        & 34.3         &   35.8          & 116.5         & 98.5
+    \\ 
+    \hline
+    CE2
+    \\
+    \hline
+    ET-B
+    \\ 
+  \end{tabular}
+  \caption{%%
+    Matched filter signal-to-noise ratio (SNR) of the simulated waveforms for
+    the different GW detectors considered in this study. The source is located
+    at 10 kpc and is optimally oriented with respect to the detector.
+    %%
+  }
+  \label{Tab:SNR}
+\end{table}

### 2020-09-01 simulations.tex
Added the 8th 2d model to section 2 and modified the text using Marie-Anne's input.

@@
 
-\textbf{Martin's simulations and code description}\\
-\textbf{1D simlation data to fit the ratio vs frequency model - AA and CoConut outputs}
+% \textbf{Martin's simulations and code description}\\
+% \textbf{1D simlation data to fit the ratio vs frequency model - AA and CoConut outputs}
 
-\maob{%%
-  We apply our analysis to spherically symmetric
-  \citep{Torres:2019} and two-dimensional axisymmetric models of
+The algorithm proposed in the article does not require accurate
+waveforms but relies on the evolution of the frequencies of
+oscillations depending on the PNS mass and radius.  To parametrize
+this dependence, we have considered spherically symmetric
+\citep{Torres:2019a} and two-dimensional axisymmetric models of
 stellar core collapse simulated with two codes, CoCoNuT
@@
 combines special relativistic (magneto-)hydrodynamics, a modified
-  Newtonian gravitational potential approximating the effects of
-  general relativity \citep{Marek_etal__2006__AA__TOV-potential}, and
-  a spectral two-moment neutrino transport solver
+Newtonian gravitational potential approximating the effects of general
+relativity \citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
+two-moment neutrino transport solver
 \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
-  We included the relevant reactions between matter and neutrinos of
-  all flavours, i.e., emission and absorption by nucleons and nuclei,
+We included the relevant reactions between matter and neutrinos of all
+flavours, i.e., emission and absorption by nucleons and nuclei,
 electron-positron pair annihilation, nucleonic bremsstrahlung, and
 scattering off nucleons, nuclei, and electrons.
-  %%
-}
 
-\maob{%%
-  We use two sets of 25 models in the range of initial stellar masses
-  $M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two
-  codes.  They were carried out using six equations of state (EOSs).
-  They are complemented by a set of 7 two-dimensional axisymmetric
-  models consisting of stellar core collapse of five stars with
-  zero-age main-sequence masses of
-  $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$ evolved through the
-  hydrostatic phases by
+To tune the parameters of the algorithm, we use two sets of 25 models
+in the range of initial stellar masses
+$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two codes.
+They were carried out using six equations of state (EOSs).  They are
+complemented by a set of 7 two-dimensional axisymmetric models
+consisting of stellar core collapse of five stars with zero-age
+main-sequence masses of $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
+evolved through the hydrostatic phases by
 \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
 We performed one simulation of each stellar model using the equation
-  of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS}
-  with an incompressibility of $K = 220 \, \mathrm{MeV}$ and added
-  comparison models with the SFHo EOS
+of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
+an incompressibility of $K = 220 \, \mathrm{MeV}$ (LS220) and added
+comparison simulations with the SFHo EOS
 \cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
@@
 \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-  for the one with $M_{\mathrm{ZAMS}} = 15 \, \Msol$ (see Table
-  \ref{Tab:2dSimList} for a list of models).
-  %%
-}
+(GShen) for the one with $M_{\mathrm{ZAMS}} = 15 \, \Msol$ (see Table
+\ref{Tab:2dSimList} for a list of models).  To this set of
+simulations, we add the waveform of a two-dimensional model used in
+\cite{Torres:2019a}, denoted \texttt{s20S}.  It corresponds to a star
+with the same initial mass, $M_{\mathrm{ZAMS}} = 20 \, \Msol$, as for
+one of the other 7 axisymmetric simulations, but was taken from a
+newer set of stellar-evolution models
+\cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
+It was evolved with the SFHo EOS.
 
-\maob{%%
 We mapped the pre-collapse state of the stars to a spherical
 coordinate system with $n_r = 400$ zones in radial direction
-  distributed logarithmically with a minimum grid width of $(\Delta
-  r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
+distributed logarithmically with a minimum grid width of
+$(\Delta r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
 $r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and
 $n_{\theta} = 128$ equidistant cells in angular direction.  For the
-  neutrino energies, we used a logarithmic grid with $n_e = 10$
-  bins up to $240 \, \mathrm{MeV}$.
-  %%
-}
+neutrino energies, we used a logarithmic grid with $n_e = 10$ bins up
+to $240 \, \mathrm{MeV}$.
 
-\maob{%%
 All spherical and most axisymmetric models fail to achieve shock
 revival during the time of our simulations.  Only the two stars with
-  the highest masses, \texttt{s25} and \texttt{s40}, develop
-  relatively late explosions in axisymmetry.  Consequently, mass
-  accretion onto the PNSs proceeds at high rates for a long time in
-  all cases and causes them to oscillate with their characteristic
-  frequencies.  The final masses of the PNSs are in the range of
+the highest masses, \texttt{s25} and \texttt{s40}, develop relatively
+late explosions in axisymmetry.  Consequently, mass accretion onto the
+PNSs proceeds at high rates for a long time in all cases and causes
+them to oscillate with their characteristic frequencies.  The final
+masses of the PNSs are in the range of
 $M_{\mathrm{PNS}} = 1.47 - 2.33 \, \Msol$, i.e., insufficient for
 producing a black hole.
-  %%
-}
 
@@
     \texttt{s20} & 20.0 & LS220 & 1.53 & $\times$ & 1.75
+    \\ 
+    \texttt{s20S} & 20.0 & SFHo & 0.87 & $\times$ & 2.05
     \\ 

### 2020-09-02 results.tex
New plots added to results

@@
 gravitational wave detector data, we have added {\texttt s20S} GW signal to 
-Gaussian noise realisations whose power spectral density follows advanced LIGO
+100 Gaussian noise realisations whose power spectral density follows advanced LIGO
 spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}. 
@@
 For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
-of length $N$ and compute two quantities that compare {$r_i$} to ratio {$r_i^0$} derived from
+of length $N$ and compare {$r_i$} to ratio {$r_i^0$} derived from
 the PNS mass and radius generated by the simulation code that produces {\texttt s20S}.
 
-The first quantity, $coverage$, is the fraction of the
-ratio {$r_i^0$} values that fall within the 95\% confidence interval of {$r_i$}.
-We also compute the $precision$ value given by
+Figure \ref{fig:s20results} is showing the fraction of the ratio {$r_i^0$} values that fall
+within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
+when the source is located within few kpc and then decreases with the distance.
+
+To better quantity how well we reconstruct the ratio, we have considered the mean over the track
+of the relative error of $r_i$.
+
 \begin{equation}
-precision=\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
+\Delta=\frac{1}{N}\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
 \end{equation}
 
-Figure \ref{fig:s20results} is showing the median of $coverage$ and $precision$ as
-function of the distance of the source as well as the confidence belt corresponding
-to the median absolute deviation. As expected, the estimation of $r$ is maximal when
-the source is nearby and decreases with the distance. At small distance the $precision$
-value is small but not null. This reflects the approximtaion of the model used for $r$.
+$\Delta$ values of each of the 100 noise realisation are shown as well as function of the distance
+on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
+remains smaller than 20\%. At small distance $\delta$ is small but not null. This reflects the
+approximation of the model used for $r$.
 It is nevertheless remarkable that one can reconstruct the ratio time series with a good
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/s20-gw_covpbb_prec_aLIGO}
- \caption{$coverage$ and $precision$ for {\texttt s20S} signal embedded in aLIGO noise at different distance from the Earth.
-The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
+  \begin{tabular}{c}
+    \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_covpbb_boxplot_aLIGO} \\
+    \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_error_boxplot_aLIGO} \\
+  \end{tabular}
+
+ \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.} \label{fig:s20results}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{plots/covppb_all_aLIGO}
- \caption{$coverage$ for 7 waveforms embedded in aLIGO noise at different distance from the Earth. } \label{fig:aLIGOall}
+ \caption{Median of $coverage$ for 7 waveforms embedded in aLIGO noise and located at different distance from the Earth. } \label{fig:aLIGOall}
 \end{figure}

### 2020-09-03 simulations.tex
update

@@
 
-To tune the parameters of the algorithm, we use two sets of 25 models
-in the range of initial stellar masses
+We use two sets of 25 models in the range of initial stellar masses
 $M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two codes.
-They were carried out using six equations of state (EOSs).  They are
-complemented by a set of 7 two-dimensional axisymmetric models
+They were carried out using six equations of state (EOSs).
+In addition to these simulations data we have considered 8 waveforms.
+7 of them are from two-dimensional axisymmetric models
 consisting of stellar core collapse of five stars with zero-age

### 2020-09-09 introduction.tex

--- 2020-06-30 introduction.tex
@@
 The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case) by the explosion of the star as a supernova.
-Core-collapse supernova (CCSN) explosions are one most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories.
+Core-collapse supernova (CCSN) explosions are one of the expected sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories.
 This is because even the most common type of CCSN, the neutrino-driven explosion, have a rate about three per century \cite{Gossan:2016} within our galaxy.
-The other main type of explosion, those produced by the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016}.
+The other main type of explosion, the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016} \textcolor{red} Add better reference.
 However, the rate of events of this kind is much lower than the one for the neutrino driven mechanism $\sim 10^{-4} \rm{yr}^{-1}$, which represents less than $1 \%$ of all CCSNe.
-Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the explosion.
+Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the collapse.
 
-In the past years an impressive progress has been made in the development of numerical codes, which allows to obtain more accurate CCSN simulations.
+In the past years impressive progresses have been made in the development of numerical codes, which allow to obtain more accurate CCSN simulations.
 The waveforms produced by the magneto-rotational mechanism in particular is well understood.
 The core-bounce signal can be directly related with the rotational properties of the core \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
-However, the low rate of this kind of events and is low amplitude and high frequency of the bounce signal in the slow-rotation case will probably impede its detection.
+However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.
 
-In the case of most common neutrino-driven mechanism, the GW emission is manly produced during the hydrodynamical bounce and the unstable evolution of the fluid inside the region formed by the recently formed proto-neutron star (PNS) and the accretion shock.
+In the case of the neutrino-driven explosion mechanism, the GW emission is mainly produced during the hydrodynamical bounce and the unstable evolution of the fluid inside the region formed by the recently formed proto-neutron star (PNS) and the accretion shock.
 The dynamics excite the different modes of oscillation of the PNS \cite{kokkotas, Friedman:2013}.
-Unluckily, in this case is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
-There are many reasons that explain this issue.
-The large number of physical processes involved whose role in the escenario is not completely understood.
-For instance, exist uncertainties in the stellar evolution of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) or the neutrino interactions.
-Furthermore, the stochastic and chaotic nature of the instabilities is transferred to the GW emission, resulting in the same progenitor leading to a significantly different waveform.
-These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
-One simulation of one single progenitor in 3D with accurate neutrino transport and realistic equation of state (EoS) can take several months of intense calculations on a scientific supercomputer facility, which complicates the systematic exploration of the progenitor parameters.
+Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
+A large number of physical processes are involved and their role is not completely understood.
+For instance,  uncertainties in the stellar evolution models of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) of nuclear matter or the neutrino interactions.
+Furthermore, the stochastic and chaotic nature of the instabilities is transferred to the GW emission, resulting in the same progenitor leading significantly different waveforms.
+The large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
+One simulation of a single progenitor explosion  in 3D with accurate neutrino transport and realistic EoS can take several months of intense calculations on a scientific supercomputer facility.
+This complicates the systematic exploration of the progenitor parameters.
 
-Common features in the GW signal, that have been interpreted as g modes of the PNS, have been reported in many works \cite{murphy:09, Cerda:2013, mueller:13gw, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+Common features in the GW signal, that have been interpreted as gravity modes (g-modes) oscillations of the PNS, have been reported in many articles \cite{murphy:09, Cerda:2013, mueller:13gw, Yakunin:2015, Kuroda:2016, Andresen:2017}.
 Typically, the frequencies associated with the modes rise monotonically with time during the contraction of the PNS.
-The characteristic frequencies of the modes associated to the PNS make them promising candidates for detection in ground-based interferometers.
+The characteristic frequencies of the modes associated to the PNS make them promising features for detection in ground-based interferometers.
+The presence of g-modes in hot PNS has been studied since the end of last century.
+The oscillation modes related with the surface of hot PNS was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
+Additionally, the stratified structure of the PNS allows the presence of different types of g-modes related with the fluid core \cite{Reisenegger:1992}.
+Many posterior works used simplified neutron star models  assuming equilibrium configurations, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
+Sotani \& Takiwaki \cite{Sotani:2016} studied the oscillation modes before the explosion using a simplified fits to numerical simulations.
 
-\textbf{Mode analysis and relations}\\
+In previous works \cite{Torres:2018, Torres:2019a}, we explore the eigenmode spectrum using results of CCSN numerical simulations and the theoretical model of the cavity form by the center of the PNS and the shock.
+We showed that the GW time-frequency distribution corresponds with the frequencies of oscillation of different families of p- and g-modes.
+These works reveal that is posible to perform CCSN asteroseismology and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
+In this line of research, in \cite{Torres:2019b} we derived the relations between the different types of modes with some with the evolution of mass and radius of the PNS.
+These relations are universal in the sense that they not depend of the EOS, the mass of the progenitor or the code used to perform the simulation.
 
-\textbf{This paper method description and previous works}\\
-\textbf{Previous work in identifying modes from spectrogram}\\
-\textbf{Paper organization}\\
+In this paper, we present a method to extract from the GW data the mass and the radius of the PNS as function of time using the universal relations.
+We show how the algorithm is able to extract the time-frequency evolution of the main arc of GW emission, which corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b}.
+The universal relation for this mode is inverted to obtain the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$.
+Using 2D CCSN waveform corresponding to different progenitor masses we estimate teh performance of the algorithm for current and future generation of ground-based GW detectors.
+
+This paper is organised as follows.
+Section II describes the details of the 2D CCSN used.
+Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode.
+Section IV shows the performance of the data analysis method for different GW detectors.
+Finally, we discuss the results in section V.

### 2020-09-09 results.tex

@@
 gravitational wave detector data, we have added {\texttt s20S} GW signal to 
-100 Gaussian noise realisations whose power spectral density follows advanced LIGO
+100 Gaussian noise realisations whose power spectral density follows advanced LIGO (aLIGO)
 spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}. 
@@
 For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
-of length $N$ and compare {$r_i$} to ratio {$r_i^0$} derived from
-the PNS mass and radius generated by the simulation code that produces {\texttt s20S}.
+of length $N$ and compare it to ratio {$r_i^0$} derived from
+the PNS mass and radius generated by the simulation code that produced {\texttt s20S}.
 
@@
 
-To better quantity how well we reconstruct the ratio, we have considered the mean over the track
-of the relative error of $r_i$.
+To better quantity how well we reconstruct the ratio, we have also considered $\Delta$ the mean
+over the track of the relative error of $r_i$.
 
@@
 on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
-remains smaller than 20\%. At small distance $\delta$ is small but not null. This reflects the
+remains smaller than 20\%. At small distance $\Delta$ is small but not null. This reflects the
 approximation of the model used for $r$.
@@
 that the method does not depend on features of {\texttt s20S} using 7 other waveforms
-described in section \ref{sec:simulation} covering a large range of progenitor masses.
+described in section \ref{sec:simulations} covering a large range of progenitor masses.
 Figure \ref{fig:aLIGOall} shows that apart {\tt s11.2--LS220}, the ratio is well
@@
 
- \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.} \label{fig:s20results}
+ \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
+100 noise realisations is considered for each distance.} \label{fig:s20results}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/covppb_all_aLIGO}
- \caption{Median of $coverage$ for 7 waveforms embedded in aLIGO noise and located at different distance from the Earth. } \label{fig:aLIGOall}
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_allwvfs}
+ \caption{Median of $coverage$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth. } \label{fig:aLIGOall}
 \end{figure}

### 2020-09-09 results.tex
add new plots

@@
 
-To better quantity how well we reconstruct the ratio, we have also considered $\Delta$ the mean
+To better quantify how well we reconstruct the ratio, we have also considered $\Delta$ the mean
 over the track of the relative error of $r_i$.
@@
 It is nevertheless remarkable that one can reconstruct the ratio time series with a good
-precision at distance up to $\sim$ 10 kpc for this particular waveform. We have tested
-that the method does not depend on features of {\texttt s20S} using 7 other waveforms
+precision at distance up to $\sim$ 10 kpc for this particular waveform, with $coverage$ value
+larger than 80\%.
+We have tested that the method does not depend on features of {\texttt s20S} using 7 other waveforms
 described in section \ref{sec:simulations} covering a large range of progenitor masses.
-Figure \ref{fig:aLIGOall} shows that apart {\tt s11.2--LS220}, the ratio is well
-reconstructed for all waveforms up to $\sim$ 10kpc.
+Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220}, the ratio is well
+reconstructed for all waveforms up to $\sim$ 10kpc. On this figure we also show the $coverage$
+value in case of absence of signal. The median value is significantly different from 0 because
+the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
+in the spectrogram, starting between 0 and 200 Hz. In Table \ref{tab:} we are reporting the
+distance at which $coverage$ median is lower than 80\%.
 
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/aLIGO_allwvfs}
- \caption{Median of $coverage$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth. } \label{fig:aLIGOall}
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_cov_allwvfs}
+ \caption{Median of $coverage$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
+The ``no signal'' line and band show the median and first and third quartile of $coverage$ in absence of any signal.} \label{fig:aLIGO_cov_allwvf}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_prec_allwvfs}
+ \caption{Median of $Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
+The ``no signal'' line and band show the median and first and third quartile of $Delta$ in absence of any signal.} \label{fig:aLIGO_prec_allwvf}
 \end{figure}
@@
     \hline
-    aLIGO      & 19.5         & 55.4         & 59.0          &   60.0        & 34.3         &   35.8          & 116.5         & 98.5
+    aLIGO char. distance & 7     &            &                &               &   15.6           &    11.3   &  8.9             &  11.5
     \\ 
     \hline
-    CE2
+    SNR in aLIGO         & 19.5 & 55.4        & 59.0          &   60.0        & 34.3         &   35.8          & 116.5         & 98.5
     \\
     \hline
-    ET-B
+    \hline
+    CE2  max. distance  &      &            &                &               &              &                 &               & 
     \\
+    \hline
+    SNR in CE2      &      &            &                &               &              &                 &               &
+    \\
+    \hline
+    \hline
   \end{tabular}
@@
   }
-  \label{Tab:SNR}
+  \label{Tab:results}
 \end{table}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
+  \caption{Median of $coverage$ for s20.0--SFHo CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
+\end{figure}

### 2020-09-10 results.tex
updated aLIGO all wvf plots & results

@@
 For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
-of length $N$ and compare it to ratio {$r_i^0$} derived from
-the PNS mass and radius generated by the simulation code that produced {\texttt s20S}.
+of length $N$ starting from the left side of the spectrogram and constraining the beginning of the
+track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
+{$r_i^0$} derived from the PNS mass and radius generated by the simulation code that produced {\texttt s20S}.
 
@@
     \hline
-    aLIGO char. distance & 7     &            &                &               &   15.6           &    11.3   &  8.9             &  11.5
+    aLIGO char. & 7  &      26      &    25         &      21       &   16         &    11         &  9           &  12
+    \\
+    distance (kpc) & & & & & & & &
     \\ 
@@
     \hline
-    CE2  max. distance  &      &            &                &               &              &                 &               & 
+    CE2  char.  &      &            &                &               &              &                 &               & 
+    \\
+    distance (kpc) & & & & & & & &
     \\

### 2020-09-10 introduction.tex
Done with first two paragraphs of the intro

@@
+% !TEX root = ccsn.tex
+
 \section{Introduction}
 
-The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case) by the explosion of the star as a supernova.
-Core-collapse supernova (CCSN) explosions are one of the expected sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories.
-This is because even the most common type of CCSN, the neutrino-driven explosion, have a rate about three per century \cite{Gossan:2016} within our galaxy.
-The other main type of explosion, the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016} \textcolor{red} Add better reference.
-However, the rate of events of this kind is much lower than the one for the neutrino driven mechanism $\sim 10^{-4} \rm{yr}^{-1}$, which represents less than $1 \%$ of all CCSNe.
-Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the collapse.
+The life of massive stars \pcd{(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of \pcd{the} iron core under \pcd{its} own gravity, leading \pcd{to} the formation of a neutron star \pcd{(NS)} or a black hole (BH), \pcd{and} followed (typically but not necessarily in the BH case) by \pcd{a supernova} explosion.
+\pcd{Nearby} core-collapse supernova (CCSN) explosions \pcd{are expected to be sources of gravitational waves (GW) and are one of the main candidates for the next great discovery by current ground-based observatories.
+However, these are relative rare events.
+A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
+This event could be detected with advanced GW detectors within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}.
+Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
+For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful signal \pcd{that could} be detected \pcd{within $50$~kpc and for some extreme models
+up to $5-30$ Mpc \cite{Gossan:2016,TargetedSNSearchO12}.
+However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs
+\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$. Therefore, we focus this work only in neutrino-driven CCSNe. 
+Despite the low rates, CCSN are of great scientific interest because they produce} a complex GW signals which could provide significant clues about the physical processes that occur in the moments after the collapse.
 
-In the past years impressive progresses have been made in the development of numerical codes, which allow to obtain more accurate CCSN simulations.
-The waveforms produced by the magneto-rotational mechanism in particular is well understood.
-The core-bounce signal can be directly related with the rotational properties of the core \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
-However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.
+In the \pcd{last decade, a significant} progresses have been made in the development of numerical codes, \pcd{in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} \pcd{[PCD: I don't think this sentence is necessary since
+we have already discarded magneto-rotational events for this work.
+]\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
+The core-bounce signal can be directly related with the rotational properties of the core} \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
+\sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
 
-In the case of the neutrino-driven explosion mechanism, the GW emission is mainly produced during the hydrodynamical bounce and the unstable evolution of the fluid inside the region formed by the recently formed proto-neutron star (PNS) and the accretion shock.
-The dynamics excite the different modes of oscillation of the PNS \cite{kokkotas, Friedman:2013}.
-Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
+In the case of the neutrino-driven \pcd{explosions}, the GW emission is \pcd{primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface.
+This} dynamics excite the different modes of oscillation of the PNS\pcd{, which ultimately leads to the emission of GWs.
+[UNFINISHED!
+CONTINUE HERE]}Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
 A large number of physical processes are involved and their role is not completely understood.
@@
 
-In previous works \cite{Torres:2018, Torres:2019a}, we explore the eigenmode spectrum using results of CCSN numerical simulations and the theoretical model of the cavity form by the center of the PNS and the shock.
+\cite{kokkotas, Friedman:2013} In previous works \cite{Torres:2018, Torres:2019a}, we explore the eigenmode spectrum using results of CCSN numerical simulations and the theoretical model of the cavity form by the center of the PNS and the shock.
 We showed that the GW time-frequency distribution corresponds with the frequencies of oscillation of different families of p- and g-modes.

### 2020-09-11 ccsn.tex
Edits in methods, biblio,
minor changes
Move gmode reconstruction algo to appendix
small changes
Update for section results
Added a short description of the simulations in section II with a table of the 2d models and a few new references.
small modification
Done with first two paragraphs of the intro
Done with the introduction

@@
 \usepackage{graphicx}
+\graphicspath{{figures/}} % Directory in which figures are stored
+
 \usepackage{amsmath}
@@
 \usepackage{lineno}
-%\usepackage{ulem}
-%\usepackage{subfigure}
+\usepackage{units}
+\usepackage[title]{appendix}
+\usepackage{aas_macros}
+\usepackage{ulem}
+\usepackage{subfigure}
 \newcommand{\note}[1]{{\color{red}#1}}
 \newcommand{\second}{\ensuremath{\mathrm{\,s}}}
+%\newcommand{\mnras}{MNRAS}
+\def\apjl{Astrophys. J. Lett.}
+\def\prd{Phys.~Rev.~D}
+\def\aap{A\&A}
+
+\definecolor{maob}{rgb}{0.6,0.1,0.2}
+\newcommand{\maob}[1]{\textcolor{maob}{#1}}
+\newcommand{\pcd}[1]{\textcolor{blue}{#1}}
+
+\newcommand{\Msol}{M_{\odot}}
 
@@
 
-\title{Identification of protoneutron star g-modes in gravitational-wave data}
+\title{Identification of protoneutron star g-modes in gravitational-wave data
+\pcd{[PCD: alt. title]Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}}
 
@@
 
-\input{introduction}
-\input{methods}
-\input{noisestudy}
-\input{discussion}
+\input{introduction}\bigskip
+
+\input{simulations}\bigskip
+
+\input{methods}\bigskip
+
+\input{results}\bigskip
+
+\input{discussion}\bigskip
 
@@
 
+\begin{appendices}
+\input{appendixA}
+\end{appendices}
 %\bibliographystyle{unsrt}

### 2020-09-11 introduction.tex
Done with the introduction

@@
 A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
-This event could be detected with advanced GW detectors within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}.
+This event could be detected with advanced ground-based GW detectors within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}.
 Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
@@
 
-In the \pcd{last decade, a significant} progresses have been made in the development of numerical codes, \pcd{in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} \pcd{[PCD: I don't think this sentence is necessary since
-we have already discarded magneto-rotational events for this work.
-]\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
+In the \pcd{last decade, a significant progress has} been made in the development of numerical codes, \pcd{in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} \pcd{[PCD: I don't think the next sentence is necessary since
+we have already discarded magneto-rotational events for this work.]\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
 The core-bounce signal can be directly related with the rotational properties of the core} \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
 \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
+In the case of  neutrino-driven \pcd{explosions}, the GW emission is \pcd{primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+This} dynamics excite the different modes of oscillation of the PNS\pcd{, which ultimately leads to the emission of GWs.
+The frequency and time evolution of these modes carry information about the
+properties of the GW emitter and could allow to perform PNS asteroseismology. }
 
-In the case of the neutrino-driven \pcd{explosions}, the GW emission is \pcd{primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface.
-This} dynamics excite the different modes of oscillation of the PNS\pcd{, which ultimately leads to the emission of GWs.
-[UNFINISHED!
-CONTINUE HERE]}Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
+\pcd{[PCD: I would remove the discussion on the inference of the progenitor properties (commented now). ]}
+%Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
 A large number of physical processes are involved and their role is not completely understood.
@@
 Furthermore, the stochastic and chaotic nature of the instabilities is transferred to the GW emission, resulting in the same progenitor leading significantly different waveforms.
-The large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
+%The large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources.
 One simulation of a single progenitor explosion  in 3D with accurate neutrino transport and realistic EoS can take several months of intense calculations on a scientific supercomputer facility.
@@
 
-Common features in the GW signal, that have been interpreted as gravity modes (g-modes) oscillations of the PNS, have been reported in many articles \cite{murphy:09, Cerda:2013, mueller:13gw, Yakunin:2015, Kuroda:2016, Andresen:2017}.
-Typically, the frequencies associated with the modes rise monotonically with time during the contraction of the PNS.
-The characteristic frequencies of the modes associated to the PNS make them promising features for detection in ground-based interferometers.
-The presence of g-modes in hot PNS has been studied since the end of last century.
-The oscillation modes related with the surface of hot PNS was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
+\pcd{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow feature in the post bounce evolution with raising frequency
+from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
+This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \cite{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The \pcd{typical} frequencies of \pcd{these} modes make them \pcd{a} promising \pcd{source} for ground-based interferometers \pcd{(aLIGO, aVirgo, KAGRA)}.
+
+ The \pcd{properties of} g-modes in hot \pcd{PNSs} have been studied since the end of last century \pcd{by means of linear perturbation analysis of background PNS models}.
+The oscillation modes \pcd{associated to} the surface of hot PNSs was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
 Additionally, the stratified structure of the PNS allows the presence of different types of g-modes related with the fluid core \cite{Reisenegger:1992}.
-Many posterior works used simplified neutron star models  assuming equilibrium configurations, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
-Sotani \& Takiwaki \cite{Sotani:2016} studied the oscillation modes before the explosion using a simplified fits to numerical simulations.
+Many posterior works used simplified neutron star models assuming an equilibrium configuration \pcd{as a background}, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
+\pcd{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \cite{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
 
-\cite{kokkotas, Friedman:2013} In previous works \cite{Torres:2018, Torres:2019a}, we explore the eigenmode spectrum using results of CCSN numerical simulations and the theoretical model of the cavity form by the center of the PNS and the shock.
-We showed that the GW time-frequency distribution corresponds with the frequencies of oscillation of different families of p- and g-modes.
-These works reveal that is posible to perform CCSN asteroseismology and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
-In this line of research, in \cite{Torres:2019b} we derived the relations between the different types of modes with some with the evolution of mass and radius of the PNS.
-These relations are universal in the sense that they not depend of the EOS, the mass of the progenitor or the code used to perform the simulation.
+ \pcd{We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
+ using results from 2D CCSN numerical simulations as a background.
+Their results show a good match of the mode frequencies computed and the features observed in the GW spectrum of the same simulation (specially
+ when space-time perturbations are included \citep{Torres:2019a}).
+This result} reveals that it is posible to perform CCSN asteroseismology \pcd{under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
+\pcd{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the \pcd{ instantaneous frequency of the g-mode and} the mass and radius of the PNS \pcd{at each time of the evolution}.
+These relations are universal in the sense that they do not depend \pcd{on the equation of state (EOS) used} or the mass of the progenitor, \pcd{and only weakly on} the numerical code used \pcd{(see discussion in section~\ref{sec:simulations})}.
+\pcd{Similar relations have been found by \cite{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
 
-In this paper, we present a method to extract from the GW data the mass and the radius of the PNS as function of time using the universal relations.
-We show how the algorithm is able to extract the time-frequency evolution of the main arc of GW emission, which corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b}.
-The universal relation for this mode is inverted to obtain the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$.
-Using 2D CCSN waveform corresponding to different progenitor masses we estimate teh performance of the algorithm for current and future generation of ground-based GW detectors.
+\pcd{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius.
+For this purpose we have developed an algorithm to
+extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection).
+Next, we use the universal relations obtained by \cite{Torres:2019b} to infer the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$, being
+$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
 

### 2020-09-11 methods.tex
fixing ref

@@
 \begin{equation}
-\label{model1}
+\label{eq:model1}
 r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
@@
 The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to
-the Aikaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{model1}.
+the Aikaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{eq:model1}.
 The data and fit of the model including 95\% confidence bands are displayed in
@@
 the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use equation (\eqref{model1}) to obtain
+for values in between the $t_i$. We then use equation \eqref{eq:model1} to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.

### 2020-09-11 introduction.tex
Done with section II

@@
 This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection).
-Next, we use the universal relations obtained by \cite{Torres:2019b} to infer the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$, being
+Next, we use the universal relations obtained by \cite{Torres:2019b}\pcd{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
 $M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
@@
 This paper is organised as follows.
-Section II describes the details of the 2D CCSN used.
+Section II describes the details of the CCSN simulations used in the paper.
 Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode.

### 2020-09-11 simulations.tex
Done with section II

--- 2020-09-01 simulations.tex
@@
+% !TEX root = ccsn.tex
 \section{Core collapse supernova simulations}
@@
 
-The algorithm proposed in the article does not require accurate
-waveforms but relies on the evolution of the frequencies of
-oscillations depending on the PNS mass and radius.  To parametrize
-this dependence, we have considered spherically symmetric
-\citep{Torres:2019a} and two-dimensional axisymmetric models of
-stellar core collapse simulated with two codes, CoCoNuT
-(one-dimensional models) and AENUS-ALCAR
-\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
-(one- and two-dimensional models).  CoCoNuT
-\citep{Cerda-Duran__2008__AA__GRMHD-code} is a code for general
+\pcd{Unless other methods used GW astronomy, the} algorithm proposed in \pcd{this work} does not require accurate
+waveforms \pcd{in order to infer the properties of the PNS.} \pcd{Instead, it relies on the evolution of the frequency of
+oscillation of some particular modes, as seen in the GW spectrum. The frequency of these modes depends, in a universal way, 
+on the surface gravity of the PNS ($r=M_{\rm PNS}/R_{\rm PNS}^2$), in the sense that if at a given time we observe GW emission
+at a certain frequency $f$ we can determine univocally the value of the surface gravity, within a certain error, regardless of the details
+of the numerical simulation. In this work we use two sets of simulations: i) The {\it model set}, composed by 1D simulations, which is used 
+to build the universal relation (model), $r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by 2D simulations, 
+for which we know both the GW signal and the evolution of the ratio, $r (t)$, and that is used to test performance of the algorithm.}
+
+\pcd{We have used two different numerical codes in our numerical simulations.} 
+%CoCoNuT
+%(one-dimensional models) and AENUS-ALCAR
+%\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
+%(one- and two-dimensional models).  
+CoCoNuT
+\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
 relativistic hydrodynamics coupled to the Fast Multigroup Transport
@@
 
-To tune the parameters of the algorithm, we use two sets of 25 models
-in the range of initial stellar masses
-$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$ simulated with the two codes.
-They were carried out using six equations of state (EOSs).  They are
-complemented by a set of 7 two-dimensional axisymmetric models
-consisting of stellar core collapse of five stars with zero-age
-main-sequence masses of $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
+\pcd{For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
+including progenitors with zero-age main sequence (ZAMS) masses in the range 
+$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$. The set contains simulations using the
+ two numerical codes and six different equations of state. Details can be found in
+ \citep{Torres:2019a}. The reason to use one dimensional simulations for the model set
+ is that the computational cost of those is significantly smaller than the cost of multidimensional
+ simulations, so is easier to accumulate the statistics necessary to build a good model for $r(f)$.}
+
+ \begin{table}
+  \centering
+  \begin{tabular}{c|ccc|ccc}
+    \hline
+    Model & $M_\mathrm{ZAMS} $  & progenitor& EOS & $t_{\mathrm{f}}$& $t_{\rm explosion}$ & $M_{\mathrm{PNS, f}}$\\
+    name&  $[\Msol]$ & model & &  $[\mathrm{s}]$& & $[\Msol]$ 
+    \\ 
+    \hline
+    \texttt{s11} & 11.2 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.86 & $\times$ &  1.47 
+    \\ 
+    \texttt{s15} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.66 & $\times$ &  2.00 
+        \\ 
+    \texttt{s15S} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& SFHo & 1.75 & $\times$ &  2.02 
+        \\ 
+    \texttt{s15G} & 15.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& GShen & 0.97 & $\times$ &  1.86
+         \\ 
+    \texttt{s20} & 20.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.53 & $\times$ & 1.75 
+        \\ 
+    \texttt{s20S} & 20.0 & \cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity} & SFHo & 0.87 & $\times$ & 2.05 
+    \\ 
+    \texttt{s25} & 25.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.60 & $0.91$ &  2.33 
+        \\ 
+    \texttt{s40} & 40.0 & \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}& LS220 & 1.70 & $1.52$ &  2.23 
+        \\ \hline
+  \end{tabular}
+  \caption{%%
+    List of axisymmetric simulations \pcd{used for the {\it test set}}.  
+    \pcd{The last three columns show, the post-bounce time at the end of the
+    simulation, the one at the onset of the explosion (non exploding models marked
+    with $\times$), and the PNS mass at the end of the simulation.}
+    %%
+  }
+  \label{Tab:2dSimList}
+\end{table}
+
+\pcd{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the AENUS-ALCAR code
+(see Table~\ref{Tab:2dSimList} for a list of models).
+$7$ of these simulations use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
  evolved through the hydrostatic phases by
@@
 \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-(GShen) for the one with $M_{\mathrm{ZAMS}} = 15 \, \Msol$ (see Table
-\ref{Tab:2dSimList} for a list of models).  To this set of
+(GShen) for the \pcd{progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$.  To this set of
 simulations, we add the waveform of a two-dimensional model used in
@@
 
-We mapped the pre-collapse state of the stars to a spherical
+\pcd{For all the simulations,} we mapped the pre-collapse state of the stars to a spherical
 coordinate system with $n_r = 400$ zones in radial direction
@@
 
-\begin{table}
-  \centering
-  \begin{tabular}{c|cc|ccc}
-    \hline
-    Simulation & $M_\mathrm{ZAMS} [\Msol]$ & EOS & $t_{\mathrm{f}} [\mathrm{s}]$&
-    explosion & $M_{\mathrm{PNS}} [\Msol]$
-    \\ 
-    \hline
-    \texttt{s11} & 11.2 & LS220 & 1.86 & $\times$ &  1.47
-    \\ 
-    \texttt{s15} & 15.0 & LS220 & 1.66 & $\times$ &  2.00
-    \\ 
-    \texttt{s15S} & 15.0 & SFHo & 1.75 & $\times$ &  2.02
-    \\ 
-    \texttt{s15G} & 15.0 & GShen & 0.97 & $\times$ &  1.86
-    \\ 
-    \texttt{s20} & 20.0 & LS220 & 1.53 & $\times$ & 1.75
-    \\ 
-    \texttt{s20S} & 20.0 & SFHo & 0.87 & $\times$ & 2.05
-    \\ 
-    \texttt{s25} & 25.0 & LS220 & 1.60 & $0.91$ &  2.33
-    \\ 
-    \texttt{s40} & 40.0 & LS220 & 1.70 & $1.52$ &  2.23
-    \\ \hline
-  \end{tabular}
-  \caption{%%
-    List of axisymmetric simulations.  We present the name of the models, the initial
-    mass of the progenitors, and the EOS used, and the  final
-    post-bounce time of the simulations.  For models which explode,
-    we list the time at which the shock starts to expand in column
-    ``explosion''; otherwise, a $\times$ sign is displayed.  The final
-    column indicates the mass of the PNS at the end of the
-    simulation.  
-    %%
-  }
-  \label{Tab:2dSimList}
-\end{table}
-
 %%% Local Variables:

### 2020-09-11 appendixA.tex
more updates

@@
 \section{G-mode reconstruction}
-
+\label{app:gmode}
 Given the spectrogram and an specified time interval for the g-mode reconstruction, our proposal method works as follows.
@@
 
-Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:universal}.
+Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:model1}.
 The g-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios.

### 2020-09-14 introduction.tex
Blue comments removed

@@
 
-The life of massive stars \pcd{(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of \pcd{the} iron core under \pcd{its} own gravity, leading \pcd{to} the formation of a neutron star \pcd{(NS)} or a black hole (BH), \pcd{and} followed (typically but not necessarily in the BH case) by \pcd{a supernova} explosion.
-\pcd{Nearby} core-collapse supernova (CCSN) explosions \pcd{are expected to be sources of gravitational waves (GW) and are one of the main candidates for the next great discovery by current ground-based observatories.
+The life of massive stars {(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion.
+{Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GW) and are one of the main candidates for the next great discovery by current ground-based observatories.
 However, these are relative rare events.
@@
 Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
-For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful signal \pcd{that could} be detected \pcd{within $50$~kpc and for some extreme models
+For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful signal {that could} be detected {within $50$~kpc and for some extreme models
 up to $5-30$ Mpc \cite{Gossan:2016,TargetedSNSearchO12}.
@@
 
-In the \pcd{last decade, a significant progress has} been made in the development of numerical codes, \pcd{in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} \pcd{[PCD: I don't think the next sentence is necessary since
-we have already discarded magneto-rotational events for this work.]\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
+In the {last decade, a significant progress has} been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
+% {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
+%\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
 The core-bounce signal can be directly related with the rotational properties of the core} \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
 \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
-In the case of  neutrino-driven \pcd{explosions}, the GW emission is \pcd{primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
-This} dynamics excite the different modes of oscillation of the PNS\pcd{, which ultimately leads to the emission of GWs.
+In the case of  neutrino-driven {explosions}, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs.
 The frequency and time evolution of these modes carry information about the
@@
 
-\pcd{[PCD: I would remove the discussion on the inference of the progenitor properties (commented now). ]}
+%{[PCD: I would remove the discussion on the inference of the progenitor properties (commented now). ]}
 %Unluckily, in this case it is not posible to relate the GW emission with the properties (mass, rotation rate, metallicity or magnetic fields) of the progenitor stars.
@@
 
-\pcd{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow feature in the post bounce evolution with raising frequency
+{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow feature in the post bounce evolution with raising frequency
 from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
 This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \cite{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
-In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The \pcd{typical} frequencies of \pcd{these} modes make them \pcd{a} promising \pcd{source} for ground-based interferometers \pcd{(aLIGO, aVirgo, KAGRA)}.
+In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers {(aLIGO, aVirgo, KAGRA)}.
 
- The \pcd{properties of} g-modes in hot \pcd{PNSs} have been studied since the end of last century \pcd{by means of linear perturbation analysis of background PNS models}.
-The oscillation modes \pcd{associated to} the surface of hot PNSs was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
+ The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}.
+The oscillation modes {associated to} the surface of hot PNSs was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
 Additionally, the stratified structure of the PNS allows the presence of different types of g-modes related with the fluid core \cite{Reisenegger:1992}.
-Many posterior works used simplified neutron star models assuming an equilibrium configuration \pcd{as a background}, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
-\pcd{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \cite{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
+Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
+{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \cite{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
 
- \pcd{We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
+ {We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
  using results from 2D CCSN numerical simulations as a background.
@@
  when space-time perturbations are included \citep{Torres:2019a}).
-This result} reveals that it is posible to perform CCSN asteroseismology \pcd{under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
-\pcd{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the \pcd{ instantaneous frequency of the g-mode and} the mass and radius of the PNS \pcd{at each time of the evolution}.
-These relations are universal in the sense that they do not depend \pcd{on the equation of state (EOS) used} or the mass of the progenitor, \pcd{and only weakly on} the numerical code used \pcd{(see discussion in section~\ref{sec:simulations})}.
-\pcd{Similar relations have been found by \cite{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
+This result} reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
+{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the { instantaneous frequency of the g-mode and} the mass and radius of the PNS {at each time of the evolution}.
+These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}.
+{Similar relations have been found by \cite{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
 
-\pcd{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius.
+{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius.
 For this purpose we have developed an algorithm to
@@
 This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection).
-Next, we use the universal relations obtained by \cite{Torres:2019b}\pcd{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
+Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
 $M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2020-09-14 simulations.tex
Blue comments removed

@@
 
-\pcd{Unless other methods used GW astronomy, the} algorithm proposed in \pcd{this work} does not require accurate
-waveforms \pcd{in order to infer the properties of the PNS.} \pcd{Instead, it relies on the evolution of the frequency of
+{Unless other methods used GW astronomy, the} algorithm proposed in {this work} does not require accurate
+waveforms {in order to infer the properties of the PNS.} {Instead, it relies on the evolution of the frequency of
 oscillation of some particular modes, as seen in the GW spectrum. The frequency of these modes depends, in a universal way, 
@@
 
-\pcd{We have used two different numerical codes in our numerical simulations.} 
+{We have used two different numerical codes in our numerical simulations.} 
 %CoCoNuT
@@
 
-\pcd{For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
+{For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
 including progenitors with zero-age main sequence (ZAMS) masses in the range 
@@
  \caption{%%
-    List of axisymmetric simulations \pcd{used for the {\it test set}}.  
-    \pcd{The last three columns show, the post-bounce time at the end of the
+  List of axisymmetric simulations {used for the {\it test set}}. 
+  {The last three columns show, the post-bounce time at the end of the
   simulation, the one at the onset of the explosion (non exploding models marked
@@
 
-\pcd{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the AENUS-ALCAR code
+{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the AENUS-ALCAR code
 (see Table~\ref{Tab:2dSimList} for a list of models).
@@
 \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-(GShen) for the \pcd{progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$.  To this set of
+(GShen) for the {progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$. To this set of
 simulations, we add the waveform of a two-dimensional model used in
@@
 
-\pcd{For all the simulations,} we mapped the pre-collapse state of the stars to a spherical
+{For all the simulations,} we mapped the pre-collapse state of the stars to a spherical
 coordinate system with $n_r = 400$ zones in radial direction

### 2020-09-15 simulations.tex
New sentence in section II

@@
 (see Table~\ref{Tab:2dSimList} for a list of models).
+\pcd{Unlike the model set, the simulations in the test set cannot be 1D becase we need to extract the gravitational wave signal, which is
+a multi-dimensional effect.. }
 $7$ of these simulations use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$

### 2020-09-15 methods.tex
Comments on section III

@@
+% !TEX root = ccsn.tex
 \section{Methods description}
@@
 
-Using 25 spherically symetric (1D) simulations obtained with the {\sc AENUS-ALCAR }code \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics} and the
-{\sc CoCoNuT} \citep{Cerda-Duran__2008__AA__GRMHD-code} code, we parametrize the ratio with a cubic polynomial
+\pcd{To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
+spherically symetric (1D) simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
+shows the data for the $25$ numerical simulations. As identified by \cite{Torres:2019b}, the only systematic
+deviation from a single universal relation is the numerical code used in the simulations. 
+To avoid any systematic effect, we only use to build our model the $18$ simulations performed with the ALCAR-AENUS
+code, which is the same code that was used in our test set. The consequences of this choice are discussed in the
+conclusions. Using this data, we} parametrize the ratio with a cubic polynomial
 regression with heteroscedastic errors
@@
  \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the {\sc AENUS-ALCAR} data points.} \label{fig:LMVAR}
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.} \label{fig:LMVAR}
 \end{figure}
 
-To develop the method we considered the gravitational wave signal
-{\texttt s20S} described in Section \ref{sec:simulations}, originally
+\pcd{We use this model to infer the properties of the simulations in the 
+{\it test set} described in in Section \ref{sec:simulations}.}
+To develop \pcd{and describe} the method we \pcd{\sout{considered} focus on} the gravitational wave signal
+\pcd{of} {\texttt s20S}, originally
 sampled at 16384 Hz but resampled at 4096 Hz.
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
-autoregressive estimates of the local spectra for successive time intervals of 
+autoregressive estimates \pcd{[CITATION?]}of the local spectra for successive time intervals of 
 \textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
@@
 developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
-taking into account that it is monotonically increasing as time goes,
-a property of the $\mbox{}^2 g_2$-mode.
+taking into account that it is monotonically increasing as time goes. 
+\pcd{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
+increases as the object becomes more massive and compact}.
 Starting from either the left- or right-most column of the time-frequency matrix
@@
 the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use equation \eqref{eq:model1} to obtain
+for values in between the $t_i$. We then use \pcd{our model given by Eq.~}
+\eqref{eq:model1} to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
@@
 the grey bands represent 95\% confidence bands. Ratio values
-computed using the mass and radius values obtained from the simulation code are shown in black.
+computed using the mass and radius values obtained from the simulation code \pcd{(true values)} are shown in black.
 
@@
 
-In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\%.
-In the next section we investigate the performance of reconstruction of $r(t)$ when the gravitational wave
+In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\% \pcd{of the true values}.
+In the next section we investigate the performance of \pcd{the} reconstruction of $r(t)$ when the gravitational wave
 signal is embedded in noise.

### 2020-09-15 simulations.tex
Comments on section III

--- 2020-09-14 simulations.tex
@@
  simulations, so is easier to accumulate the statistics necessary to build a good model for $r(f)$.}
+ \pcd{For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
+ described in \cite{Torres:2018,Torres:2019a,Torres:2019b}. }
 
@@
 to $240 \, \mathrm{MeV}$.
+\pcd{Unlike the model set, the simulations in the test set cannot be 1D becase we need to 
+extract the gravitational wave signal, which is a multi-dimensional effect. For each simulation
+we extract the gravitational wave signal, $h_+(t)$,  by means of the quadrupole formula and compute the 
+time evolution of the surface gravity, $r(t)$.}
 

### 2020-09-15 methods.tex
Comments on the rest

@@
 {\it test set} described in in Section \ref{sec:simulations}.}
-To develop \pcd{and describe} the method we \pcd{\sout{considered} focus on} the gravitational wave signal
+To \pcd{\sout{develop}  describe} the method we \pcd{\sout{considered} focus on} the gravitational wave signal
 \pcd{of} {\texttt s20S}, originally

### 2020-09-15 results.tex
update numbers and plots
Comments on the rest

@@
+% !TEX root = ccsn.tex
 \section{Detection sensitivity with Advanced gravitational wave detectors}
@@
 To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
-gravitational wave detector data, we have added {\texttt s20S} GW signal to 
-100 Gaussian noise realisations whose power spectral density follows advanced LIGO (aLIGO)
+gravitational wave detector data, we have added \pcd{the} GW signal \pcd{from {\texttt s20S}} to 
+100 Gaussian noise realisations whose power spectral density follows \pcd{the} advanced LIGO (aLIGO)
 spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}. 
@@
 track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
-{$r_i^0$} derived from the PNS mass and radius generated by the simulation code that produced {\texttt s20S}.
+{$r_i^0$} derived from the PNS mass and radius \pcd{computed from the \sout{generated by the simulation code that produced} {\texttt s20S} simulation}.
 
@@
 on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
-remains smaller than 20\%. At small distance $\Delta$ is small but not null. This reflects the
+remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance $\Delta$ is small but not null. This reflects the
 approximation of the model used for $r$.
@@
 larger than 80\%.
-We have tested that the method does not depend on features of {\texttt s20S} using 7 other waveforms
+We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7 other waveforms \pcd{of the {\it test set}}
 described in section \ref{sec:simulations} covering a large range of progenitor masses.
@@
 the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
-in the spectrogram, starting between 0 and 200 Hz. In Table \ref{tab:} we are reporting the
+in the spectrogram, starting between 0 and 200 Hz. In Table \ref{tab:results} we are reporting the
 distance at which $coverage$ median is lower than 80\%.
@@
  \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
-100 noise realisations is considered for each distance.} \label{fig:s20results}
+100 noise realisations is considered for each distance.
+ \pcd{[PCD: I guess you use "coverage probability" because we are showing the box plots. Shouln't we say $\Delta$ probability as well? Or just "coverage"? Also, 
+ how are outliers computed? Why are there so many outliers in $|delta$? If they are a significant fraction of the data points why are they considered outliers?]}}
+  \label{fig:s20results}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/aLIGO_cov_allwvfs}
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_coverage_allwvfs}
  \caption{Median of $coverage$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/aLIGO_prec_allwvfs}
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
  \caption{Median of $Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
@@
   \centering
-  \begin{tabular}{c|cccccccc}
+  \begin{tabular}{c|c|cccccccc}
+
+\multicolumn{2}{c|}{}  & \texttt{s11} & \texttt{s15} & \texttt{s15S} & \texttt{s15G} & \texttt{s20} & \texttt{s20S} & \texttt{s25}  & \texttt{s40}\\
+
+\hline
+\multirow{2}{*}{aLIGO} & $d_{max}$ (kpc) & 7.3  & 26.3 & 24.8 & 22.3 & 16.3 & 10.8 & 7.8  & 14.8\\
+\cline{2-10}
+                       & SNR            & 13.0 & 37.0 & 38.7 & 39.2 & 22.9 & 23.3 & 75.6 & 63.6\\
+\hline
 \hline
 
-    Simulation & \texttt{s11} & \texttt{s15} & \texttt{s15S} & \texttt{s15G} & \texttt{s20} & \texttt{s20S} & \texttt{s25}  & \texttt{s40}
-    \\   
-    \hline
-    aLIGO char. & 7  &      26      &    25         &      21       &   16         &    11         &  9           &  12
-    \\
-    distance (kpc) & & & & & & & &
-    \\ 
-    \hline
-    SNR in aLIGO         & 19.5 &    55.4    &    59.0       &      60.0     &   34.3       &    35.8          & 116.5         & 98.5
-    \\
-    \hline
-    \hline
-    CE2  char.  &      &            &                &               &              &                 &               & 
-    \\
-    distance (kpc) & & & & & & & &
-    \\
-    \hline
-    SNR in CE2      &      &            &                &               &              &                 &               &
-    \\
-    \hline
-    \hline
   \end{tabular}
@@
   }
-  \label{Tab:results}
+  \label{tab:results}
 \end{table}

### 2020-09-23 results.tex
add dmax

@@
 \hline
-\multirow{2}{*}{aLIGO} & $d_{max}$ (kpc) & 7.3  & 26.3 & 24.8 & 22.3 & 16.3 & 10.8 & 7.8  & 14.8\\
+\multirow{2}{*}{aLIGO} & $d_{max}$ & 7  & 26 & 24 & 22 & 16 & 10 & 13 & 28\\
 \cline{2-10}
-                       & SNR            & 13.0 & 37.0 & 38.7 & 39.2 & 22.9 & 23.3 & 75.6 & 63.6\\
+                       & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\
 \hline
 \hline
+\multirow{2}{*}{CE1} & $d_{max}$ &  75  & 285 & 279 & 230 & 179 & 111 & 164  & 349\\
+\cline{2-10}
+                       & SNR    &  149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\
+
+\hline
+\multirow{2}{*}{CE2} & $d_{max}$ & 147  & -- & -- & -- & -- & 207 & --  & --\\
+\cline{2-10}
+                       & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\
+
+\hline
+\multirow{2}{*}{ET\_B} & $d_{max}$ & 75  & 278 & 274 & 244 & 159 & 107 & 79   & 62\\
+\cline{2-10}
+                       & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\
+
+\hline
+\multirow{2}{*}{ET\_C} & $d_{max}$ & 71  & 271 & 239 & 222 & 154 & 119 & 104  & 165\\
+\cline{2-10}
+                       & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\
+
+\hline
+\multirow{2}{*}{ET\_D} & $d_{max}$ & 79  & 306 & 264 & 252 & 179 & 131 & 134  & 349\\
+\cline{2-10}
+                       & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\
 
@@
   \caption{%%
+    Maximal distance $d_{max}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
+    with good accuracy for a source optimally oreinted with respect to the GW detectors
+    considered in this study.
     Matched filter signal-to-noise ratio (SNR) of the simulated waveforms for
-    the different GW detectors considered in this study. The source is located
-    at 10 kpc and is optimally oriented with respect to the detector.
+    the different GW detectors. The source is located at 10 kpc and is optimally oriented
+    with respect to the detector.
     %%

### 2020-09-23 methods.tex
update

--- 2020-09-15 methods.tex
@@
 deviation from a single universal relation is the numerical code used in the simulations. 
-To avoid any systematic effect, we only use to build our model the $18$ simulations performed with the ALCAR-AENUS
+To avoid any systematic effect, we only use the $18$ simulations performed with the ALCAR-AENUS
 code, which is the same code that was used in our test set. The consequences of this choice are discussed in the
@@
 {\it test set} described in in Section \ref{sec:simulations}.}
-To develop \pcd{and describe} the method we \pcd{\sout{considered} focus on} the gravitational wave signal
-\pcd{of} {\texttt s20S}, originally
+To describe the method we focus on the gravitational wave signal
+of {\texttt s20S}, originally
 sampled at 16384 Hz but resampled at 4096 Hz.
@@
  \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points) \mab{To be redone in python}.} \label{fig:ratio}
 \end{figure}

### 2020-09-23 simulations.tex
update

@@
 \cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
-and the EOS of
+and the GShen EOS 
 \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-(GShen) for the {progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$. To this set of
+for the {progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$. To this set of
 simulations, we add the waveform of a two-dimensional model used in
@@
 to $240 \, \mathrm{MeV}$.
-\pcd{Unlike the model set, the simulations in the test set cannot be 1D becase we need to 
+\pcd{Unlike the model set, the simulations in the test set are not 1D because we need to 
 extract the gravitational wave signal, which is a multi-dimensional effect. For each simulation

### 2020-09-29 introduction.tex
Update

@@
 extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
-This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection).
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convention).
 Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
@@
 Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode.
-Section IV shows the performance of the data analysis method for different GW detectors.
+Section IV shows the performance of the data analysis method with simulated GW detectors data.
 Finally, we discuss the results in section V.

### 2020-09-29 results.tex
Update

--- 2020-09-15 results.tex
@@
 track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
-{$r_i^0$} derived from the PNS mass and radius \pcd{computed from the \sout{generated by the simulation code that produced} {\texttt s20S} simulation}.
+{$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
 
-Figure \ref{fig:s20results} is showing the fraction of the ratio {$r_i^0$} values that fall
+Figure \ref{fig:s20results} is showing the distribution of the fraction of the ratio {$r_i^0$} values that fall
 within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
@@
 on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
-remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance $\Delta$ is small but not null. This reflects the
-approximation of the model used for $r$.
+remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance
+$\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
 It is nevertheless remarkable that one can reconstruct the ratio time series with a good
-precision at distance up to $\sim$ 10 kpc for this particular waveform, with $coverage$ value
+precision at distance up to $\sim$ 9 kpc for this particular waveform, with $coverage$ value
 larger than 80\%.
-We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7 other waveforms \pcd{of the {\it test set}}
-described in section \ref{sec:simulations} covering a large range of progenitor masses.
+We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7
+other waveforms \pcd{of the {\it test set}} described in section \ref{sec:simulations} covering
+a large range of progenitor masses.
 Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220}, the ratio is well
-reconstructed for all waveforms up to $\sim$ 10kpc. On this figure we also show the $coverage$
-value in case of absence of signal. The median value is significantly different from 0 because
-the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
-in the spectrogram, starting between 0 and 200 Hz. In Table \ref{tab:results} we are reporting the
-distance at which $coverage$ median is lower than 80\%.
+reconstructed for all waveforms up to $\sim$ 9kpc.
+To measure the level of contamination of $coverage$ by reconstructed ratios not associated to a
+signal, we have run 100 simulations without injecting a signal. The median of $coverage$ as well as
+the 95 quartile are shown on Figure \ref{fig:aLIGO_cov_allwvf}. The noise only median value is
+significantly different from zero because the g-mode reconstruction algorithm is looking for a
+continuously frequency increasing track in the spectrogram, starting between 0 and 200 Hz.
+Outliers can also reach values as high as 80\%.
+Figure \ref{fig:aLIGO_prec_allwvf} shows $\Delta$ as function of the distance for the same signals
+as well as the result when only noise is considered.
+In Table \ref{tab:results} we are reporting the distance at which $coverage$ median is lower than
+95\% of the noise only values. These numbers are an estimate of the order of magnitude of the 
+source maximal distance at which a reconstruction of the ratio could be possible with current
+and future generations of gravitational wave detectors. We have checked that $coverage$ and $\Delta$
+provide similar values.
+
+\mab{add comment about ADV}
 
@@
 
+The same analysis has been performed using expected sensitivity curves for the third generation of
+gravitational wave detectors. The US based project Cosmic Explorer~\cite{} is proposing to reach out
+its design sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in
+Figure~\ref{fig:spectrum}. In Europe, the Einstein Telescope is composed of three nested interferometers
+in a configuration of an equilateral triange~\cite{}.
+
+\mab{explain ET config}
+\mab{comment figure 9}
+
 \begin{table}
@@
 \hline
-\multirow{2}{*}{aLIGO} & $d_{max}$ (kpc) & 7.3  & 26.3 & 24.8 & 22.3 & 16.3 & 10.8 & 7.8  & 14.8\\
+\multirow{3}{*}{aLIGO} & $d_{r}$ & 7   & 27  & 24  & 22  & 16  & 10  & 37  & 45 \\
 \cline{2-10}
-                       & SNR            & 13.0 & 37.0 & 38.7 & 39.2 & 22.9 & 23.3 & 75.6 & 63.6\\
+                       & $d_{det}$ & 11 & 36 & 26 & 27 & 21 & 16 & 74 & 61\\
+                       & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\
+
 \hline
 \hline
+\multirow{3}{*}{ADV}   & $d_{r}$ & 6   & 25  & 20  & 19  & 14  & 10  & 43  & 42 \\
+\cline{2-10}
+                       & $d_{det}$ &  &  &  &  &  &  &  & \\
+                       & SNR      &  &  &  &  &  &  &  & \\
+
+\hline
+\hline
+\multirow{2}{*}{CE1}   & $d_{r}$   & 79  & 303 & 257 & 229 & 186 & 115 & 523 & 490 \\
+\cline{2-10}
+                       & $d_{det}$ & 115 & 377 & 270 & 282 & 217 & 168 & 774  & 633\\
+                       & SNR      & 149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\
+
+\hline
+\multirow{2}{*}{CE2}   & $d_{r}$  & 135 & 518 & 464 & 405 & 305 & 183 & 391 & 700 \\
+\cline{2-10}
+                       & $d_{det}$ & 197 & 649 & 468 & 489 & 375 & 294 & 1347  & 1100\\
+                       & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\
+
+\hline
+\multirow{2}{*}{ET\_B} & $d_{r}$ & 71  & 293 & 248 & 245 & 158 & 123 & 113 & 392 \\
+\cline{2-10}
+                       & $d_{det}$ & 106 & 364 & 274 & 391 & 216 & 200 & 805 & 665\\
+                       & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\
+
+\hline
+\multirow{2}{*}{ET\_C} & $d_{r}$ & 75  & 302 & 239 & 237 & 172 & 131 & 239 & 446 \\
+\cline{2-10}
+                       & $d_{det}$ & 97 & 332 & 246 & 260 & 194 & 164 & 727  & 603\\
+                       & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\
+
+\hline
+\multirow{2}{*}{ET\_D} & $d_{r}$ & 83  & 329 & 257 & 261 & 186 & 139 & 369 & 523 \\
+\cline{2-10}
+                       & $d_{det}$ & 107 & 368 & 271 & 285 & 213 & 174 & 796  & 661\\
+                       & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\
 
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
+  \caption{} \label{fig:distances}
+\end{figure}

### 2020-09-30 results.tex
add ref+update text

@@
 The same analysis has been performed using expected sensitivity curves for the third generation of
-gravitational wave detectors. The US based project Cosmic Explorer~\cite{} is proposing to reach out
-its design sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in
+gravitational wave detectors. The US based project Cosmic Explorer~\cite{reitze2019cosmic} is proposing
+to reach out its design sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in
 Figure~\ref{fig:spectrum}. In Europe, the Einstein Telescope is composed of three nested interferometers
-in a configuration of an equilateral triange~\cite{}.
+in a configuration of an equilateral triange~\cite{Hild_2011}.
 
@@
                        & $d_{det}$ & 11 & 36 & 26 & 27 & 21 & 16 & 74 & 61\\
-                       & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\
+%                       & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\
 
@@
                        & $d_{det}$ &  &  &  &  &  &  &  & \\
-                       & SNR      &  &  &  &  &  &  &  & \\
+%                       & SNR      &  &  &  &  &  &  &  & \\
 
@@
                        & $d_{det}$ & 115 & 377 & 270 & 282 & 217 & 168 & 774  & 633\\
-                       & SNR      & 149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\
+%                       & SNR      & 149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\
 
@@
                        & $d_{det}$ & 197 & 649 & 468 & 489 & 375 & 294 & 1347  & 1100\\
-                       & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\
+%                       & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\
 
@@
                        & $d_{det}$ & 106 & 364 & 274 & 391 & 216 & 200 & 805 & 665\\
-                       & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\
+%                       & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\
 
@@
                        & $d_{det}$ & 97 & 332 & 246 & 260 & 194 & 164 & 727  & 603\\
-                       & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\
+%                       & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\
 
@@
                        & $d_{det}$ & 107 & 368 & 271 & 285 & 213 & 174 & 796  & 661\\
-                       & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\
+%                       & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\
 
@@
   \caption{%%
-    Maximal distance $d_{max}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
+    Maximal distance $d_{r}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
     with good accuracy for a source optimally oreinted with respect to the GW detectors
-    considered in this study.
-    Matched filter signal-to-noise ratio (SNR) of the simulated waveforms for
-    the different GW detectors. The source is located at 10 kpc and is optimally oriented
-    with respect to the detector.
+    considered in this study. $d_{det}$ is the distance at which one could detect a source
+    optimally oriented with a matched filter signal-to-noise ratio of 13 in the  different
+    GW detectors.
     %%

### 2020-10-09 introduction.tex
update with final results

@@
 The life of massive stars {(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion.
-{Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GW) and are one of the main candidates for the next great discovery by current ground-based observatories.
+{Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GWs) and are one of the main candidates for the next great discovery by current ground-based observatories.
 However, these are relative rare events.
 A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
-This event could be detected with advanced ground-based GW detectors within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}.
-Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
-For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful signal {that could} be detected {within $50$~kpc and for some extreme models
+\mab{The emitted GWs} could be detected with advanced ground-based GW detectors \mab{(
+  Advanced LIGO\citep{TheLIGOScientific:2014jea},
+  Advanced Virgo\citep{TheVirgo:2014hva} and
+  KAGRA\citep{Aso:2013eba})} within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
+For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful \mab{GW} signal {that could} be detected {within $50$~kpc and for some extreme models
 up to $5-30$ Mpc \cite{Gossan:2016,TargetedSNSearchO12}.
 However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs
-\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$. Therefore, we focus this work only in neutrino-driven CCSNe. 
+\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{add ref?}.
+Therefore, we focus this work only in neutrino-driven CCSNe.
 Despite the low rates, CCSN are of great scientific interest because they produce} a complex GW signals which could provide significant clues about the physical processes that occur in the moments after the collapse.
 
-In the {last decade, a significant progress has} been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
+In the last decade, significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
 % {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
@@
 \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
-In the case of  neutrino-driven {explosions}, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
 This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs.
@@
 
-{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow feature in the post bounce evolution with raising frequency
+{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow {\mab oscillation} in the post bounce evolution with raising frequency
 from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
 This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \cite{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
-In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers {(aLIGO, aVirgo, KAGRA)}.
+In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.
 
@@
 
- {We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
- using results from 2D CCSN numerical simulations as a background.
-Their results show a good match of the mode frequencies computed and the features observed in the GW spectrum of the same simulation (specially
- when space-time perturbations are included \citep{Torres:2019a}).
-This result} reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
+ % {We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored
+\mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
+using results from 2D CCSN numerical simulations as a background studied in \cite{Torres:2018, Torres:2019a}
+show a good match of the mode frequencies computed and the features observed in the GW spectrum of the same
+simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
+This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
 {\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the { instantaneous frequency of the g-mode and} the mass and radius of the PNS {at each time of the evolution}.

### 2020-10-09 results.tex
update with final results

@@
 spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}.
+
+\begin{figure}
+ \centering
+ \includegraphics[width=0.5\textwidth]{plots/spectrum}
+ \caption{Amplitude spectral density of the GW detectors advanced LIGO (aLIGO) and Advanced Virgo (ADV) at design sensitivity and of the proposed third generation detectors Cosmic Explorer and Einstein Telescope.
+Eintein Telescope sensitivity curve ET-B is obtained pushing second generation detector technology at its limit.
+ET-C and ET-D sensitivity curves correspond to a detector configuration where a low-power cryogenic low-frequency interferometer and a high-power room temperature high-frequency interferometer are sharing the same infrastructure~\cite{Hild_20211}.
+Cosmic Explorer design sensitivity will be achieved in two stages.
+Stage 1 (CE1) is expected to use the technology developed for the “A+” upgrade to advanced LIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
+} \label{fig:spectrum}
+\end{figure}
+
 We have varied the distance to the source, covering a large
@@
 
-Figure \ref{fig:s20results} is showing the distribution of the fraction of the ratio {$r_i^0$} values that fall
-within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
+Figure \ref{fig:s20results} is showing the distribution of the fraction of the ratio {$r_i^0$} values
+that fall within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
 when the source is located within few kpc and then decreases with the distance.
@@
 $\Delta$ values of each of the 100 noise realisation are shown as well as function of the distance
-on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
-remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance
-$\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
-It is nevertheless remarkable that one can reconstruct the ratio time series with a good
+on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[9]{kpc} the relative error
+remains smaller than 20\%.
+At small distance, $\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
+It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
 precision at distance up to $\sim$ 9 kpc for this particular waveform, with $coverage$ value
-larger than 80\%.
-We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7
-other waveforms \pcd{of the {\it test set}} described in section \ref{sec:simulations} covering
-a large range of progenitor masses.
-Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220}, the ratio is well
-reconstructed for all waveforms up to $\sim$ 9kpc.
-To measure the level of contamination of $coverage$ by reconstructed ratios not associated to a
-signal, we have run 100 simulations without injecting a signal. The median of $coverage$ as well as
-the 95 quartile are shown on Figure \ref{fig:aLIGO_cov_allwvf}. The noise only median value is
-significantly different from zero because the g-mode reconstruction algorithm is looking for a
-continuously frequency increasing track in the spectrogram, starting between 0 and 200 Hz.
-Outliers can also reach values as high as 80\%.
-Figure \ref{fig:aLIGO_prec_allwvf} shows $\Delta$ as function of the distance for the same signals
-as well as the result when only noise is considered.
-In Table \ref{tab:results} we are reporting the distance at which $coverage$ median is lower than
-95\% of the noise only values. These numbers are an estimate of the order of magnitude of the 
-source maximal distance at which a reconstruction of the ratio could be possible with current
-and future generations of gravitational wave detectors. We have checked that $coverage$ and $\Delta$
-provide similar values.
-
-\mab{add comment about ADV}
+larger than 80\%. There are few noise realisations for a source located at $<$ \unit[9]{kpc} for which
+$\Delta$ takes large values, indicating that the method start failing to reconstruct with accuracy the ratio.
 
@@
  \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
-100 noise realisations is considered for each distance.
- \pcd{[PCD: I guess you use "coverage probability" because we are showing the box plots. Shouln't we say $\Delta$ probability as well? Or just "coverage"? Also, 
- how are outliers computed? Why are there so many outliers in $|delta$? If they are a significant fraction of the data points why are they considered outliers?]}}
+100 noise realisations is considered for each distance.}
   \label{fig:s20results}
@@
 
-\begin{figure}
- \centering
- \includegraphics[width=0.5\textwidth]{plots/spectrum}
- \caption{} \label{fig:spectrum}
-\end{figure}
+We have tested that the method does not depend on features of {\texttt s20S} using the 7
+other waveforms of the {\it test set} described in section \ref{sec:simulations} covering
+a large range of progenitor masses.
+
+Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220} and to a lesser extent
+{\tt s20.0--SFHo}, the ratio is well reconstructed for all waveforms up to $\sim$ 15kpc. In an
+effort to better determine the maximal distance of the source at which we can reconstruct the ratio
+we have run 100 simulations without injecting a signal and have measured $coverage$ for the
+reconstructed ratios.
+The median of $coverage$ as well as the 95 quartile are shown on Figure \ref{fig:aLIGO_cov_allwvf}.
+The noise only median value is null in this case, but it can be different from zero because
+the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
+in the spectrogram, starting between 0 and 200 Hz, where we expect the GW signal to be.
+This is enhancing the probability of overlap. This effects explains why outliers can reach
+values as high as 80\%.
+Figure \ref{fig:aLIGO_prec_allwvf} shows $\Delta$ as function of the distance for the same signals
+as well as the result when only noise is considered. 
+In Table \ref{tab:results} we are reporting the distance $d_r$ at which $coverage$ median is lower
+than 95\% of the noise only values. We have checked that $coverage$ and $\Delta$ provide similar
+values.
+These numbers are an estimate of the order of magnitude of the source maximal distance at which a
+reconstruction of the ratio could be possible with current GW detectors.
+They are also upper limits as we are taking into account the detector antenna response in our
+simulation but consider the source is optimally oriented. 
+Table \ref{tab:results} reports also $d_r$ for the advanced Virgo detector at design sensitivity.
+Results are very similar to aLIGO, despite the detector sensitivity differences.
+Note that Table \ref{tab:results} provides the distance at which one could detect a source
+optimally oriented with a matched filter signal-to-noise ratio of 13.
 
@@
 The same analysis has been performed using expected sensitivity curves for the third generation of
-gravitational wave detectors. The US based project Cosmic Explorer~\cite{reitze2019cosmic} is proposing
-to reach out its design sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in
-Figure~\ref{fig:spectrum}. In Europe, the Einstein Telescope is composed of three nested interferometers
-in a configuration of an equilateral triange~\cite{Hild_2011}.
+gravitational wave detectors. In Europe the Einstein Telescope project proposes to host in a 10-km
+equilateral triangle configuration 3 low-power low-frequency cryogenic interferometers as well as 3
+high-power high-frequency interferometers. Three sensitivity curves, ET-B, ET-C and ET-D corresponding
+to different options and stages of the project~\cite{Hild_2011} are considered in this study.
+The US based project Cosmic Explorer~\cite{reitze2019cosmic} is proposing to reach its design
+sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in Figure~\ref{fig:spectrum}.
 
-\mab{explain ET config}
-\mab{comment figure 9}
+Figure \ref{fig:s20--SFHo_all3G} shows $\Delta$ as function of the source distance for {\tt s20.0--SFHo}
+waveform for the 5 3G detectors configurations. Overall, the ratio is well reconstructed up to distances
+in the range \unit{100--200}[kpc] which represents an improved of at least a factor 10 with respect to
+advanced LIGO and advanced Virgo detectors. We can also note that the Einstein Telescope results lay in
+between the 2 Cosmic Explorer results. This is confirmed for all other waveforms, expect {\tt s25.0-LS220}
+for which the maximal distance reach in CE2 is significantly lower than CE1. This is partly due to the
+small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$
+rather incertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
+\ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
+located up to several hundred of kpc. It is nevertheless important to note the rather large range obtained
+for the different waveforms probing a large range of progenitor masses.
+
+\mab{add comment on the mass progenitor dependence}\\
+\mab{say that the main depend is the SNR as shown by ddet}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
+  \caption{Median of $\Delta$ for s20.0--SFHo CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
+  \caption{Maximal distance $d_{r}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
+    with good accuracy for a source optimally oriented with respect to the GW detectors for the 7 CCSN waveforms considered in this study.} \label{fig:distances}
+\end{figure}
 
@@
 \hline
-\multirow{3}{*}{aLIGO} & $d_{r}$ & 7   & 27  & 24  & 22  & 16  & 10  & 37  & 45 \\
+\multirow{3}{*}{aLIGO} & $d_{r}$   & 7 & 28 & 24  & 22 & 16 & 11 & 38 & 46 \\
 \cline{2-10}
@@
 \hline
-\multirow{3}{*}{ADV}   & $d_{r}$ & 6   & 25  & 20  & 19  & 14  & 10  & 43  & 42 \\
+\multirow{3}{*}{ADV}   & $d_{r}$   & 7  & 26  & 20 & 19 & 15 & 10 & 43 & 42 \\
 \cline{2-10}
-                       & $d_{det}$ &  &  &  &  &  &  &  & \\
-%                       & SNR      &  &  &  &  &  &  &  & \\
+                       & $d_{det}$ &  10 & 32 & 22 & 23 & 18 & 13 & 64 & 52\\
+%                      & SNR      &  13 & 41 & 29 & 30 & 23 & 17 &  83 & 68\\
 
@@
 \hline
-\multirow{2}{*}{CE1}   & $d_{r}$   & 79  & 303 & 257 & 229 & 186 & 115 & 523 & 490 \\
+\multirow{2}{*}{CE1}   & $d_{r}$   & 79  & 304 & 258 & 229 & 187 & 115 & 524 & 490 \\
 \cline{2-10}
@@
 \hline
-\multirow{2}{*}{CE2}   & $d_{r}$  & 135 & 518 & 464 & 405 & 305 & 183 & 391 & 700 \\
+\multirow{2}{*}{CE2}   & $d_{r}$  & 135 & 499 & 451 & 405 & 305 & 183 & 391 & 898 \\
 \cline{2-10}
@@
 \end{table}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
-  \caption{Median of $coverage$ for s20.0--SFHo CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
-  \caption{} \label{fig:distances}
-\end{figure}

### 2020-10-10 conclusion.tex
complete results + add conclusion

\section{Conclusion}
The algorithm presented in this paper is a first attemp to infer the time evolution of a
combinaison of the mass of the PNS and its radius based on the universal relations found
in PNS asteroseismology. More precisely, we have considered in this paper the ratio
$r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$
oscillation mode in the GW data. We have especially investigated the performance of the algorithm
in the case of an optimally oriented source detected in a singe GW detector. For Advanced LIGO
or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
that this is true for a wide range of progenitor mass and that the quality of the inference
mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
as Einstein Telescope and Cosmic Explorer, the $\mbox{}^2g_2$ will be reconstructible for sources
at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
for all waveforms considered here thanks to its excellent sensitivity in the \unit{100-1000}[Hz] range.
Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
especially for the waveforms with the highest progenitor mass (25 $\Msol$ and 40 $\Msol$). Comparing
$d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
below \unit{200}[Hz] is important for massive mass progenitor signals.

This study does not include the realistic case of operating within a network of detectors. We defer this
for a forcoming publication. The sources of GWs we have considered here are optimally oriented. The
reported distance at which we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an
upper limit that may be lower by a factor 2--3 on average for a source located anywhere on the sky.

Finally, this method can be adapted to other PNS oscillation modes, changing few parameters such as
the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
reconstruct several modes in a same GW signal would allow to infer individually each of the PNS property.

### 2020-10-10 results.tex
complete results + add conclusion

@@
 Eintein Telescope sensitivity curve ET-B is obtained pushing second generation detector technology at its limit.
-ET-C and ET-D sensitivity curves correspond to a detector configuration where a low-power cryogenic low-frequency interferometer and a high-power room temperature high-frequency interferometer are sharing the same infrastructure~\cite{Hild_20211}.
+ET-C and ET-D sensitivity curves correspond to a detector configuration where a low-power cryogenic low-frequency interferometer and a high-power room temperature high-frequency interferometer are sharing the same infrastructure~\cite{Hild_2011}.
 Cosmic Explorer design sensitivity will be achieved in two stages.
@@
 \ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
-located up to several hundred of kpc. It is nevertheless important to note the rather large range obtained
+located up to several hundred of kpc. It is nevertheless important to note the rather wide range obtained
 for the different waveforms probing a large range of progenitor masses.
-
-\mab{add comment on the mass progenitor dependence}\\
-\mab{say that the main depend is the SNR as shown by ddet}
+We did not find any correlation between the mass of the progenitor and $d_r$. On the other hand, the quality
+of the ratio reconstruction depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by
+$d_{det}$.
 

### 2020-10-11 ccsn.tex
udpdate
add dmax
complete results + add conclusion
add abstrcat

@@
 \usepackage{subfigure}
-\newcommand{\note}[1]{{\color{red}#1}}
+\usepackage{multirow}
+%\newcommand{\note}[1]{{\color{red}#1}}
 \newcommand{\second}{\ensuremath{\mathrm{\,s}}}
@@
 \newcommand{\pcd}[1]{\textcolor{blue}{#1}}
+\newcommand{\mab}[1]{\textcolor{red}{#1}}
 
@@
 
-\title{Identification of protoneutron star g-modes in gravitational-wave data
-\pcd{[PCD: alt. title]Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}}
+\title{Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}
+
+\author{M.~A.~Bizouard}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+
+\author{P.~Cerd\'a-Dur\'an}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{N.~Christensen}
+\affiliation{Carleton College, Northfield, MN 55057, USA}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+
+\author{J.~A.~Font}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{P.~Maturana Russel}
+\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
+\affiliation{Department of Mathematical Sciences, Auckland University of Technology, Auckland, New Zealand}
+
+\author{R.~Meyer}
+\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
+
+\author{M.~Obergaulinger}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{A.~Torres-Forn\'e}
+\affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
+
+\mab{Alphabetical order ? Better to move forward the youngest}
 
 \begin{abstract}
+  Core collapse supernovae are very important phenomena in the Universe and their expected gravitational
+  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
+  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
+  makes the source parameter inference problem very challenging. In this paper, we have considered the
+  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
+  whose frequency time evolution is linked to the radius and the mass of the proton neutron star through
+  universal relations demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract
+  from the gravitational wave data the frequency evolution of the gravity mode. We have used a set of 1D
+  core collapse supernova simulations to build a model of evolution of the proto neutron star properties
+  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
+  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
+  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
+  11 and 40 solar mass and different equation of state. We have shown that for Advanced LIGO and Advanced
+  Virgo detectors at design sensitivity it will be possible to infer poroton neutron star properties for
+  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
+  to test distances of several hundreds of kiloparsec.
 
@@
 
-\input{discussion}\bigskip
+\input{conclusion}\bigskip
 

### 2020-10-11 introduction.tex
cosmetic

@@
 This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
-{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the { instantaneous frequency of the g-mode and} the mass and radius of the PNS {at each time of the evolution}.
+{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}.
 These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}.

### 2020-10-13 ccsn.tex
add NLC comments

@@
 \affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
 
@@
   proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
-  whose frequency time evolution is linked to the radius and the mass of the proton neutron star through
-  universal relations demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract
-  from the gravitational wave data the frequency evolution of the gravity mode. We have used a set of 1D
-  core collapse supernova simulations to build a model of evolution of the proto neutron star properties
+  whose frequency time evolution is linked to the system properties through universal relations as
+  demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract from the
+  gravitational wave data the frequency evolution of the gravity mode. A set of 1D core collapse
+  supernova simulations is used to build a model of evolution of the proto neutron star properties
   with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
@@
   performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
-  11 and 40 solar mass and different equation of state. We have shown that for Advanced LIGO and Advanced
-  Virgo detectors at design sensitivity it will be possible to infer poroton neutron star properties for
+  11 and 40 solar mass and different equations of state. We have shown that for Advanced LIGO and Advanced
+  Virgo detectors at design sensitivity it will be possible to infer proton neutron star properties for
   a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow

### 2020-10-13 conclusion.tex
add NLC comments

@@
 \section{Conclusion}
-The algorithm presented in this paper is a first attemp to infer the time evolution of a
+The algorithm presented in this paper is a first attempt to infer the time evolution of a
 combinaison of the mass of the PNS and its radius based on the universal relations found
@@
 or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
-that this is true for a wide range of progenitor mass and that the quality of the inference
+that this is true for a wide range of progenitor masses and that the quality of the inference
 mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
@@
 at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
-for all waveforms considered here thanks to its excellent sensitivity in the \unit{100-1000}[Hz] range.
+for all waveforms considered here thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
 Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
@@
 
+%The model to infer the ratio as function of the frequency is based on 1D {\sc AENUS-ALCAR} simulations.
+%We have observed that using the other numerical code {\sc CoCoNut} inputs lead to systematic bias of the
+%ratio reconstruction.
+
 This study does not include the realistic case of operating within a network of detectors. We defer this
-for a forcoming publication. The sources of GWs we have considered here are optimally oriented. The
+for a forthcoming publication. The sources of GWs we have considered here are optimally oriented. The
 reported distance at which we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an
@@
 the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
-reconstruct several modes in a same GW signal would allow to infer individually each of the PNS property.
+reconstruct several modes in the same GW signal would allow to infer individually each of the PNS property.

### 2020-10-13 introduction.tex
add NLC comments

@@
 For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful \mab{GW} signal {that could} be detected {within $50$~kpc and for some extreme models
-up to $5-30$ Mpc \cite{Gossan:2016,TargetedSNSearchO12}.
+up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
 However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs
-\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{add ref?}.
+\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}.
 Therefore, we focus this work only in neutrino-driven CCSNe.
@@
 %\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
-The core-bounce signal can be directly related with the rotational properties of the core} \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
+The core-bounce signal can be directly related with the rotational properties of the core} \citep{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
 \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
@@
 
-{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow {\mab oscillation} in the post bounce evolution with raising frequency
+{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} in the post bounce evolution with raising frequency
 from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
-This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \cite{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
 In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.
@@
  The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}.
-The oscillation modes {associated to} the surface of hot PNSs was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}.
-Additionally, the stratified structure of the PNS allows the presence of different types of g-modes related with the fluid core \cite{Reisenegger:1992}.
-Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}.
-{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \cite{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
+The oscillation modes {associated to} the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}.
+Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}.
+Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transition \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}.
+{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
 
- % {We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored
+ % {We base this work in the PNS mode analysis performed by \citep{Torres:2018, Torres:2019a}, which explored
 \mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
-using results from 2D CCSN numerical simulations as a background studied in \cite{Torres:2018, Torres:2019a}
-show a good match of the mode frequencies computed and the features observed in the GW spectrum of the same
+using results from 2D CCSN numerical simulations as a background studied in \citep{Torres:2018, Torres:2019a}
+shows a good match to the mode frequencies computed and the features observed in the GW spectrum of the same
 simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
 This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
-{\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}.
+Authors in {\citep{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}.
 These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}.
-{Similar relations have been found by \cite{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
+{Similar relations have been found by \citep{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
 
-{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius.
+{In this work, we present a method to infer from the GW data alone, the time evolution of some properties of the PNS, namely a combination of its mass and radius.
 For this purpose we have developed an algorithm to
 extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
-This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention).
+Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
 $M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2020-10-13 methods.tex
Update
update with final results
add NLC comments

@@
 In this section, we outline a strategy for estimating the time evolution of the
-ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
-(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
-oscillation mode in the gravitational wave detector data.
+ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
+oscillation mode in the GW detector data.
 An integral part of this strategy is the universal relations that relate the
@@
 
-\pcd{To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
+To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
 spherically symetric (1D) simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
@@
 deviation from a single universal relation is the numerical code used in the simulations. 
-To avoid any systematic effect, we only use the $18$ simulations performed with the ALCAR-AENUS
-code, which is the same code that was used in our test set. The consequences of this choice are discussed in the
-conclusions. Using this data, we} parametrize the ratio with a cubic polynomial
+To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
+code, which is the same code that was used in our test set.
+%The consequences of this choice are discussed in the conclusions.
+Using this data, we parametrize the discretized ratio $r_i$ with a cubic polynomial
 regression with heteroscedastic errors
@@
  \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.} \label{fig:LMVAR}
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.
+\mab{To be redone in python}} \label{fig:LMVAR}
 \end{figure}
 
-\pcd{We use this model to infer the properties of the simulations in the 
+{We use this model to infer the properties of the simulations in the 
 {\it test set} described in in Section \ref{sec:simulations}.}
-To describe the method we focus on the gravitational wave signal
+To describe the method we focus on the GW signal
 of {\texttt s20S}, originally
@@
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
-autoregressive estimates \pcd{[CITATION?]}of the local spectra for successive time intervals of 
+autoregressive estimates \pcd{[add ref?]}of the local spectra for successive time intervals of 
 \textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
 The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
-developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
+developed a time-frequency method to track the ridge $m(t) $in the spectrogram,
 taking into account that it is monotonically increasing as time goes. 
-\pcd{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
+{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
 increases as the object becomes more massive and compact}.
@@
  \centering
- \includegraphics[width=0.5\textwidth]{plots/spectrogram}
- \caption{Spectrogram of the gravitational wave signal {\texttt s20S} sampled at \unit[4096]{Hz}.
+ \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
+ \caption{Spectrogram of the GW signal {\texttt s20S} sampled at \unit[4096]{Hz}.
    The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
-   with each other.} \label{fig:spectrogram}
+   with each other. \mab{To be redone in python}} \label{fig:spectrogram}
 \end{figure}
@@
 the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use \pcd{our model given by Eq.~}
-\eqref{eq:model1} to obtain
+for values in between the $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
-An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
+An example is given in Figure \ref{fig:ratio} where the red points are the point estimates and
 the grey bands represent 95\% confidence bands. Ratio values
-computed using the mass and radius values obtained from the simulation code \pcd{(true values)} are shown in black.
+computed using the mass and radius values obtained from the simulation code \pcd{(true values)}
+are shown in black.
+In this example of a GW signal without noise the coverage of our 95\% confidence band is 100\%
+of the true values.
+In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
+signal is embedded in noise.
 
@@
 \end{figure}
-
-In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\% \pcd{of the true values}.
-In the next section we investigate the performance of \pcd{the} reconstruction of $r(t)$ when the gravitational wave
-signal is embedded in noise.

### 2020-10-13 results.tex
Add EOS comment
add NLC comments

@@
 
-To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
-gravitational wave detector data, we have added \pcd{the} GW signal \pcd{from {\texttt s20S}} to 
-100 Gaussian noise realisations whose power spectral density follows \pcd{the} advanced LIGO (aLIGO)
+To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in a single
+detector GW data, we have added the GW signal {\texttt s20S} to 
+100 Gaussian noise realisations whose power spectral density follows the Advanced LIGO (aLIGO)
 spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}.
@@
  \includegraphics[width=0.5\textwidth]{plots/spectrum}
- \caption{Amplitude spectral density of the GW detectors advanced LIGO (aLIGO) and Advanced Virgo (ADV) at design sensitivity and of the proposed third generation detectors Cosmic Explorer and Einstein Telescope.
+ \caption{Amplitude spectral density of the GW detectors Advanced LIGO (aLIGO) and Advanced Virgo (ADV) at design sensitivity and of the proposed third generation detectors Cosmic Explorer and Einstein Telescope.
 Eintein Telescope sensitivity curve ET-B is obtained pushing second generation detector technology at its limit.
@@
 Cosmic Explorer design sensitivity will be achieved in two stages.
-Stage 1 (CE1) is expected to use the technology developed for the “A+” upgrade to advanced LIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
+Stage 1 (CE1) is expected to use the technology developed for the “A+” upgrade to Advanced LIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
 } \label{fig:spectrum}
@@
 
-We have varied the distance to the source, covering a large
-range of distances for which a detection in second generation of gravitational wave detectors
-is feasible. The source is optimally oriented with
-respect to the gravitational wave detector. We are assuming a GW signal from a core collapse
+We have covered a large range of distances for which a detection in second generation of GW
+detectors is feasible. The source is optimally oriented with
+respect to the GW single detector. We are assuming a GW signal from a core collapse
 phenomena has been identified in the data and that the beginning of the GW signal is known within $O(10~ms)$.
@@
 
-For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
+For each of the noise realizations, we reconstruct the ratio time series {$r_i$}
 of length $N$ starting from the left side of the spectrogram and constraining the beginning of the
-track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
+track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compared to the ``true'' ratio
 {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
 
-Figure \ref{fig:s20results} is showing the distribution of the fraction of the ratio {$r_i^0$} values
+Figure \ref{fig:s20results} shows the distribution of the fraction of the ratio {$r_i^0$} values
 that fall within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
@@
 remains smaller than 20\%.
-At small distance, $\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
+At small distances, $\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
 It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
 precision at distance up to $\sim$ 9 kpc for this particular waveform, with $coverage$ value
-larger than 80\%. There are few noise realisations for a source located at $<$ \unit[9]{kpc} for which
+larger than 80\%. There are few noise realizations for a source located at $<$ \unit[9]{kpc} for which
 $\Delta$ takes large values, indicating that the method start failing to reconstruct with accuracy the ratio.
@@
 
- \caption{Boxplots of $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
-100 noise realisations is considered for each distance.}
+ \caption{Boxplots of the $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
+100 noise realizations are considered for each distance.}
   \label{fig:s20results}
@@
 
-Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220} and to a lesser extent
-{\tt s20.0--SFHo}, the ratio is well reconstructed for all waveforms up to $\sim$ 15kpc. In an
+Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from {\texttt s11} and to a lesser extent
+{\texttt s20S}, the ratio is well reconstructed for all waveforms up to $\sim$ 15kpc. In an
 effort to better determine the maximal distance of the source at which we can reconstruct the ratio
@@
 simulation but consider the source is optimally oriented. 
-Table \ref{tab:results} reports also $d_r$ for the advanced Virgo detector at design sensitivity.
+Table \ref{tab:results} reports also $d_r$ for the Advanced Virgo detector at design sensitivity.
 Results are very similar to aLIGO, despite the detector sensitivity differences.
@@
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
- \caption{Median of $Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
-The ``no signal'' line and band show the median and first and third quartile of $Delta$ in absence of any signal.} \label{fig:aLIGO_prec_allwvf}
+ \caption{Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
+The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.} \label{fig:aLIGO_prec_allwvf}
 \end{figure}
@@
 The same analysis has been performed using expected sensitivity curves for the third generation of
-gravitational wave detectors. In Europe the Einstein Telescope project proposes to host in a 10-km
+GW detectors. In Europe the Einstein Telescope project proposes to host in a 10-km
 equilateral triangle configuration 3 low-power low-frequency cryogenic interferometers as well as 3
@@
 
-Figure \ref{fig:s20--SFHo_all3G} shows $\Delta$ as function of the source distance for {\tt s20.0--SFHo}
-waveform for the 5 3G detectors configurations. Overall, the ratio is well reconstructed up to distances
-in the range \unit{100--200}[kpc] which represents an improved of at least a factor 10 with respect to
-advanced LIGO and advanced Virgo detectors. We can also note that the Einstein Telescope results lay in
-between the 2 Cosmic Explorer results. This is confirmed for all other waveforms, expect {\tt s25.0-LS220}
+Figure \ref{fig:s20--SFHo_all3G} shows $\Delta$ as function of the source distance for {\texttt s20S}
+waveform for the five 3G detector configurations. Overall, the ratio is well reconstructed up to distances
+in the range \unit[100--200]{kpc} which represents an improvement of a factor 10 with respect to
+Advanced LIGO and Advanced Virgo detectors. We can also note that the Einstein Telescope results lay in
+between the 2 Cosmic Explorer results. This is confirmed for all other waveforms, expect {\texttt s25}
 for which the maximal distance reach in CE2 is significantly lower than CE1. This is partly due to the
 small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$
-rather incertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
+rather uncertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
 \ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
@@
 for the different waveforms probing a large range of progenitor masses.
-We did not find any correlation between the mass of the progenitor and $d_r$. On the other hand, the quality
+We did not find any correlation between the mass of the progenitor and $d_r$, nor the equation of state.
+On the other hand, the quality
 of the ratio reconstruction depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by
@@
   \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
-  \caption{Median of $\Delta$ for s20.0--SFHo CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
+  \caption{Median of $\Delta$ for {\texttt s20S} CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
-  \caption{Maximal distance $d_{r}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
+  \caption{Maximal distance $d_{r}$ in kpc at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
     with good accuracy for a source optimally oriented with respect to the GW detectors for the 7 CCSN waveforms considered in this study.} \label{fig:distances}
@@
     optimally oriented with a matched filter signal-to-noise ratio of 13 in the  different
-    GW detectors.
+    GW detectors. All distances are expressed in kpc.
     %%

### 2020-10-13 simulations.tex
add NLC comments

@@
 
-{Unless other methods used GW astronomy, the} algorithm proposed in {this work} does not require accurate
-waveforms {in order to infer the properties of the PNS.} {Instead, it relies on the evolution of the frequency of
-oscillation of some particular modes, as seen in the GW spectrum. The frequency of these modes depends, in a universal way, 
-on the surface gravity of the PNS ($r=M_{\rm PNS}/R_{\rm PNS}^2$), in the sense that if at a given time we observe GW emission
-at a certain frequency $f$ we can determine univocally the value of the surface gravity, within a certain error, regardless of the details
-of the numerical simulation. In this work we use two sets of simulations: i) The {\it model set}, composed by 1D simulations, which is used 
-to build the universal relation (model), $r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by 2D simulations, 
-for which we know both the GW signal and the evolution of the ratio, $r (t)$, and that is used to test performance of the algorithm.}
+{Unlike other methods used GW astronomy, the} algorithm proposed in {this work} does not require accurate
+waveforms {in order to infer the properties of the PNS.} {Instead, it relies on the evolution of the
+frequency of oscillation of some particular modes, as seen in the GW spectrum.
+The frequency of these modes depends, in a universal way, on the surface gravity of the PNS
+($r=M_{\rm PNS}/R_{\rm PNS}^2$), in the sense that if at a given time we observe GW emission at a certain
+frequency $f$ we can determine unequivocally the value of the surface gravity, within a certain error,
+regardless of the details of the numerical simulation. In this work we use two sets of simulations: i)
+The {\it model set}, composed by 1D simulations, which is used to build the universal relation (model),
+$r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by
+2D simulations, 
+for which we know both the GW signal and the evolution of the ratio, $r (t)$, and that is used to test
+performance of the algorithm.}
 
@@
  simulations, so is easier to accumulate the statistics necessary to build a good model for $r(f)$.}
- \pcd{For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
+ {For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
  described in \cite{Torres:2018,Torres:2019a,Torres:2019b}. }
@@
 to $240 \, \mathrm{MeV}$.
-\pcd{Unlike the model set, the simulations in the test set are not 1D because we need to 
-extract the gravitational wave signal, which is a multi-dimensional effect. For each simulation
-we extract the gravitational wave signal, $h_+(t)$,  by means of the quadrupole formula and compute the 
+{Unlike the model set, the simulations in the test set are not 1D because we need to 
+extract the GW signal, which is a multi-dimensional effect. For each simulation
+the GW signal, $h_+(t)$, is extracted by means of the quadrupole formula and we compute the 
 time evolution of the surface gravity, $r(t)$.}
@@
 masses of the PNSs are in the range of
-$M_{\mathrm{PNS}} = 1.47 - 2.33 \, \Msol$, i.e., insufficient for
+$M_{\mathrm{PNS}} = 1.47 - 2.33 \, \Msol$, i.e., likely insufficient for
 producing a black hole.

### 2020-10-15 conclusion.tex
Update

@@
 \section{Conclusion}
+\label{sec:conclusion}
 The algorithm presented in this paper is a first attempt to infer the time evolution of a
@@
 $d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
-below \unit{200}[Hz] is important for massive mass progenitor signals.
+below \unit[200]{Hz} is important for massive mass progenitor signals.
 
@@
 
-This study does not include the realistic case of operating within a network of detectors. We defer this
-for a forthcoming publication. The sources of GWs we have considered here are optimally oriented. The
-reported distance at which we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an
-upper limit that may be lower by a factor 2--3 on average for a source located anywhere on the sky.
+This study does not include the realistic case of operating within a network of detectors.
+The sources of GWs we have considered here are optimally oriented. The reported distance at which
+we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an upper limit that may be lower
+by a factor 2--3 on average for a source located anywhere on the sky. We defer a more realistic
+simulation implementation for a forthcoming publication.
+
+Furthermore, the results presented here are based on a model of the frequency evolution of the
+ratio that is making use of AENUS-ALCAR code simulations. Using simulations produced
+with the CoCoNut code, the model fit is a bit different (see Figure \ref{fig:LMVAR}). This systematic
+difference of the two simulation codes is discussed in \cite{Torres:2019b}. If we consider a model based
+on CoCoNut simulations and test the method on test simulations obtained with the AENUS-ALACR code,
+we observe systematic effects in the ratio reconstruction leading to systematically worsen performance.
+As the goal of this paper is to characterize the methods and because both simulation codes have pros
+and cons we have reported results obtained with a model set and test set waveforms obtained with the same
+code.
 

### 2020-10-22 conclusion.tex
Remove Coconut inputs and discussion

@@
 
-Furthermore, the results presented here are based on a model of the frequency evolution of the
-ratio that is making use of AENUS-ALCAR code simulations. Using simulations produced
-with the CoCoNut code, the model fit is a bit different (see Figure \ref{fig:LMVAR}). This systematic
-difference of the two simulation codes is discussed in \cite{Torres:2019b}. If we consider a model based
-on CoCoNut simulations and test the method on test simulations obtained with the AENUS-ALACR code,
-we observe systematic effects in the ratio reconstruction leading to systematically worsen performance.
-As the goal of this paper is to characterize the methods and because both simulation codes have pros
-and cons we have reported results obtained with a model set and test set waveforms obtained with the same
-code.
+%Furthermore, the results presented here are based on a model of the frequency evolution of the
+%ratio that is making use of AENUS-ALCAR code simulations. Using simulations produced
+%with the CoCoNut code, the model fit is a bit different (see Figure \ref{fig:LMVAR}). This systematic
+%difference of the two simulation codes is discussed in \cite{Torres:2019b}. If we consider a model based
+%on CoCoNut simulations and test the method on test simulations obtained with the AENUS-ALACR code,
+%we observe systematic effects in the ratio reconstruction leading to systematically worsen performance.
+%As the goal of this paper is to characterize the methods and because both simulation codes have pros
+%and cons we have reported results obtained with a model set and test set waveforms obtained with the same
+%code.
 

### 2020-10-22 methods.tex
Update
Remove Coconut inputs and discussion

@@
 deviation from a single universal relation is the numerical code used in the simulations. 
-To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
-code, which is the same code that was used in our test set.
+%To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
+%code, which is the same code that was used in our test set.
+%\mab{We defer a discussion of the possible consequences of this choice in Section \ref{sec:conclusion}.}
 %The consequences of this choice are discussed in the conclusions.
@@
  \includegraphics[width=0.5\textwidth]{plots/model}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
+% \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
 The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.
 \mab{To be redone in python}} \label{fig:LMVAR}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations using {\sc AENUS-ALCAR} code.
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
+} \label{fig:LMVAR}
+
 \end{figure}
@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
+ \includegraphics[width=0.5\textwidth]{plots/spectrogram}
  \caption{Spectrogram of the GW signal {\texttt s20S} sampled at \unit[4096]{Hz}.
    The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
-   with each other. \mab{To be redone in python}} \label{fig:spectrogram}
+   with each other.} \label{fig:spectrogram}
 \end{figure}
@@
  \centering
- \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points) \mab{To be redone in python}.} \label{fig:ratio}
+ \includegraphics[width=0.5\textwidth]{plots/ratio}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
 \end{figure}

### 2020-10-22 simulations.tex
Remove Coconut inputs and discussion

@@
 
-{We have used two different numerical codes in our numerical simulations.} 
+%{We have used two different numerical codes in our numerical simulations.} 
 %CoCoNuT
@@
 %(one- and two-dimensional models). 
-CoCoNuT
-\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
-relativistic hydrodynamics coupled to the Fast Multigroup Transport
-scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
-description of the emission and transport of neutrinos. AENUS-ALCAR
+%CoCoNuT
+%\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
+%relativistic hydrodynamics coupled to the Fast Multigroup Transport
+%scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
+%description of the emission and transport of neutrinos.
+\mab{Both the {\it model set} and {\it test set} simulations have been generated using the numerical
+  relativity code AENUS-ALCAR
 \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
-combines special relativistic (magneto-)hydrodynamics, a modified
+which }combines special relativistic (magneto-)hydrodynamics, a modified
 Newtonian gravitational potential approximating the effects of general
@@
 
-{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the AENUS-ALCAR code
+{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the {\sc AENUS-ALCAR} code
 (see Table~\ref{Tab:2dSimList} for a list of models).

### 2020-10-23 ccsn.tex
Update
G-mode reconstruction update
Moving average added in the text
Work on the abstract.

--- 2020-09-11 ccsn.tex
@@
 \newcommand{\pcd}[1]{\textcolor{blue}{#1}}
+\newcommand{\mab}[1]{\textcolor{red}{#1}}
+\newcommand{\tf}[1]{\textcolor{green}{#1}}
 
@@
 
-\title{Identification of protoneutron star g-modes in gravitational-wave data
-\pcd{[PCD: alt. title]Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}}
+\title{Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae.}
+
+\author{M.~A.~Bizouard}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+
+\author{P.~Maturana-Russel}
+\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
+\affiliation{Department of Mathematical Sciences, Auckland University of Technology, Auckland, New Zealand}
+
+\author{A.~Torres-Forn\'e}
+\affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{M.~Obergaulinger}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{P.~Cerd\'a-Dur\'an}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+
+\author{N.~Christensen}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+\affiliation{Carleton College, Northfield, MN 55057, USA}
+
+\author{J.~A.~Font}
+\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
+\affiliation{Observatori Astron\`omic, Universitat de Val\`encia, E-46980, Paterna, Val\`encia, Spain}
+
+\author{R.~Meyer}
+\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
 
 \begin{abstract}
+The eventual detection of gravitational waves from core-collapse supernovae (CCSN) may help improve our current understanding of the explosion mechanism of massive stars.
+The stochastic nature of the late post-bounce gravitational wave signal due to matter effects and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
+In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with convective oscillations of proto-neutron stars (PNS).
+Numerical simulations of CCNS have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
+We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.
+The model is used to infer the time evolution of combinations of the mass and the radius of the PNS.
+The  performance of the method is estimated employing  simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
+Our results indicate that it is possible to infer PNS properties for a galactic source using Advanced LIGO and Advanced Virgo at design sensitivities.
+Third generation detectors such as Einstein Telescope and Cosmic Explorer will allow to test distances of ${\cal O}(100\, {\rm kpc})$.
+\tf{Caveats?
+Gaussian noise?}
+% 
+%  Core-collapse supernovae are very important phenomena in the Universe and their expected gravitational
+%  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
+%  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
+%  makes the source parameter inference problem very challenging. In this paper, we have considered the
+%  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
+%  whose frequency time evolution is linked to the system properties through universal relations as
+%  demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract from the
+%  gravitational wave data the frequency evolution of the gravity mode. A set of 1D core collapse
+%  supernova simulations is used to build a model of evolution of the proto neutron star properties
+%  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
+%  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
+%  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
+%  11 and 40 solar mass and different equations of state. We have shown that for Advanced LIGO and Advanced
+%  Virgo detectors at design sensitivity it will be possible to infer proton neutron star properties for
+%  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
+%  to test distances of several hundreds of kiloparsec.
 

### 2020-10-28 introduction.tex
Some work on the intro.

@@
 
-The life of massive stars {(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion.
-{Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GWs) and are one of the main candidates for the next great discovery by current ground-based observatories.
+The life of sufficiently massive stars, i.e.~those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$, ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion.
+Nearby core-collapse supernova (CCSN) explosions are expected to be sources of gravitational waves (GWs) and they could be
+%are one of the main candidates for 
+the next great discovery of current ground-based observatories.
 However, these are relative rare events.
-A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
-\mab{The emitted GWs} could be detected with advanced ground-based GW detectors \mab{(
-Advanced LIGO\citep{TheLIGOScientific:2014jea},
-Advanced Virgo\citep{TheVirgo:2014hva} and
-KAGRA\citep{Aso:2013eba})} within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
-For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful \mab{GW} signal {that could} be detected {within $50$~kpc and for some extreme models
-up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
-However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs
-\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}.
-Therefore, we focus this work only in neutrino-driven CCSNe.
-Despite the low rates, CCSN are of great scientific interest because they produce} a complex GW signals which could provide significant clues about the physical processes that occur in the moments after the collapse.
+A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors.
+The emitted GWs could be detected with the advanced ground-based GW detector network (Advanced LIGO~\citep{TheLIGOScientific:2014jea}, Advanced Virgo~\citep{TheVirgo:2014hva} and
+KAGRA~\citep{Aso:2013eba}) within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
+For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within $50$~kpc and, for some extreme models, up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
+However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs
+\citep{Chapman:2007}) making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}.
+For the results discussed in this work we only consider neutrino-driven CCSN.
+\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
 

### 2020-10-29 introduction.tex
A little more work on the intro ...

@@
 
-In the last decade, significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
-% {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
-%\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood.
-The core-bounce signal can be directly related with the rotational properties of the core} \citep{Dimmelmeier:2007, abdikamalov:14, Richers:2017}.
-\sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
-In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
-This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs.
-The frequency and time evolution of these modes carry information about the
-properties of the GW emitter and could allow to perform PNS asteroseismology. }
+In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+This} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs.
+The frequency and time evolution of these modes carry information about the properties of the GW emitter and could allow to perform PNS asteroseismology.
 
@@
 
-{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} in the post bounce evolution with raising frequency
+All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
+%
+%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
+%
+during the post-bounce evolution of the system, with raising frequency
 from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
 This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
-In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.
+In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
+%promising {source} for 
+ground-based GW interferometers.
 
- The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}.
-The oscillation modes {associated to} the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}.
+ The {properties of} g-modes in hot {PNSs} have been studied since the 1990s
+ %end of last century 
+ {by means of linear perturbation analysis of background PNS models}.
+The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}.
 Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}.
-Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transition \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}.
-{Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
+Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transitions \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}.
+{Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
 

### 2020-10-30 methods.tex
update
Description Figure 3

@@
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
-An example is given in Figure \ref{fig:ratio} where the red points are the point estimates and
-the grey bands represent 95\% confidence bands. Ratio values
+An example is given in Figure \ref{fig:ratio} where the red triangles \sout{points} are the point estimates and
+the grey bands represent 95\% confidence bands.
+{\color{red} The size of the red triangles are proportional to the magnitude of the $\mbox{}^2 g_2$-mode estimates.
+Note that high $\mbox{}^2 g_2$-mode values have associated more uncertainty (bigger intervals) because our model allows heterogeneous variance.} Ratio values
 computed using the mass and radius values obtained from the simulation code \pcd{(true values)}
@@
  \includegraphics[width=0.5\textwidth]{plots/ratio}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time {\color{red}estimated} \sout{extracted} from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red {\color{red}triangles} \sout{points} and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).
+Estimated ratios are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode estimates.
+} \label{fig:ratio}
 \end{figure}

### 2020-10-30 introduction.tex
Finished my revision of the introduction

@@
 
- % {We base this work in the PNS mode analysis performed by \citep{Torres:2018, Torres:2019a}, which explored
-\mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
-using results from 2D CCSN numerical simulations as a background studied in \citep{Torres:2018, Torres:2019a}
-shows a good match to the mode frequencies computed and the features observed in the GW spectrum of the same
-simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
-This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.
-Authors in {\citep{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}.
-These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}.
-{Similar relations have been found by \citep{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}
+Using results from 2D CCSN numerical simulations as a background with respect to which to solve an eigenvalue problem~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
+shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
+\tf{This last sentence might be improved.}
+This reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. 
+Further work was presented in {\citep{Torres:2019b} who found that it is possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the numerical evolutions}.
+These relations are universal as they do not depend {on the equation of state (EOS) or on the mass of the progenitor {and they only depend weakly on} the numerical code used {(see discussion in Section~\ref{sec:simulations})}.
+{Similar universal relations have been discussed by \citep{Sotani:2020a,Sotani:2020b} who also found that they do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.
 
-{In this work, we present a method to infer from the GW data alone, the time evolution of some properties of the PNS, namely a combination of its mass and radius.
-For this purpose we have developed an algorithm to
+{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
 extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
 This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being
-$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
+Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.} Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
 
-This paper is organised as follows.
+This paper is organized as follows.
 Section II describes the details of the CCSN simulations used in the paper.
-Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode.
-Section IV shows the performance of the data analysis method with simulated GW detectors data.
-Finally, we discuss the results in section V.
+The algorithm that we employ to extract the time evolution of the PNS surface gravity is discussed in Section III.
+Section IV shows the performance of our inference method with simulated GW detectors data and presents our main results.
+Our findings are summarized in section V.
+Appendix A discusses specific details related to the reconstruction of the g-mode.

### 2020-10-30 introduction.tex
Minor changes.

@@
 This paper is organized as follows.
-Section II describes the details of the CCSN simulations used in the paper.
+Section II provides details of the CCSN simulations used in our work.
 The algorithm that we employ to extract the time evolution of the PNS surface gravity is discussed in Section III.
-Section IV shows the performance of our inference method with simulated GW detectors data and presents our main results.
-Our findings are summarized in section V.
+Section IV shows the performance of our inference method and presents our main results.
+Finally, our findings are summarized in Section V.
 Appendix A discusses specific details related to the reconstruction of the g-mode.

### 2020-11-06 simulations.tex
Work on the simulations section.

@@
 
-% \textbf{Martin's simulations and code description}\\
-% \textbf{1D simlation data to fit the ratio vs frequency model - AA and CoConut outputs}
+Unlike other methods used in GW astronomy the algorithm proposed in {this work} does not require accurate
+waveforms {in order to infer the properties of the PNS.} Instead, it relies on the evolution of the oscillation
+frequency of some particular modes visible in the GW spectrum.
+The frequency of these modes depends in a universal way on the surface gravity of the PNS, $r\equiv M_{\rm PNS}/R_{\rm PNS}^2$ \citep{Torres:2019b}.
+Therefore, if at a given time GW emission is observed at a certain
+frequency $f$ then the value of the surface gravity can be determined unequivocally, within a certain error,
+regardless of the details of the numerical simulation.
 
-{Unlike other methods used GW astronomy, the} algorithm proposed in {this work} does not require accurate
-waveforms {in order to infer the properties of the PNS.} {Instead, it relies on the evolution of the
-frequency of oscillation of some particular modes, as seen in the GW spectrum.
-The frequency of these modes depends, in a universal way, on the surface gravity of the PNS
-($r=M_{\rm PNS}/R_{\rm PNS}^2$), in the sense that if at a given time we observe GW emission at a certain
-frequency $f$ we can determine unequivocally the value of the surface gravity, within a certain error,
-regardless of the details of the numerical simulation. In this work we use two sets of simulations: i)
-The {\it model set}, composed by 1D simulations, which is used to build the universal relation (model),
-$r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by
+In this work we use two sets of simulations: i) the {\it model set}, composed by 1D simulations, which is used to build the 
+universal relation (model), $r(f)$, linking the ratio $r$ with the observed frequency $f$, and ii) the {\it test set}, composed by
 2D simulations, 
 for which we know both the GW signal and the evolution of the ratio, $r (t)$, and that is used to test
-performance of the algorithm.}
+performance of the algorithm.
 
@@
 %description of the emission and transport of neutrinos.
-\mab{Both the {\it model set} and {\it test set} simulations have been generated using the numerical
-  relativity code AENUS-ALCAR
+Both the {\it model set} and {\it test set} simulations have been generated using the AENUS-ALCAR code
 \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
-which }combines special relativistic (magneto-)hydrodynamics, a modified
+which combines special relativistic (magneto-)hydrodynamics, a modified
 Newtonian gravitational potential approximating the effects of general
@@
 \citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
-We included the relevant reactions between matter and neutrinos of all
+All siimulations include the relevant reactions between matter and neutrinos of all
 flavours, i.e., emission and absorption by nucleons and nuclei,
@@
 
-{For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
+For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
 including progenitors with zero-age main sequence (ZAMS) masses in the range 
-$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$. The set contains simulations using the
- two numerical codes and six different equations of state. Details can be found in
- \citep{Torres:2019a}. The reason to use one dimensional simulations for the model set
- is that the computational cost of those is significantly smaller than the cost of multidimensional
- simulations, so is easier to accumulate the statistics necessary to build a good model for $r(f)$.}
+$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$. The set contains simulations using  
+six different EOS. Details can be found in
+ \citep{Torres:2019a}. The reason to use 1D simulations for the model set
+ is that their computational cost is significantly smaller than that of multidimensional
+ simulations which allows us to accumulate the statistics necessary to build a good model for $r(f)$.
  {For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
@@
 
-{For the {\it test set}, we use $8$ axisymmetric (2D) simulations using the {\sc AENUS-ALCAR} code
+{For the {\it test set}, we use $8$ axisymmetric (2D) simulations performed with the {\sc AENUS-ALCAR} code
 (see Table~\ref{Tab:2dSimList} for a list of models).
-$7$ of these simulations use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
+All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
  evolved through the hydrostatic phases by
 \cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
-We performed one simulation of each stellar model using the equation
-of state of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
+We performed one simulation of each stellar model using the EOS of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
 an incompressibility of $K = 220 \, \mathrm{MeV}$ (LS220) and added
@@
 for the {progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$. To this set of
-simulations, we add the waveform of a two-dimensional model used in
+simulations we add the waveform of a 2D model used in
 \cite{Torres:2019a}, denoted \texttt{s20S}. It corresponds to a star
 with the same initial mass, $M_{\mathrm{ZAMS}} = 20 \, \Msol$, as for
-one of the other 7 axisymmetric simulations, but was taken from a
+one of the other seven axisymmetric simulations, but was taken from a
 newer set of stellar-evolution models
@@
 {For all the simulations,} we mapped the pre-collapse state of the stars to a spherical
-coordinate system with $n_r = 400$ zones in radial direction
+coordinate system with $n_r = 400$ zones in the radial direction
 distributed logarithmically with a minimum grid width of
@@
 $r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and
-$n_{\theta} = 128$ equidistant cells in angular direction. For the
+$n_{\theta} = 128$ equidistant cells in the angular (polar) direction. For the
 neutrino energies, we used a logarithmic grid with $n_e = 10$ bins up
@@
 
-All spherical and most axisymmetric models fail to achieve shock
+All spherical and most axisymmetric models we evolved fail to achieve shock
 revival during the time of our simulations. Only the two stars with

### 2020-11-06 methods.tex
Work on the methods section.

--- 2020-10-22 methods.tex
@@
 
-In this section, we outline a strategy for estimating the time evolution of the
-ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
-oscillation mode in the GW detector data.
-An integral part of this strategy is the universal relations that relate the
-characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
-and the radius of the PNS, the shock radius and the total mass inside the shock as
-demonstrated in \cite{Torres:2019b}.
+\begin{figure}[t]
+ \centering
+ \includegraphics[width=0.45\textwidth]{plots/model}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 25 1D simulations of the model set.
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
+\tf{18 or 25?} \tf{We could remove the legend as only one code is used.} }
+ \label{fig:LMVAR}
+\end{figure}
 
+We next outline our strategy for estimating the time evolution of $r(t)$
+%ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) 
+from the observation of the $\mbox{}^2g_2$ oscillation mode in the GW detector data.
+%An integral part of this strategy is the universal relations that relate the
+%characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
+%and the radius of the PNS, the shock radius and the total mass inside the shock as
+%demonstrated in \cite{Torres:2019b}.
 To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
-spherically symetric (1D) simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
-shows the data for the $25$ numerical simulations. As identified by \cite{Torres:2019b}, the only systematic
-deviation from a single universal relation is the numerical code used in the simulations. 
+1D simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
+shows the data for the $25$ numerical simulations. 
+%As identified by \cite{Torres:2019b}, the only systematic deviation from a single universal relation is the numerical code used in the simulations. 
 %To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
@@
 %The consequences of this choice are discussed in the conclusions.
-Using this data, we parametrize the discretized ratio $r_i$ with a cubic polynomial
+Using these data we parametrize the discretized ratio $r_i$ with a cubic polynomial
 regression with heteroscedastic errors
-%Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn
-
 \begin{equation}
 \label{eq:model1}
-r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
+r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i\,,
 \end{equation}
@@
 \begin{equation}
-\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
+\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i\,,
 \end{equation}
@@
 
-\begin{figure}
- \centering
- \includegraphics[width=0.5\textwidth]{plots/model}
-% \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.
-\mab{To be redone in python}} \label{fig:LMVAR}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations using {\sc AENUS-ALCAR} code.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
-} \label{fig:LMVAR}
-
-\end{figure}
-
 {We use this model to infer the properties of the simulations in the 
-{\it test set} described in in Section \ref{sec:simulations}.}
+{\it test set} discussed in Section \ref{sec:simulations}.}
 To describe the method we focus on the GW signal
-of {\texttt s20S}, originally
-sampled at 16384 Hz but resampled at 4096 Hz.
+of model {\texttt s20S}, originally
+sampled at 16384 Hz but downsampled at 4096 Hz.
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
-autoregressive estimates \pcd{[add ref?]}of the local spectra for successive time intervals of 
-\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
+autoregressive estimates \pcd{[add ref?]} of the local spectra for successive time intervals of 
+length 200 with a 90\% overlap.
 The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
-developed a time-frequency method to track the ridge $m(t) $in the spectrogram,
-taking into account that it is monotonically increasing as time goes. 
-{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
-increases as the object becomes more massive and compact}.
+developed a time-frequency method to track the ridge $m(t) $ in the spectrogram,
+taking into account that it is monotonically increasing with time. 
+This is a property of the $\mbox{}^2 g_2$-mode whose frequency   
+increases as the PNS becomes more massive and compact.
 Starting from either the left- or right-most column of the time-frequency matrix
 we identify and trace the sequence of amplitude peaks within a certain frequency
-band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
-details on the reconstruction of the $g$ mode ridge.
+band given the monotonicity constraint. Specific details about the reconstruction of the $\mbox{}^2 g_2$-mode ridge 
+are provided in Appendix \ref{app:gmode}.
 
@@
  \caption{Spectrogram of the GW signal {\texttt s20S} sampled at \unit[4096]{Hz}.
-   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
-   with each other.} \label{fig:spectrogram}
+   The spectrogram is obtained using a data streach of 200 samples overlapping at 90\%
+   with each other. The open circles track the ridge $m(t) $ of the $\mbox{}^2 g_2$-mode. } \label{fig:spectrogram}
 \end{figure}
@@
 We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
-the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
-for values in between the $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
+the midpoint $t_i$ of each local time interval of the spectrogram and interpolate $f(t)$
+for values in between $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
 estimates of the time evolution of the ratio together with 95\% confidence intervals.
-An example is given in Figure \ref{fig:ratio} where the red points are the point estimates and
-the grey bands represent 95\% confidence bands. Ratio values
-computed using the mass and radius values obtained from the simulation code \pcd{(true values)}
-are shown in black.
-In this example of a GW signal without noise the coverage of our 95\% confidence band is 100\%
-of the true values.
-In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
+An example is given in Figure \ref{fig:ratio} where the red triangles are the point estimates and
+the grey bands represent 95\% confidence bands.
+The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode estimates.
+Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance.
+Ratio values
+computed using the mass and radius values obtained from the simulation code (i.e.~the true values)
+are shown in black. In this example of a noise-free GW signal the coverage of our 95\% confidence band is 100\%
+of the true values. In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
 signal is embedded in noise.
+\tf{We also note that despite we have only explicitely shown results for the GW signal of the {\texttt s20S} model the same conclusions hold for any of the other waveforms of our test set.}
 
@@
  \includegraphics[width=0.5\textwidth]{plots/ratio}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
+ \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles).
+The ratios are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode estimates.}
+ \label{fig:ratio}
 \end{figure}

### 2020-11-07 introduction.tex
Further work on the paper.

@@
 A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors.
-The emitted GWs could be detected with the advanced ground-based GW detector network (Advanced LIGO~\citep{TheLIGOScientific:2014jea}, Advanced Virgo~\citep{TheVirgo:2014hva} and
+The emitted GWs could be detected with the advanced ground-based GW detector network (Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
 KAGRA~\citep{Aso:2013eba}) within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.

### 2020-11-07 results.tex
Further work on the paper.

@@
 
-To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in a single
-detector GW data, we have added the GW signal {\texttt s20S} to 
-100 Gaussian noise realisations whose power spectral density follows the Advanced LIGO (aLIGO)
-spectrum~\cite{aLIGOsens:2018} shown on Figure~\ref{fig:spectrum}.
+To estimate how accurately we can infer the time evolution of $r(t)$ in the GW data of a single
+detector we inject the GW signal for model {\texttt s20S} into 
+100 Gaussian noise realisations whose power spectral density (PSD) follows the aLIGO
+spectrum~\cite{aLIGOsens:2018}. The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
 
-\begin{figure}
+\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{plots/spectrum}
- \caption{Amplitude spectral density of the GW detectors Advanced LIGO (aLIGO) and Advanced Virgo (ADV) at design sensitivity and of the proposed third generation detectors Cosmic Explorer and Einstein Telescope.
-Eintein Telescope sensitivity curve ET-B is obtained pushing second generation detector technology at its limit.
+ \caption{Amplitude spectral density of the GW detectors aLIGO and AdV at design sensitivity and that of the proposed third-generation detectors Cosmic Explorer and Einstein Telescope.
+Einstein Telescope sensitivity curve ET-B is obtained pushing second-generation detector technology at its limit.
 ET-C and ET-D sensitivity curves correspond to a detector configuration where a low-power cryogenic low-frequency interferometer and a high-power room temperature high-frequency interferometer are sharing the same infrastructure~\cite{Hild_2011}.
 Cosmic Explorer design sensitivity will be achieved in two stages.
-Stage 1 (CE1) is expected to use the technology developed for the “A+” upgrade to Advanced LIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
+Stage 1 (CE1) is expected to use the technology developed for the ``A+'' upgrade to aLIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
 } \label{fig:spectrum}
@@
 
-We have covered a large range of distances for which a detection in second generation of GW
-detectors is feasible. The source is optimally oriented with
-respect to the GW single detector. We are assuming a GW signal from a core collapse
-phenomena has been identified in the data and that the beginning of the GW signal is known within $O(10~ms)$.
-The data (signal embedded in noise) are whitened using the function {\it prewhiten} of the R-package TSA.
-An auto-regressive model with maximal \textcolor{red}{100} coefficients has been used.
+We cover a large range of distances for which a CCSN detection in second-generation GW detectors is feasible.
+We assume that the source is optimally oriented with respect to our single detector.
+Moreover, we also assume that a CCSN GW signal has been identified in the data and that the beginning of the signal is known within {$\mathcal O$}(10 ms).
+The data (signal embedded in noise) are whitened using the function {\tt prewhiten} of the R-package {\tt TSA}.
+An auto-regressive model with a maximum of maximal 100 coefficients is used.
 
-For each of the noise realizations, we reconstruct the ratio time series {$r_i$}
-of length $N$ starting from the left side of the spectrogram and constraining the beginning of the
-track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compared to the ``true'' ratio
-{$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
-
-Figure \ref{fig:s20results} shows the distribution of the fraction of the ratio {$r_i^0$} values
-that fall within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
-when the source is located within few kpc and then decreases with the distance.
-
-To better quantify how well we reconstruct the ratio, we have also considered $\Delta$ the mean
-over the track of the relative error of $r_i$.
-
-\begin{equation}
-\Delta=\frac{1}{N}\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
-\end{equation}
-
-$\Delta$ values of each of the 100 noise realisation are shown as well as function of the distance
-on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[9]{kpc} the relative error
-remains smaller than 20\%.
-At small distances, $\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
-It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
-precision at distance up to $\sim$ 9 kpc for this particular waveform, with $coverage$ value
-larger than 80\%. There are few noise realizations for a source located at $<$ \unit[9]{kpc} for which
-$\Delta$ takes large values, indicating that the method start failing to reconstruct with accuracy the ratio.
-
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
  \caption{Boxplots of the $coverage$ (upper panel) and $\Delta$ (lower panel) for {\texttt s20S} signal embedded in aLIGO noise at different distances from the Earth.
-100 noise realizations are considered for each distance.}
+100 noise realizations are considered for each distance.
+\tf{We should explain a bit more the symbols and colours that appear in the plots.}}
   \label{fig:s20results}
@@
 
-We have tested that the method does not depend on features of {\texttt s20S} using the 7
-other waveforms of the {\it test set} described in section \ref{sec:simulations} covering
-a large range of progenitor masses.
+For each of the noise realizations we reconstruct the ratio time series {$r_i$} of length $N$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}.
+The reconstructed ratio is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
+The top panel of Fig.~\ref{fig:s20results} shows the distribution of the fraction of the ratio values {$r_i^0$} that fall within the 95\% confidence interval of {$r_i$} as a function of the distance to the source.
+This quantity, hence, gives information about the {\it coverage} of the reconstructed ratio.
+The coverage takes maximum values when the source is located within a few kpc and then decreases with the distance.
 
-Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from {\texttt s11} and to a lesser extent
+To better quantify how well we reconstruct the ratio we also consider  the mean of the relative error of $r_i$ along 
+the track of the spectrogram, $\Delta$,  
+\begin{equation}
+\Delta=\frac{1}{N}\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}\,.
+\end{equation}
+The values of $\Delta$ for each of the 100 noise realizations are shown as a function of the distance
+in the bottom panel of Fig.~\ref{fig:s20results}. For a source located up to $\sim$\unit[9]{kpc} the relative error
+remains smaller than 20\%.
+At closer distances $\Delta$ is smaller but it does not vanish.
+This reflects the approximate nature of the model used for $r$.
+It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
+precision up to distances of $\sim$ 9 kpc (for this particular waveform) with a coverage value
+larger than 80\%. We note that there are a few noise realizations for distances below 9 kpc for which
+$\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those case.
+\tf{Do we know why that is the case?
+What is intrinsically different from one noise realisation to another that makes the method more prone to failure?
+I think we should provide some explanation here, if we have it.}
+
+We have tested that the method does not depend on the specific features of the waveform of model {\texttt s20S} by repeating the procedure for the remaining seven waveforms of the {\it test set} described in Section \ref{sec:simulations} covering
+a large range of progenitor masses. Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from {\texttt s11} and to a lesser extent
 {\texttt s20S}, the ratio is well reconstructed for all waveforms up to $\sim$ 15kpc. In an

### 2020-11-07 methods.tex
Further work on the paper.
Some further work on the paper.

@@
 % !TEX root = ccsn.tex
-\section{Methods description}
+\section{Description of the method}
 \label{methods}
@@
 signal is embedded in noise.
-\tf{We also note that despite we have only explicitely shown results for the GW signal of the {\texttt s20S} model the same conclusions hold for any of the other waveforms of our test set.}
+We also note that despite we have only explicitely shown results for the GW signal of the {\texttt s20S} model the same conclusions hold for any of the other waveforms of our test set.
 

### 2020-11-07 results.tex
Some further work on the paper.

@@
 For each of the noise realizations we reconstruct the ratio time series {$r_i$} of length $N$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}.
+\tf{Perhaps we should mention that this value is chosen using the information on the initial mode frequency from the simulations.}
 The reconstructed ratio is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_coverage_allwvfs}
+ \caption{Median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source.
+The ``no signal'' line and band show the median and first and third quartile of coverage in absence of any signal.
+The line of the latter is zero and overlaps with the horizontal axis (see main text for details).
+\tf{Perhaps is more consistent to use in the legend the same model names as in Table 1.
+Some of the colours are really hard to see, particularly for s20S.
+That could perhaps be improved if we use thicker lines.
+I'm confused about the quartiles shown.}} \label{fig:aLIGO_cov_allwvf}
+\end{figure}
+
 We have tested that the method does not depend on the specific features of the waveform of model {\texttt s20S} by repeating the procedure for the remaining seven waveforms of the {\it test set} described in Section \ref{sec:simulations} covering
-a large range of progenitor masses. Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from {\texttt s11} and to a lesser extent
-{\texttt s20S}, the ratio is well reconstructed for all waveforms up to $\sim$ 15kpc. In an
-effort to better determine the maximal distance of the source at which we can reconstruct the ratio
-we have run 100 simulations without injecting a signal and have measured $coverage$ for the
-reconstructed ratios.
-The median of $coverage$ as well as the 95 quartile are shown on Figure \ref{fig:aLIGO_cov_allwvf}.
-The noise only median value is null in this case, but it can be different from zero because
-the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
+a large range of progenitor masses.
+Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from model {\texttt s11} and to a lesser extent model {\texttt s20S}, the ratio is well reconstructed for all waveforms up to a distance of $\sim$ 15kpc.
+In an
+effort to better determine the maximal distance of the source at which we can reconstruct the ratio we have run 100 simulations without injecting a signal and have measured the corresponding coverage for the reconstructed ratios.
+The median of the coverage as well as the 95 quartile are shown in Figure \ref{fig:aLIGO_cov_allwvf}.
+\tf{This seems inconsistent with what is written in the caption, which indicates the third quartile.}
+The noise only median value is identically zero in this case. However, note that it could be different from zero because
+the g-mode reconstruction algorithm is looking for a continuously  increasing frequency track
 in the spectrogram, starting between 0 and 200 Hz, where we expect the GW signal to be.
-This is enhancing the probability of overlap. This effects explains why outliers can reach
-values as high as 80\%.
-Figure \ref{fig:aLIGO_prec_allwvf} shows $\Delta$ as function of the distance for the same signals
+This is enhancing the probability of overlap. This effect explains why outliers can reach
+values as high as 80\%. \tf{I don't understand this last comment.}
+
+Figure \ref{fig:aLIGO_prec_allwvf} shows the relative error $\Delta$ as a function of the distance for the signals of the {\it test set}
 as well as the result when only noise is considered.
-In Table \ref{tab:results} we are reporting the distance $d_r$ at which $coverage$ median is lower
-than 95\% of the noise only values. We have checked that $coverage$ and $\Delta$ provide similar
-values.
+This quantity follows the same trend than that followed by the coverage, since all signals but models {\texttt s11} and {\texttt s20S} are reconstructed with relative errors below 20\% up to distances of $\sim$ 15kpc.
+Correspondingly, the no-signal case yields the largest error, as expected.
+
+Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors.
+We have checked that using either the median of the coverage or the median of $\Delta$ provide similar results for the distance.
 These numbers are an estimate of the order of magnitude of the source maximal distance at which a
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{plots/aLIGO_coverage_allwvfs}
- \caption{Median of $coverage$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
-The ``no signal'' line and band show the median and first and third quartile of $coverage$ in absence of any signal.} \label{fig:aLIGO_cov_allwvf}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
- \caption{Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
-The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.} \label{fig:aLIGO_prec_allwvf}
+ \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$. \tf{Same comments as for Fig.~\ref{fig:aLIGO_cov_allwvf}.}}
+% Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
+The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.}
+\label{fig:aLIGO_prec_allwvf}
 \end{figure}

### 2020-11-07 results.tex
A bit more work.

@@
 \label{sec:results}
-
-To estimate how accurately we can infer the time evolution of $r(t)$ in the GW data of a single
-detector we inject the GW signal for model {\texttt s20S} into 
-100 Gaussian noise realisations whose power spectral density (PSD) follows the aLIGO
-spectrum~\cite{aLIGOsens:2018}. The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
 
@@
 \end{figure}
-
-We cover a large range of distances for which a CCSN detection in second-generation GW detectors is feasible.
-We assume that the source is optimally oriented with respect to our single detector.
-Moreover, we also assume that a CCSN GW signal has been identified in the data and that the beginning of the signal is known within {$\mathcal O$}(10 ms).
-The data (signal embedded in noise) are whitened using the function {\tt prewhiten} of the R-package {\tt TSA}.
-An auto-regressive model with a maximum of maximal 100 coefficients is used.
 
@@
 \end{figure}
+
+\tf{We probably need a better title for this section.}
+
+To estimate how accurately we can infer the time evolution of $r(t)$ in the GW data of a single
+detector we inject the GW signal for model {\texttt s20S} into 
+100 Gaussian noise realisations whose power spectral density (PSD) follows the aLIGO
+spectrum~\cite{aLIGOsens:2018}. The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
+
+We cover a large range of distances for which a CCSN detection in second-generation GW detectors is feasible.
+We assume that the source is optimally oriented with respect to our single detector.
+Moreover, we also assume that a CCSN GW signal has been identified in the data and that the beginning of the signal is known within {$\mathcal O$}(10 ms).
+The data (signal embedded in noise) are whitened using the function {\tt prewhiten} of the R-package {\tt TSA}.
+An auto-regressive model with a maximum of maximal 100 coefficients is used.
 
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
+ \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$. \tf{Same comments as for Fig.~\ref{fig:aLIGO_cov_allwvf}.}}
+% Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
+The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.}
+\label{fig:aLIGO_prec_allwvf}
+\end{figure}
+
 We have tested that the method does not depend on the specific features of the waveform of model {\texttt s20S} by repeating the procedure for the remaining seven waveforms of the {\it test set} described in Section \ref{sec:simulations} covering
@@
 
+We perform the same analysis using the design sensitivity curve of AdV and expected sensitivity curves for the third-generation 
+GW detectors.
+The results for the former are reported in Table \ref{tab:results}.
+We focus now on third-generation detectors whose results are reported in Table \ref{tab:results} and shown in Fig.~\ref{fig:s20--SFHo_all3G}.
+In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power low-frequency cryogenic interferometers as well as three high-power high-frequency interferometers.
+Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study (cf.~Fig.~\ref{fig:spectrum}).
+The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
+sensitivity circa 2040 through two phases labeled CE1 and CE2, whose sensitivity curves are also shown in Figure~\ref{fig:spectrum}.
+
+%%%
+
 Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors.
-We have checked that using either the median of the coverage or the median of $\Delta$ provide similar results for the distance.
-These numbers are an estimate of the order of magnitude of the source maximal distance at which a
-reconstruction of the ratio could be possible with current GW detectors.
+We have checked that using either the median of the coverage or the median of $\Delta$ yield similar results for the distance.
+The numbers on the table are an estimate of the order of magnitude of the source maximal distance at which a
+reconstruction of the ratio could be possible with current and planned GW detectors.
 They are also upper limits as we are taking into account the detector antenna response in our
-simulation but consider the source is optimally oriented. 
-Table \ref{tab:results} reports also $d_r$ for the Advanced Virgo detector at design sensitivity.
-Results are very similar to aLIGO, despite the detector sensitivity differences.
-Note that Table \ref{tab:results} provides the distance at which one could detect a source
-optimally oriented with a matched filter signal-to-noise ratio of 13.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
- \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$. \tf{Same comments as for Fig.~\ref{fig:aLIGO_cov_allwvf}.}}
-% Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
-The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.}
-\label{fig:aLIGO_prec_allwvf}
-\end{figure}
-
-The same analysis has been performed using expected sensitivity curves for the third generation of
-GW detectors. In Europe the Einstein Telescope project proposes to host in a 10-km
-equilateral triangle configuration 3 low-power low-frequency cryogenic interferometers as well as 3
-high-power high-frequency interferometers. Three sensitivity curves, ET-B, ET-C and ET-D corresponding
-to different options and stages of the project~\cite{Hild_2011} are considered in this study.
-The US based project Cosmic Explorer~\cite{reitze2019cosmic} is proposing to reach its design
-sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in Figure~\ref{fig:spectrum}.
+simulation but assume that the source is optimally oriented with a matched filter signal-to-noise ratio of 13. 
+Table \ref{tab:results} shows that the results for the Advanced Virgo detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity.
+%Note that Table \ref{tab:results} provides the distance at which one could detect a source
+%optimally oriented with a matched filter signal-to-noise ratio of 13.
 
 Figure \ref{fig:s20--SFHo_all3G} shows $\Delta$ as function of the source distance for {\texttt s20S}
-waveform for the five 3G detector configurations. Overall, the ratio is well reconstructed up to distances
+waveform for the five third-generation detector configurations we analyze. Overall, the ratio is well reconstructed up to distances
 in the range \unit[100--200]{kpc} which represents an improvement of a factor 10 with respect to

### 2020-11-08 results.tex
More work. Almost done.

@@
 % !TEX root = ccsn.tex
-\section{Detection sensitivity with Advanced gravitational wave detectors}
+%\section{Detection sensitivity with Advanced gravitational wave detectors}
+\section{Detectability prospects}
 \label{sec:results}
@@
 \end{figure}
-
-\tf{We probably need a better title for this section.}
 
@@
 
-We perform the same analysis using the design sensitivity curve of AdV and expected sensitivity curves for the third-generation 
+We perform the same analysis using the design sensitivity curve of AdV and expected sensitivity curves for third-generation 
 GW detectors.
-The results for the former are reported in Table \ref{tab:results}.
-We focus now on third-generation detectors whose results are reported in Table \ref{tab:results} and shown in Fig.~\ref{fig:s20--SFHo_all3G}.
-In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power low-frequency cryogenic interferometers as well as three high-power high-frequency interferometers.
+The results for the former are reported in Table \ref{tab:results} and are discussed below.
+We focus now on third-generation detectors, presenting our findings in Table \ref{tab:results} and in Fig.~\ref{fig:s20--SFHo_all3G}.
+In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers.
 Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study (cf.~Fig.~\ref{fig:spectrum}).
 The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
-sensitivity circa 2040 through two phases labeled CE1 and CE2, whose sensitivity curves are also shown in Figure~\ref{fig:spectrum}.
+sensitivity circa 2040 through two phases labeled CE1 and CE2. Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.
 
-%%%
+Figure \ref{fig:s20--SFHo_all3G} displays $\Delta$ as a function of the source distance for the five third-generation detector configurations we analyze.
+As before, we use the {\texttt s20S} waveform model as a reference case.
+This figure shows that,  overall, the ratio is well reconstructed up to distances in the range \unit[100--200]{kpc} which represents an improvement of a factor 10 with respect to Advanced LIGO and Advanced Virgo detectors.
+We can also note that the Einstein Telescope results lay in between the 2 Cosmic Explorer results.
+This is confirmed for all other waveforms, expect {\texttt s25}
+for which the maximal distance reach in CE2 is significantly lower than CE1. This is partly due to the
+small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$
+rather uncertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
+\ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
+located up to several hundred of kpc. It is nevertheless important to note the rather wide range obtained
+for the different waveforms probing a large range of progenitor masses.
+We did not find any correlation between the mass of the progenitor and $d_r$, nor the equation of state.
+On the other hand, the quality
+of the ratio reconstruction depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by
+$d_{det}$.
 
@@
 %optimally oriented with a matched filter signal-to-noise ratio of 13.
-
-Figure \ref{fig:s20--SFHo_all3G} shows $\Delta$ as function of the source distance for {\texttt s20S}
-waveform for the five third-generation detector configurations we analyze. Overall, the ratio is well reconstructed up to distances
-in the range \unit[100--200]{kpc} which represents an improvement of a factor 10 with respect to
-Advanced LIGO and Advanced Virgo detectors. We can also note that the Einstein Telescope results lay in
-between the 2 Cosmic Explorer results. This is confirmed for all other waveforms, expect {\texttt s25}
-for which the maximal distance reach in CE2 is significantly lower than CE1. This is partly due to the
-small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$
-rather uncertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
-\ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
-located up to several hundred of kpc. It is nevertheless important to note the rather wide range obtained
-for the different waveforms probing a large range of progenitor masses.
-We did not find any correlation between the mass of the progenitor and $d_r$, nor the equation of state.
-On the other hand, the quality
-of the ratio reconstruction depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by
-$d_{det}$.
 

### 2020-11-08 results.tex
Revised manuscript up to conclusions. Moving to those next.

@@
 Stage 1 (CE1) is expected to use the technology developed for the ``A+'' upgrade to aLIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
-} \label{fig:spectrum}
+\tf{Use AdV in the legend instead of ADV.}}
+ \label{fig:spectrum}
 \end{figure}
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
+  \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.}
+  \label{fig:s20--SFHo_all3G}
+\end{figure}
+
 Figure \ref{fig:s20--SFHo_all3G} displays $\Delta$ as a function of the source distance for the five third-generation detector configurations we analyze.
 As before, we use the {\texttt s20S} waveform model as a reference case.
-This figure shows that,  overall, the ratio is well reconstructed up to distances in the range \unit[100--200]{kpc} which represents an improvement of a factor 10 with respect to Advanced LIGO and Advanced Virgo detectors.
-We can also note that the Einstein Telescope results lay in between the 2 Cosmic Explorer results.
-This is confirmed for all other waveforms, expect {\texttt s25}
-for which the maximal distance reach in CE2 is significantly lower than CE1. This is partly due to the
-small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$
-rather uncertain for this waveform. All results are summarized in Table \ref{tab:results} and Figure
-\ref{fig:distances}. It is remarkable that with 3G detectors the ratio could be reconstructed for sources
-located up to several hundred of kpc. It is nevertheless important to note the rather wide range obtained
-for the different waveforms probing a large range of progenitor masses.
-We did not find any correlation between the mass of the progenitor and $d_r$, nor the equation of state.
-On the other hand, the quality
-of the ratio reconstruction depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by
-$d_{det}$.
+This figure shows that,  overall, the ratio is well reconstructed up to distances in the range \unit[100--200]{kpc} which represents an order of magnitude improvement with respect to aLIGO and AdV.
+We also note that the results for the various Einstein Telescope configurations lay in between those of the two Cosmic Explorer designs.
+Moreover,  the detectability prospects for the former depend weakly on the detector configuration while the arrangement of CE2 yields better results than CE1.
+These results are confirmed for all other waveforms of our sample except for model {\texttt s25} for which the maximal distance reach in CE2 is significantly lower than CE1, as shown in Fig.~\ref{fig:distances}.
+This is partly due to the small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$ rather uncertain for this particular waveform.
+\tf{Is this really understood?
+What's particular in this waveform?}
 
-Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors.
-We have checked that using either the median of the coverage or the median of $\Delta$ yield similar results for the distance.
-The numbers on the table are an estimate of the order of magnitude of the source maximal distance at which a
-reconstruction of the ratio could be possible with current and planned GW detectors.
-They are also upper limits as we are taking into account the detector antenna response in our
-simulation but assume that the source is optimally oriented with a matched filter signal-to-noise ratio of 13. 
-Table \ref{tab:results} shows that the results for the Advanced Virgo detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity.
-%Note that Table \ref{tab:results} provides the distance at which one could detect a source
-%optimally oriented with a matched filter signal-to-noise ratio of 13.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
-  \caption{Median of $\Delta$ for {\texttt s20S} CCSN waveform embedded in 3G detectors noise and located at different distance from the Earth. } \label{fig:s20--SFHo_all3G}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[t]
   \centering
   \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
-  \caption{Maximal distance $d_{r}$ in kpc at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
-    with good accuracy for a source optimally oriented with respect to the GW detectors for the 7 CCSN waveforms considered in this study.} \label{fig:distances}
+  \caption{Maximal distance at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
+    with good accuracy for all GW detectors analyzed in this study.
+The values are shown for all eight CCSN waveforms of our sample assuming that the source is optimally oriented with respect to the detector.
+\tf{I guess that for CE2 s15G is overlapping with s25.
+Should we mention this in the caption?
+Use AdV instead of ADV?}}
+\label{fig:distances}
 \end{figure}
@@
 \cline{2-10}
-                       & $d_{det}$ & 11 & 36 & 26 & 27 & 21 & 16 & 74 & 61\\
+                       & $d_{\rm det}$ & 11 & 36 & 26 & 27 & 21 & 16 & 74 & 61\\
 %                      & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\
@@
 \hline
-\multirow{3}{*}{ADV}   & $d_{r}$   & 7  & 26  & 20 & 19 & 15 & 10 & 43 & 42 \\
+\multirow{3}{*}{AdV}   & $d_{r}$   & 7  & 26  & 20 & 19 & 15 & 10 & 43 & 42 \\
 \cline{2-10}
-                       & $d_{det}$ &  10 & 32 & 22 & 23 & 18 & 13 & 64 & 52\\
+                       & $d_{\rm det}$ &  10 & 32 & 22 & 23 & 18 & 13 & 64 & 52\\
 %                      & SNR      &  13 & 41 & 29 & 30 & 23 & 17 &  83 & 68\\
@@
 \cline{2-10}
-                       & $d_{det}$ & 115 & 377 & 270 & 282 & 217 & 168 & 774  & 633\\
+                       & $d_{\rm det}$ & 115 & 377 & 270 & 282 & 217 & 168 & 774  & 633\\
 %                      & SNR      & 149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\
@@
 \cline{2-10}
-                       & $d_{det}$ & 197 & 649 & 468 & 489 & 375 & 294 & 1347  & 1100\\
+                       & $d_{\rm det}$ & 197 & 649 & 468 & 489 & 375 & 294 & 1347  & 1100\\
 %                      & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\
@@
 \cline{2-10}
-                       & $d_{det}$ & 106 & 364 & 274 & 391 & 216 & 200 & 805 & 665\\
+                       & $d_{\rm det}$ & 106 & 364 & 274 & 391 & 216 & 200 & 805 & 665\\
 %                      & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\
@@
 \cline{2-10}
-                       & $d_{det}$ & 97 & 332 & 246 & 260 & 194 & 164 & 727  & 603\\
+                       & $d_{\rm det}$ & 97 & 332 & 246 & 260 & 194 & 164 & 727  & 603\\
 %                      & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\
@@
 \cline{2-10}
-                       & $d_{det}$ & 107 & 368 & 271 & 285 & 213 & 174 & 796  & 661\\
+                       & $d_{\rm det}$ & 107 & 368 & 271 & 285 & 213 & 174 & 796  & 661\\
 %                      & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\
@@
   Maximal distance $d_{r}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
-    with good accuracy for a source optimally oreinted with respect to the GW detectors
-    considered in this study. $d_{det}$ is the distance at which one could detect a source
-    optimally oriented with a matched filter signal-to-noise ratio of 13 in the  different
-    GW detectors. All distances are expressed in kpc.
+    with good accuracy for all GW detectors analyzed in this study, assuming optimal orientation between the source and the detector.
+Correspondingly, $d_{\rm det}$ is the distance at which different interferometers could detect a source
+    optimally oriented with a matched filter signal-to-noise ratio of 13. All distances are expressed in kpc.
     %%
@@
 \end{table}
+
+All results for both second-generation and third-generation detectors are summarized in Table \ref{tab:results} and Figure
+\ref{fig:distances}.
+Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors.
+This same information is displayed in Fig.~\ref{fig:distances}.
+We have checked that using either the median of the coverage or the median of $\Delta$ yields similar results for the distance.
+On the other hand, the quality of the ratio reconstruction and, thus, of the distance range, depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by $d_{\rm det}$.
+The numbers reported on the table for $d_r$ are an estimate of the order of magnitude of the maximal distance of the source  at which a
+reconstruction of the ratio could be possible with current and planned GW detectors.
+We also provide upper limits for $d_{\rm det}$ by taking into account the detector antenna response in our simulations and assuming that the source is optimally oriented with a matched filter signal-to-noise ratio of 13.
+Table \ref{tab:results} shows that the results for the Advanced Virgo detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity.
+It is remarkable that for third-generation detectors the PNS surface gravity  could be reconstructed for sources located up to a few hundreds of kpc.
+It is nevertheless important to note the rather wide range in distances we obtain for the different waveforms of our {\it test set} that probe a large range of progenitor masses.
+We do not find any correlation between either the mass of the progenitor nor the EOS with $d_r$.

### 2020-11-08 conclusion.tex
More work on the paper.

@@
-\section{Conclusion}
+% !TEX root = ccsn.\TeX
+\section{Conclusions}
 \label{sec:conclusion}
-The algorithm presented in this paper is a first attempt to infer the time evolution of a
-combinaison of the mass of the PNS and its radius based on the universal relations found
-in PNS asteroseismology. More precisely, we have considered in this paper the ratio
-$r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$
-oscillation mode in the GW data. We have especially investigated the performance of the algorithm
-in the case of an optimally oriented source detected in a singe GW detector. For Advanced LIGO
-or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
-that this is true for a wide range of progenitor masses and that the quality of the inference
-mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
-as Einstein Telescope and Cosmic Explorer, the $\mbox{}^2g_2$ will be reconstructible for sources
-at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
-for all waveforms considered here thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
-Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
-especially for the waveforms with the highest progenitor mass (25 $\Msol$ and 40 $\Msol$). Comparing
-$d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
-below \unit[200]{Hz} is important for massive mass progenitor signals.
 
-%The model to infer the ratio as function of the frequency is based on 1D {\sc AENUS-ALCAR} simulations.
-%We have observed that using the other numerical code {\sc CoCoNut} inputs lead to systematic bias of the
-%ratio reconstruction.
+The detection of GW from CCSN may help improve our current understanding of the explosion mechanism of massive stars.  
+In this paper we have provided estimates of the detectability prospects of CCSN signals using an approach based on GW associated with convective oscillations of PNS.
+As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.
+Such modes are responsible for a significant fraction of the highly stochastic GW emitted after core bounce.
+The findings reported in this paper suggest that PNS asteroseismology might be within reach of current and third-generation GW detectors.
 
-This study does not include the realistic case of operating within a network of detectors.
-The sources of GWs we have considered here are optimally oriented. The reported distance at which
-we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an upper limit that may be lower
-by a factor 2--3 on average for a source located anywhere on the sky. We defer a more realistic
-simulation implementation for a forthcoming publication.
+In our study we have used a set of 1D CCSN simulations to build a model that relates the  evolution of PNS properties with the frequency of the dominant g-mode, namely the $\mbox{}^2g_2$ mode.
+This relationship is extracted from the GW data using an algorithm developed for this investigation.
+This algorithm is  a first attempt to infer the time evolution of a particular combination of the PNS mass and radius based on the universal relations found in~\cite{Torres:2019b}.
+More precisely, we have considered the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$ oscillation mode in the numerically generated GW data.
+The  performance of our method has been estimated employing  simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
+
+We have investigated the performance of the algorithm in the case of an optimally oriented source detected in a singe GW detector.
+The signals of our sample have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors we have analyzed.
+For Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~distances of ${\cal O}(10\, {\rm kpc})$).
+This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the signal.
+For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(100\, {\rm kpc})$.
+In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
+Among the three configuration of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the higher progenitor masses (25 $\Msol$ and 40 $\Msol$).
+Comparing the estimated distances for ET-B and the other third-generations detectors, having a good sensitivity below \unit[200]{Hz} seems the most important factor to detect high mass progenitor signals.
+
+In the present study we have assumed that the sources are optimally oriented. The reported distances at which
+we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits.
+Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
+Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network.
+We defer an improved implementation of our approach for a forthcoming publication.
+Finally, we note that the method discussed in this work can be adapted to other PNS oscillation modes, by simply changing a few parameters such as the initial frequency range of the mode and its monotonic raise or descent.
+Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS.
 
@@
 %code.
-
-Finally, this method can be adapted to other PNS oscillation modes, changing few parameters such as
-the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
-reconstruct several modes in the same GW signal would allow to infer individually each of the PNS property.

### 2020-11-08 ccsn.tex
Some work on the intro.
Some further work on the paper.
More work. Almost done.
Finished my revision of the entire paper.

@@
 \newcommand{\mab}[1]{\textcolor{red}{#1}}
-\newcommand{\tf}[1]{\textcolor{green}{#1}}
+\newcommand{\tf}[1]{\textcolor{red}{TF: #1}}
 
@@
 \begin{document}
-\linenumbers
+%\linenumbers
 
-\title{Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae.}
+\title{Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae}
 
-\author{M.~A.~Bizouard}
+\author{Marie-Anne~Bizouard}
 \affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
 
-\author{P.~Maturana-Russel}
+\author{Patricio~Maturana-Russel}
 \affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
@@
 
-\author{A.~Torres-Forn\'e}
+\author{Alejandro~Torres-Forn\'e}
 \affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
@@
 
-\author{M.~Obergaulinger}
+\author{Martin~Obergaulinger}
 \affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
 
-\author{P.~Cerd\'a-Dur\'an}
+\author{Pablo~Cerd\'a-Dur\'an}
 \affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
 
-\author{N.~Christensen}
+\author{Nelson~Christensen}
 \affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
@@
 
-\author{J.~A.~Font}
+\author{Jos\'e~A.~Font}
 \affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}
@@
 
-\author{R.~Meyer}
+\author{Renate~Meyer}
 \affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
@@
 
-\bigskip\noindent\textit{Acknowledgments} ---
+\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071)
+and by  the European Union’s Horizon 2020 research and innovation (RISE)  programme   H2020-MSCA-RISE-2017  GrantNo.  FunFiCO-777740.  \tf{I have included my grants. Please include yours.}
 

### 2020-11-08 conclusion.tex
Finished my revision of the entire paper.

@@
 This algorithm is  a first attempt to infer the time evolution of a particular combination of the PNS mass and radius based on the universal relations found in~\cite{Torres:2019b}.
-More precisely, we have considered the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$ oscillation mode in the numerically generated GW data.
+More precisely, we have considered the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity) derived from the observation of the $\mbox{}^2g_2$ oscillation mode in the numerically generated GW data.
 The  performance of our method has been estimated employing  simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
 
-We have investigated the performance of the algorithm in the case of an optimally oriented source detected in a singe GW detector.
-The signals of our sample have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors we have analyzed.
-For Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~distances of ${\cal O}(10\, {\rm kpc})$).
-This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the signal.
+We have investigated the performance of the algorithm in the case of an optimally oriented source detected by a singe GW detector.
+Our numerical signals have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors analyzed.
+We have found that for Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~for distances of ${\cal O}(10\, {\rm kpc})$).
+This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the event.
 For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(100\, {\rm kpc})$.
 In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
-Among the three configuration of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the higher progenitor masses (25 $\Msol$ and 40 $\Msol$).
+Among the three configurations of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the highest progenitor masses (25 $\Msol$ and 40 $\Msol$).
 Comparing the estimated distances for ET-B and the other third-generations detectors, having a good sensitivity below \unit[200]{Hz} seems the most important factor to detect high mass progenitor signals.
@@
 Finally, we note that the method discussed in this work can be adapted to other PNS oscillation modes, by simply changing a few parameters such as the initial frequency range of the mode and its monotonic raise or descent.
-Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS.
+Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS in core-collapse supernova explosions.
 

### 2020-11-08 introduction.tex
Finished my revision of the entire paper.

@@
 In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
-This} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs.
+These} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs.
 The frequency and time evolution of these modes carry information about the properties of the GW emitter and could allow to perform PNS asteroseismology.
@@
 %
-during the post-bounce evolution of the system, with raising frequency
+during the post-bounce evolution of the system, with frequency rising 
 from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.

### 2020-11-08 methods.tex
Finished my revision of the entire paper.

@@
 The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to
-the Aikaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{eq:model1}.
+the Akaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{eq:model1}.
 The data and fit of the model including 95\% confidence bands are displayed in
@@
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
-autoregressive estimates \pcd{[add ref?]} of the local spectra for successive time intervals of 
+autoregressive estimates~\cite{BrockwellPeterJ1991Tsta} of the local spectra for successive time intervals of 
 length 200 with a 90\% overlap.

### 2020-11-09 ccsn.tex
answer to TF comments

@@
 \newcommand{\pcd}[1]{\textcolor{blue}{#1}}
-\newcommand{\mab}[1]{\textcolor{red}{#1}}
+\newcommand{\mab}[1]{\textcolor{green}{#1}}
 \newcommand{\tf}[1]{\textcolor{red}{TF: #1}}
@@
 \begin{abstract}
-The eventual detection of gravitational waves from core-collapse supernovae (CCSN) may help improve our current understanding of the explosion mechanism of massive stars.
+The eventual detection of gravitational waves from core-collapse supernovae (CCSN) will help improve our current understanding of the explosion mechanism of massive stars.
 The stochastic nature of the late post-bounce gravitational wave signal due to matter effects and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
@@
 The performance of the method is estimated employing simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
-Our results indicate that it is possible to infer PNS properties for a galactic source using Advanced LIGO and Advanced Virgo at design sensitivities.
+Considering signals embedded in Gaussian gravitational wave detector noise, we show that it is possible to infer PNS properties for a galactic source using Advanced LIGO and Advanced Virgo data at design sensitivities.
 Third generation detectors such as Einstein Telescope and Cosmic Explorer will allow to test distances of ${\cal O}(100\, {\rm kpc})$.
-\tf{Caveats?
+%\tf{Caveats?
 Gaussian noise?}
@@
 and by  the European Union’s Horizon 2020 research and innovation (RISE)  programme   H2020-MSCA-RISE-2017  GrantNo.  FunFiCO-777740.  \tf{I have included my grants. Please include yours.}
+N.C. acknowledges support from National Science Foundation grant PHY-1806990.
 

### 2020-11-09 introduction.tex

@@
 A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors.
-The emitted GWs could be detected with the advanced ground-based GW detector network (Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
-KAGRA~\citep{Aso:2013eba}) within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
+The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
+KAGRA~\citep{Aso:2013eba}, within $\sim 5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
 For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within $50$~kpc and, for some extreme models, up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
@@
 For the results discussed in this work we only consider neutrino-driven CCSN.
-\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
+\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
 
@@
 shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
-\tf{This last sentence might be improved.}
+\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations.
+Using results from 2D CCSN numerical simulations \citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
+}
+
 This reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. 

### 2020-11-09 results.tex

@@
 100 noise realizations are considered for each distance.
-\tf{We should explain a bit more the symbols and colours that appear in the plots.}}
+The orange lines indicate the median values while the empty rectangles indicate the first and third quartiles.
+The ``+'' markers are outliers outside the first and third quartiles.}
   \label{fig:s20results}
@@
 For each of the noise realizations we reconstruct the ratio time series {$r_i$} of length $N$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}.
-\tf{Perhaps we should mention that this value is chosen using the information on the initial mode frequency from the simulations.}
+\mab{This value is chosen using the information on the initial mode frequency from the simulations.}
 The reconstructed ratio is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
@@
 larger than 80\%. We note that there are a few noise realizations for distances below 9 kpc for which
-$\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those case.
-\tf{Do we know why that is the case?
+$\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those cases.
+\mab{The main cause of failure is the split of the ridge in few blobs when the signal becomes weaker and weaker.}
+%\tf{Do we know why that is the case?
 What is intrinsically different from one noise realisation to another that makes the method more prone to failure?
@@
  \caption{Median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source.
-The ``no signal'' line and band show the median and first and third quartile of coverage in absence of any signal.
-The line of the latter is zero and overlaps with the horizontal axis (see main text for details).
-\tf{Perhaps is more consistent to use in the legend the same model names as in Table 1.
+\mab{The ``no signal'' line shows the median of coverage in absence of any signal.
+In this example, the median is null and overlaps with the horizontal axis.
+The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal.} \tf{Perhaps is more consistent to use in the legend the same model names as in Table 1.
 Some of the colours are really hard to see, particularly for s20S.
@@
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
- \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$. \tf{Same comments as for Fig.~\ref{fig:aLIGO_cov_allwvf}.}}
+ \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$.}
 % Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
@@
 effort to better determine the maximal distance of the source at which we can reconstruct the ratio we have run 100 simulations without injecting a signal and have measured the corresponding coverage for the reconstructed ratios.
-The median of the coverage as well as the 95 quartile are shown in Figure \ref{fig:aLIGO_cov_allwvf}.
-\tf{This seems inconsistent with what is written in the caption, which indicates the third quartile.}
-The noise only median value is identically zero in this case. However, note that it could be different from zero because
+\mab{The median of the coverage as well as the band defined by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles are shown in Figure \ref{fig:aLIGO_cov_allwvf}.}
+The noise only median value is identically zero in this case. However, note that it could be different
+from zero because
 the g-mode reconstruction algorithm is looking for a continuously  increasing frequency track
 in the spectrogram, starting between 0 and 200 Hz, where we expect the GW signal to be.
-This is enhancing the probability of overlap. This effect explains why outliers can reach
-values as high as 80\%. \tf{I don't understand this last comment.}
+This is enhancing the probability of overlap. This effect explains why certain values of overlap can reach
+values as high as 80\% even when no signal is added to the noise. \tf{I don't understand this last comment.} \mab{Toni: is that clearer?}
 
@@
 The values are shown for all eight CCSN waveforms of our sample assuming that the source is optimally oriented with respect to the detector.
-\tf{I guess that for CE2 s15G is overlapping with s25.
-Should we mention this in the caption?
-Use AdV instead of ADV?}}
+Some waveform maximal distance markers are not visible because the do overlap (for instance \texttt{s15G} is overlapping with \texttt{s25} for CE2).
+\tf{Use AdV instead of ADV?}}
 \label{fig:distances}

### 2020-11-09 introduction.tex
update plots & text

@@
 For the results discussed in this work we only consider neutrino-driven CCSN.
-\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
+%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
 

### 2020-11-09 results.tex
update plots & text

@@
 
-\begin{figure}[t]
- \centering
- \includegraphics[width=0.5\textwidth]{plots/spectrum}
- \caption{Amplitude spectral density of the GW detectors aLIGO and AdV at design sensitivity and that of the proposed third-generation detectors Cosmic Explorer and Einstein Telescope.
+%\begin{figure}[t]
+% \centering
+% \includegraphics[width=0.5\textwidth]{plots/spectrum}
+% \caption{Amplitude spectral density of the GW detectors aLIGO and AdV at design sensitivity and that of the proposed third-generation detectors Cosmic Explorer and Einstein Telescope.
 Einstein Telescope sensitivity curve ET-B is obtained pushing second-generation detector technology at its limit.
@@
 Cosmic Explorer design sensitivity will be achieved in two stages.
-Stage 1 (CE1) is expected to use the technology developed for the ``A+'' upgrade to aLIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.
-\tf{Use AdV in the legend instead of ADV.}}
- \label{fig:spectrum}
-\end{figure}
+Stage 1 (CE1) is expected to use the technology developed for the ``A+'' upgrade to aLIGO but scaled up to a 40 km detector while stage 2 (CE2) will implement state-of-the-art technology to decrease quantum and thermal noises~\cite{reitze2019cosmic}.}
+% \label{fig:spectrum}
+%\end{figure}
 
@@
 100 Gaussian noise realisations whose power spectral density (PSD) follows the aLIGO
-spectrum~\cite{aLIGOsens:2018}. The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
+spectrum~\cite{aLIGOsens:2018}. %The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
 
@@
 In this example, the median is null and overlaps with the horizontal axis.
-The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal.} \tf{Perhaps is more consistent to use in the legend the same model names as in Table 1.
-Some of the colours are really hard to see, particularly for s20S.
-That could perhaps be improved if we use thicker lines.
-I'm confused about the quartiles shown.}} \label{fig:aLIGO_cov_allwvf}
+The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal.}} \label{fig:aLIGO_cov_allwvf}
 \end{figure}
@@
 In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers.
-Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study (cf.~Fig.~\ref{fig:spectrum}).
+Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study.
+%(cf.~Fig.~\ref{fig:spectrum}).
 The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
-sensitivity circa 2040 through two phases labeled CE1 and CE2. Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.
+sensitivity circa 2040 through two phases labeled CE1 and CE2. %Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.
 
@@
 The values are shown for all eight CCSN waveforms of our sample assuming that the source is optimally oriented with respect to the detector.
-Some waveform maximal distance markers are not visible because the do overlap (for instance \texttt{s15G} is overlapping with \texttt{s25} for CE2).
-\tf{Use AdV instead of ADV?}}
+Some waveform maximal distance markers are not visible because the do overlap (for instance \texttt{s15G} is overlapping with \texttt{s25} for CE2).}
 \label{fig:distances}

### 2020-11-09 ccsn.tex
My last comments.

@@
 The eventual detection of gravitational waves from core-collapse supernovae (CCSN) will help improve our current understanding of the explosion mechanism of massive stars.
-The stochastic nature of the late post-bounce gravitational wave signal due to matter effects and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
-In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with convective oscillations of proto-neutron stars (PNS).
+The stochastic nature of the late post-bounce gravitational wave signal due to \pcd{\sout{matter effects} the non-linear dynamics of the matter involved} and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
+In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with \pcd{\sout{convective}} oscillations of proto-neutron stars (PNS)
+\pcd{[PCD: the oscillations are driven by convection and other instabilities but I would not call them convective oscillations.]}.
 Numerical simulations of CCNS have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
 We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.
-The model is used to infer the time evolution of combinations of the mass and the radius of the PNS.
+The model is used to infer the time evolution of combinations \pcd{[PCD: "of a combination"?
+(there is only one)]} of the mass and the radius of the PNS.
 The performance of the method is estimated employing simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
@@
 
-\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071)
-and by  the European Union’s Horizon 2020 research and innovation (RISE)  programme   H2020-MSCA-RISE-2017  GrantNo.  FunFiCO-777740.  \tf{I have included my grants. Please include yours.}
+\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071) and by the European Union’s Horizon 2020 research and innovation (RISE) programme H2020-MSCA-RISE-2017 GrantNo.
+FunFiCO-777740.
 N.C. acknowledges support from National Science Foundation grant PHY-1806990.
-
+R.M.
+gratefully acknowledges support by the James Cook Fellowship from Government funding, administered by the Royal Society Te  Ap\={a}rangi.
+R.M.
+and P.M.
+also acknowledge funding from  DFG Grant KI 1443/3-2 and thank the Centre for eResearch at the University of Auckland for their technical support.
+P.C.-D. acknowledges the support of the Spanish {\it Ramon y Cajal} programme (RYC-2015-19074) supporting his research.
 \begin{appendices}

### 2020-11-09 conclusion.tex
My last comments.

@@
-% !TEX root = ccsn.\TeX
+% !TEX root = ccsn.tex
 \section{Conclusions}
@@
 The detection of GW from CCSN may help improve our current understanding of the explosion mechanism of massive stars.  
-In this paper we have provided estimates of the detectability prospects of CCSN signals using an approach based on GW associated with convective oscillations of PNS.
+In this paper we have provided estimates of the detectability prospects of CCSN signals using an approach based on GW associated with convective oscillations of PNS \pcd{[PCD: I wouldn't say "detectability prospects" since we are doing parameter estimation]}.
 As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.

### 2020-11-09 introduction.tex
My last comments.

@@
 However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs
-\citep{Chapman:2007}) making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}.
-For the results discussed in this work we only consider neutrino-driven CCSN.
+\citep{Chapman:2007}) making this possibility a subdominant channel of detection with \pcd{a galactic} event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}\pcd{[PCD: This number is just ~1\% of the galactic rate of 2-3 per century, so no reference is needed]}.
+For the results discussed in this work we only consider neutrino-driven CCSN \pcd{of non-rotating stellar cores} .
 %\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
@@
 
-Using results from 2D CCSN numerical simulations as a background with respect to which to solve an eigenvalue problem~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
+Using results from 2D CCSN numerical simulations as a background \pcd{\sout{with respect to which to solve an eigenvalue problem}}~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
 shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
@@
 Using results from 2D CCSN numerical simulations \citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
-}
+} \pcd{[PCD: I am not convinced about the first sentence of mab's proposal.
+Maybe if we just delete "with respect to which to solve an eigenvalue problem"
+the sentence becomes more readable]}
 

### 2020-11-09 methods.tex
update plots & text
My last comments.

@@
  \includegraphics[width=0.45\textwidth]{plots/model}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 25 1D simulations of the model set.
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our \pcd{18} 1D simulations of the model set.
 The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
-\tf{18 or 25?} \tf{We could remove the legend as only one code is used.} }
+}
  \label{fig:LMVAR}
@@
 with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}
-\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.
+\cite{lmvar:2019} \pcd{[PCD: reference is empty]} that implements a maximum likelihood approach was used to fit the model.
 
@@
 An example is given in Figure \ref{fig:ratio} where the red triangles are the point estimates and
-the grey bands represent 95\% confidence bands.
-The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode estimates.
-Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance.
-Ratio values
+the grey bands represent 95\% confidence bands. The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode \pcd{frequency?} estimates.  
+\pcd{[PCD: By magnitude you mean the value of the frequency? ]}
+Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance. Ratio values
 computed using the mass and radius values obtained from the simulation code (i.e.~the true values)
@@
  \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles).
-The ratios are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode estimates.}
+The ratios \pcd{[PCD: size of the triangles?]} are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode \pcd{frequency?} estimates.}
  \label{fig:ratio}

### 2020-11-09 results.tex
My last comments.

@@
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_coverage_allwvfs}
- \caption{Median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source.
+ \caption{Median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source\pcd{[PCD: if the solid line is the median, what are the symbols?
+Same for figures 6 and 7.]}.
 \mab{The ``no signal'' line shows the median of coverage in absence of any signal.
@@
   \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
-  \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.}
+  \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.
+\pcd{[PCD: Should we show the corresponding
+  blue ranges for the pure noise case?]}}
   \label{fig:s20--SFHo_all3G}

### 2020-11-09 simulations.tex
My last comments.

@@
 
-For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
+For the {\it model set}, we use the \pcd{$18$ }spherically symmetric (1D) simulations of \citep{Torres:2019a}
 including progenitors with zero-age main sequence (ZAMS) masses in the range 

### 2020-11-09 ccsn.tex
Pablo s comemnts answer

@@
 The eventual detection of gravitational waves from core-collapse supernovae (CCSN) will help improve our current understanding of the explosion mechanism of massive stars.
-The stochastic nature of the late post-bounce gravitational wave signal due to \pcd{\sout{matter effects} the non-linear dynamics of the matter involved} and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
-In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with \pcd{\sout{convective}} oscillations of proto-neutron stars (PNS)
-\pcd{[PCD: the oscillations are driven by convection and other instabilities but I would not call them convective oscillations.]}.
+The stochastic nature of the late post-bounce gravitational wave signal due to the non-linear dynamics of the matter involved and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
+In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS)
 Numerical simulations of CCNS have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
 We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.
-The model is used to infer the time evolution of combinations \pcd{[PCD: "of a combination"?
-(there is only one)]} of the mass and the radius of the PNS.
+The model is used to infer the time evolution of a combination of the mass and the radius of the PNS.
 The performance of the method is estimated employing simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
@@
 R.M.
-and P.M.
+and P.M.-R.
 also acknowledge funding from  DFG Grant KI 1443/3-2 and thank the Centre for eResearch at the University of Auckland for their technical support.

### 2020-11-09 conclusion.tex
Pablo s comemnts answer

--- 2020-11-08 conclusion.tex
@@
 The detection of GW from CCSN may help improve our current understanding of the explosion mechanism of massive stars.  
-In this paper we have provided estimates of the detectability prospects of CCSN signals using an approach based on GW associated with convective oscillations of PNS.
+In this paper we have proposed an explorary method to infer PNS properties using an approach based on GW associated with convective oscillations of PNS.
 As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.

### 2020-11-09 introduction.tex
Pablo s comemnts answer

--- 2020-11-09 introduction.tex
@@
 For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within $50$~kpc and, for some extreme models, up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
-However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs
-\citep{Chapman:2007}) making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}.
-For the results discussed in this work we only consider neutrino-driven CCSN.
-\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN} Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
+However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim 10^{-4} \rm{yr}^{-1}$.
+%For the results discussed in this work we only consider neutrino-driven CCSN of non-rotating stellar cores.%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN}
+Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
 
@@
 
-Using results from 2D CCSN numerical simulations as a background with respect to which to solve an eigenvalue problem~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
+Using results from 2D CCSN numerical simulations as a background~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
 shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
-\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations.
+%\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations.
 Using results from 2D CCSN numerical simulations \citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
-}
+} \pcd{[PCD: I am not convinced about the first sentence of mab's proposal.
+Maybe if we just delete "with respect to which to solve an eigenvalue problem" the sentence becomes more readable]}
 

### 2020-11-09 results.tex
Pablo s comemnts answer

@@
 $\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those cases.
-\mab{The main cause of failure is the split of the ridge in few blobs when the signal becomes weaker and weaker.}
+The main cause of failure is the split of the ridge in few blobs when the signal becomes weaker and weaker.
 %\tf{Do we know why that is the case?
@@
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_coverage_allwvfs}
- \caption{Median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source\pcd{[PCD: if the solid line is the median, what are the symbols?
-Same for figures 6 and 7.]}.
-\mab{The ``no signal'' line shows the median of coverage in absence of any signal.
+ \caption{Data points (``+'') show the median of the coverage for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the  distance to the source.
+The solid lines are smoothing splines.
+The ``no signal'' line shows the median of coverage in absence of any signal.
 In this example, the median is null and overlaps with the horizontal axis.
-The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal.}} \label{fig:aLIGO_cov_allwvf}
+The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal.} \label{fig:aLIGO_cov_allwvf}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{plots/aLIGO_delta_allwvfs}
- \caption{Same as Fig.~\ref{fig:aLIGO_cov_allwvf} but for the relative error $\Delta$.}
-% Median of $\Delta$ for 8 CCSN waveforms embedded in aLIGO noise and located at different distance from the Earth.
-The ``no signal'' line and band show the median and first and third quartile of $\Delta$ in absence of any signal.}
-\label{fig:aLIGO_prec_allwvf}
+  \caption{Data points (``+'') show the median of the relative error $\Delta$ for the eight CCSN waveforms of the {\it test set} embedded in aLIGO noise as a function of the distance to the source.
+The solid lines are smoothing splines.
+The ``no signal'' line shows the median of $\Delta$ in absence of any signal.
+The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of $\Delta$ in absence of any signal.}\label{fig:aLIGO_prec_allwvf}
 \end{figure}
@@
 effort to better determine the maximal distance of the source at which we can reconstruct the ratio we have run 100 simulations without injecting a signal and have measured the corresponding coverage for the reconstructed ratios.
-\mab{The median of the coverage as well as the band defined by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles are shown in Figure \ref{fig:aLIGO_cov_allwvf}.}
+The median of the coverage as well as the band defined by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles are shown in Figure \ref{fig:aLIGO_cov_allwvf}.
 The noise only median value is identically zero in this case. However, note that it could be different
@@
   \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.
-\pcd{[PCD: Should we show the corresponding
-  blue ranges for the pure noise case?]}}
+\pcd{[PCD: Should we show the corresponding blue ranges for the pure noise case?]}\mab{MAB: We would need to show 5 noise ranges.
+I haven't tried what it looks like.
+Could give a try}}
   \label{fig:s20--SFHo_all3G}

### 2020-11-10 conclusion.tex
prep version to be circulated in LVK

@@
 This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the event.
-For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(100\, {\rm kpc})$.
+For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(\unit[100]{kpc})$.
 In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.

### 2020-11-10 introduction.tex
prep version to be circulated in LVK

@@
 The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
-KAGRA~\citep{Aso:2013eba}, within $\sim 5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
-For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within $50$~kpc and, for some extreme models, up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}.
-However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim 10^{-4} \rm{yr}^{-1}$.
+KAGRA~\citep{Aso:2013eba}, within $\sim$ \unit[5]{kpc} \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
+For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within \unit[50]{kpc} and, for some extreme models, up to \unit[5--30]{Mpc} \citep{Gossan:2016,TargetedSNSearchO12}.
+However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim$ \unit[$10^{-4}$]{$\rm {yr}^{-1}$}.
 %For the results discussed in this work we only consider neutrino-driven CCSN of non-rotating stellar cores.%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN}
@@
 during the post-bounce evolution of the system, with frequency rising 
-from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s.
+from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}.
 This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.

### 2020-11-10 methods.tex
Pablo s comemnts answer
prep version to be circulated in LVK

--- 2020-11-08 methods.tex
@@
  \includegraphics[width=0.45\textwidth]{plots/model}
- \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 25 1D simulations of the model set.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
-\tf{18 or 25?} \tf{We could remove the legend as only one code is used.} }
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set.
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.}
  \label{fig:LMVAR}
@@
 1D simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
-shows the data for the $25$ numerical simulations. 
+shows the data for the $18$ numerical simulations. 
 %As identified by \cite{Torres:2019b}, the only systematic deviation from a single universal relation is the numerical code used in the simulations. 
@@
 of model {\texttt s20S}, originally
-sampled at 16384 Hz but downsampled at 4096 Hz.
+sampled at \unit[16384]{Hz} but downsampled at \unit[4096]{Hz}.
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
@@
 An example is given in Figure \ref{fig:ratio} where the red triangles are the point estimates and
-the grey bands represent 95\% confidence bands.
-The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode estimates.
-Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance.
-Ratio values
+the grey bands represent 95\% confidence bands. The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.
+Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance. Ratio values
 computed using the mass and radius values obtained from the simulation code (i.e.~the true values)
@@
  \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles).
-The ratios are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode estimates.}
+The size of the triangles are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.}
  \label{fig:ratio}

### 2020-11-10 results.tex
prep version to be circulated in LVK

@@
 
-For each of the noise realizations we reconstruct the ratio time series {$r_i$} of length $N$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}.
-\mab{This value is chosen using the information on the initial mode frequency from the simulations.}
-The reconstructed ratio is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
+For each of the noise realizations we reconstruct the track ridge $m(t_i)$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}.
+This value is chosen using the information on the initial mode frequency from the simulations.
+We derive the ratio time series $r_i$ of length $N$ which is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.
 The top panel of Fig.~\ref{fig:s20results} shows the distribution of the fraction of the ratio values {$r_i^0$} that fall within the 95\% confidence interval of {$r_i$} as a function of the distance to the source.
@@
 It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
-precision up to distances of $\sim$ 9 kpc (for this particular waveform) with a coverage value
-larger than 80\%. We note that there are a few noise realizations for distances below 9 kpc for which
+precision up to distances of $\sim$ \unit[9]{kpc} (for this particular waveform) with a coverage value
+larger than 80\%. We note that there are a few noise realizations for distances below \unit[9]{kpc} for which
 $\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those cases.
@@
 This is enhancing the probability of overlap. This effect explains why certain values of overlap can reach
-values as high as 80\% even when no signal is added to the noise. \tf{I don't understand this last comment.} \mab{Toni: is that clearer?}
+values as high as 80\% even when no signal is added to the noise.
+%\tf{I don't understand this last comment.} \mab{Toni: is that clearer?}
 
@@
 In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers.
-Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study.
+Three sensitivity curves, ET\_B, ET\_C and ET\_D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study.
 %(cf.~Fig.~\ref{fig:spectrum}).
@@
   \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
-  \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.
-\pcd{[PCD: Should we show the corresponding blue ranges for the pure noise case?]}\mab{MAB: We would need to show 5 noise ranges.
-I haven't tried what it looks like.
-Could give a try}}
+  \caption{Data points (``+'') show the median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise.
+The solid lines are smoothing splines.
+The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal for ET\_D noise realisations.}
   \label{fig:s20--SFHo_all3G}
@@
 This is partly due to the small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$ rather uncertain for this particular waveform.
-\tf{Is this really understood?
+%\tf{Is this really understood?
 What's particular in this waveform?}
@@
 We also provide upper limits for $d_{\rm det}$ by taking into account the detector antenna response in our simulations and assuming that the source is optimally oriented with a matched filter signal-to-noise ratio of 13.
-Table \ref{tab:results} shows that the results for the Advanced Virgo detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity.
+Table \ref{tab:results} shows that the results for the AdV detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity.
 It is remarkable that for third-generation detectors the PNS surface gravity  could be reconstructed for sources located up to a few hundreds of kpc.

### 2020-11-10 simulations.tex
prep version to be circulated in LVK

--- 2020-11-06 simulations.tex
@@
 
-For the {\it model set}, we use the $25$ spherically symmetric (1D) simulations of \citep{Torres:2019a}
+For the {\it model set}, we use the \pcd{$18$ }spherically symmetric (1D) simulations of \citep{Torres:2019a}
 including progenitors with zero-age main sequence (ZAMS) masses in the range 
-$M_{\mathrm{ZAMS}} = 11.2 - 75 \, \Msol$. The set contains simulations using  
+$M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 75]{$\Msol$}. The set contains simulations using  
 six different EOS. Details can be found in
@@
 (see Table~\ref{Tab:2dSimList} for a list of models).
-All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}} = 11.2 - 40 \, \Msol$
+All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 40]{$\Msol$}
  evolved through the hydrostatic phases by
@@
 We performed one simulation of each stellar model using the EOS of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
-an incompressibility of $K = 220 \, \mathrm{MeV}$ (LS220) and added
+an incompressibility of $K=$ \unit[220]{MeV} (LS220) and added
 comparison simulations with the SFHo EOS
@@
 \cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-for the {progenitor} with $M_{\mathrm{ZAMS}} = 15 \, \Msol$. To this set of
+for the {progenitor} with $M_{\mathrm{ZAMS}}=$ \unit[15]{$\Msol$}. To this set of
 simulations we add the waveform of a 2D model used in
 \cite{Torres:2019a}, denoted \texttt{s20S}. It corresponds to a star
-with the same initial mass, $M_{\mathrm{ZAMS}} = 20 \, \Msol$, as for
+with the same initial mass, $M_{\mathrm{ZAMS}}=$ \unit[20]{$\Msol$}, as for
 one of the other seven axisymmetric simulations, but was taken from a
@@
 distributed logarithmically with a minimum grid width of
-$(\Delta r)_{\mathrm{min}} = 400 \, \mathrm{m}$ and an outer radius of
-$r_{\mathrm{max}} = 8.3 \times 10^{9} \, \mathrm{cm}$ and
+$(\Delta r)_{\mathrm{min}}=$ \unit[400]{m} and an outer radius of
+$r_{\mathrm{max}} =$ \unit[$8.3 \times 10^{9}$]{cm} and
 $n_{\theta} = 128$ equidistant cells in the angular (polar) direction. For the
 neutrino energies, we used a logarithmic grid with $n_e = 10$ bins up
-to $240 \, \mathrm{MeV}$.
+to \unit[240]{MeV}.
 {Unlike the model set, the simulations in the test set are not 1D because we need to 
@@
 masses of the PNSs are in the range of
-$M_{\mathrm{PNS}} = 1.47 - 2.33 \, \Msol$, i.e., likely insufficient for
+$M_{\mathrm{PNS}} =$ \unit[1.47 -- 2.33]{$\Msol$}, i.e., likely insufficient for
 producing a black hole.

### 2020-11-10 ccsn.tex
Typo corrected in the abstract.

@@
 The stochastic nature of the late post-bounce gravitational wave signal due to the non-linear dynamics of the matter involved and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
-In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS)
+In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS).
 Numerical simulations of CCNS have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.

### 2020-11-10 conclusion.tex
Fixed typo in the conclusions.

@@
 The detection of GW from CCSN may help improve our current understanding of the explosion mechanism of massive stars.  
-In this paper we have proposed an explorary method to infer PNS properties using an approach based on GW associated with convective oscillations of PNS.
+In this paper we have proposed an exploratory method to infer PNS properties using an approach based on GW associated with convective oscillations of PNS.
 As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.

### 2020-11-10 ccsn.tex
Fixing typo in the acknowledgements.

@@
 
-\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071) and by the European Union’s Horizon 2020 research and innovation (RISE) programme H2020-MSCA-RISE-2017 GrantNo.
-FunFiCO-777740.
+\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071) and by the European Union’s Horizon 2020 research and innovation (RISE) programme (H2020-MSCA-RISE-2017 GrantNo.
+FunFiCO-777740).
 N.C. acknowledges support from National Science Foundation grant PHY-1806990.

### 2020-11-14 introduction.tex
Update introduction.tex
Fixed typo

@@
 
-In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensional effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
 These} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs.

### 2020-11-18 ccsn.tex
Update ccsn.tex
add a PE paragraph in the introduction

@@
 \affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
-\affiliation{Carleton College, Northfield, MN 55057, USA}
+\affiliation{Physics and Astronomy, Carleton College, Northfield, MN 55057, USA}
 
@@
 In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS).
-Numerical simulations of CCNS have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
+Numerical simulations of CCSN have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
 We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.

### 2020-11-18 introduction.tex
add a PE paragraph in the introduction

@@
 from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}.
-This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015wra, Kuroda:2016, Andresen:2017}.
 In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
@@
 
+Many previous data analysis efforts focused on the reconstruction of the gravitational wave strain amplitude
+without assuming a particular signal model \citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As numerical
+simulations became more available other methods using catalogs of waveforms have been proposed to
+identify the supernova explosion mechanism.
+Among them, principal component analysis helps at reducing the complexity of CCSN waveform catalogs to fewer
+parameters \citep{Heng:2009,roever:09,Edwards:2014,powell:2016,powell:2017,Roma:2019kcd}.
+The classification challenge is also well addressed with deep learning methods \citep{astone:2018,Chan:2019fuz}.
+
 {In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  

### 2020-11-20 ccsn.tex
cleanup biblio

@@
 %\newcommand{\mnras}{MNRAS}
+\def\mnras{Mon. Not. Roy. Astron. Soc.}
 \def\apjl{Astrophys. J. Lett.}
+\def\apj{Astrophys. J.}
 \def\prd{Phys.~Rev.~D}
@@
 In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS).
-Numerical simulations of CCSN have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\cite{Torres:2019b}.
+Numerical simulations of CCSN have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\citep{Torres:2019b}.
 We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.

### 2020-11-20 conclusion.tex
cleanup biblio

@@
 In this paper we have proposed an exploratory method to infer PNS properties using an approach based on GW associated with convective oscillations of PNS.
-As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.
+As shown by~\citep{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations.
 Such modes are responsible for a significant fraction of the highly stochastic GW emitted after core bounce.
@@
 This relationship is extracted from the GW data using an algorithm developed for this investigation.
-This algorithm is  a first attempt to infer the time evolution of a particular combination of the PNS mass and radius based on the universal relations found in~\cite{Torres:2019b}.
+This algorithm is  a first attempt to infer the time evolution of a particular combination of the PNS mass and radius based on the universal relations found in~\citep{Torres:2019b}.
 More precisely, we have considered the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity) derived from the observation of the $\mbox{}^2g_2$ oscillation mode in the numerically generated GW data.

### 2020-11-20 introduction.tex
cleanup biblio

--- 2020-11-10 introduction.tex
@@
 The life of sufficiently massive stars, i.e.~those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$, ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion.
-Nearby core-collapse supernova (CCSN) explosions are expected to be sources of gravitational waves (GWs) and they could be
-%are one of the main candidates for 
-the next great discovery of current ground-based observatories.
+Nearby core-collapse supernova (CCSN) explosions are expected to be sources of gravitational waves (GWs) and they could be the next great discovery of current ground-based observatories.
 However, these are relative rare events.
-A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors.
-The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
-KAGRA~\citep{Aso:2013eba}, within $\sim$ \unit[5]{kpc} \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
-For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within \unit[50]{kpc} and, for some extreme models, up to \unit[5--30]{Mpc} \citep{Gossan:2016,TargetedSNSearchO12}.
-However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim$ \unit[$10^{-4}$]{$\rm {yr}^{-1}$}.
-%For the results discussed in this work we only consider neutrino-driven CCSN of non-rotating stellar cores.%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN}
+A neutrino-driven explosion~\citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors.
+The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and KAGRA~\citep{Aso:2013eba}, within $\sim$ \unit[5]{kpc}~\citep{Gossan:2016,TargetedSNSearchO12}.
+Such a galactic event has a rate of about $2-3$ per century~\citep{Adams:2013,Rozwadowska:2021}.
+For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within \unit[50]{kpc} and, for some extreme models, up to \unit[5--30]{Mpc}~\citep{Gossan:2016,TargetedSNSearchO12}.
+However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe~\citep{Li:2011b} or events associated with long GRBs~\citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim$ \unit[$10^{-4}$]{$\rm {yr}^{-1}$}.
 Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.
 
-In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.
+In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensional effects~\citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface~\citep{Kotake:2017}.
 These} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs.
@@
 from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}.
-This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}.
+This feature has been interpreted as a continuously excited gravity mode (g-mode, see~\citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS~\citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015wra, Kuroda:2016, Andresen:2017}.
 In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
@@
  {by means of linear perturbation analysis of background PNS models}.
-The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}.
-Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}.
-Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transitions \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}.
-{Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
+The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.}~\citep{McDermott:1983}.
+Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core~\citep{Reisenegger:1992}.
+Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation~\citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities~\citep{Dimmelmeier:2006}, phase transitions~\citep{Kruger:2015} and realistic equation of state~\citep{Camelio:2017}.
+{Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution~\citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
 
 Using results from 2D CCSN numerical simulations as a background~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region)
-shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
+shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included~\citep{Torres:2019a}).
 %\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations.
@@
 This reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. 
-Further work was presented in {\citep{Torres:2019b} who found that it is possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the numerical evolutions}.
+Further work was presented in~\citep{Torres:2019b} who found that it is possible to derive simple relations between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the numerical evolutions}.
 These relations are universal as they do not depend {on the equation of state (EOS) or on the mass of the progenitor {and they only depend weakly on} the numerical code used {(see discussion in Section~\ref{sec:simulations})}.
-{Similar universal relations have been discussed by \citep{Sotani:2020a,Sotani:2020b} who also found that they do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.
+{Similar universal relations have been discussed by~\citep{Sotani:2020a,Sotani:2020b} who also found that they do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.
+
+Many previous data analysis efforts focused on the reconstruction of the gravitational wave strain amplitude
+without assuming a particular signal model~\citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As numerical
+simulations became more available other methods using catalogs of waveforms have been proposed to identify
+the supernova explosion mechanism.
+Among them, principal component analysis helps at reducing the complexity of CCSN waveform catalogs to fewer
+parameters~\citep{Heng:2009,roever:09,Edwards:2014,powell:2016,powell:2017,Roma:2019kcd}.
+The classification challenge is also well addressed with deep learning methods~\citep{astone:2018,Chan:2019fuz}.
 
@@
 extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
-This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.} Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention).
+Next, we use the universal relations obtained by~\citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.} Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
 

### 2020-11-20 methods.tex
cleanup biblio

@@
 \end{equation}
-with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}
-\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.
+with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}~\citep{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.
 
@@
 A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
-autoregressive estimates~\cite{BrockwellPeterJ1991Tsta} of the local spectra for successive time intervals of 
-length 200 with a 90\% overlap.
+autoregressive estimates~\citep{BrockwellPeterJ1991Tsta} of the local spectra for successive time intervals of length 200 with a 90\% overlap.
 The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
@@
 band given the monotonicity constraint. Specific details about the reconstruction of the $\mbox{}^2 g_2$-mode ridge 
-are provided in Appendix \ref{app:gmode}.
+are provided in Appendix~\ref{app:gmode}.
 

### 2020-11-20 results.tex
cleanup biblio

@@
 100 Gaussian noise realisations whose power spectral density (PSD) follows the aLIGO
-spectrum~\cite{aLIGOsens:2018}. %The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
+spectrum~\citep{aLIGOsens:2018}. %The PSD of aLIGO and AdV at design sensitivities, along with those of planned third-generation detectors, are shown in Figure~\ref{fig:spectrum}.
 
@@
 In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers.
-Three sensitivity curves, ET\_B, ET\_C and ET\_D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study.
+Three sensitivity curves, ET\_B, ET\_C and ET\_D corresponding to different options and stages of the project~\citep{Hild_2011} are considered in our study.
 %(cf.~Fig.~\ref{fig:spectrum}).
-The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
+The US based project Cosmic Explorer~\citep{reitze2019cosmic} is expecting to reach its design
 sensitivity circa 2040 through two phases labeled CE1 and CE2. %Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.

### 2020-11-20 simulations.tex
cleanup biblio

@@
 frequency of some particular modes visible in the GW spectrum.
-The frequency of these modes depends in a universal way on the surface gravity of the PNS, $r\equiv M_{\rm PNS}/R_{\rm PNS}^2$ \citep{Torres:2019b}.
+The frequency of these modes depends in a universal way on the surface gravity of the PNS, $r\equiv M_{\rm PNS}/R_{\rm PNS}^2$~\citep{Torres:2019b}.
 Therefore, if at a given time GW emission is observed at a certain
@@
 %description of the emission and transport of neutrinos.
-Both the {\it model set} and {\it test set} simulations have been generated using the AENUS-ALCAR code
-\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
-which combines special relativistic (magneto-)hydrodynamics, a modified
-Newtonian gravitational potential approximating the effects of general
-relativity \citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
-two-moment neutrino transport solver
-\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
+Both the {\it model set} and {\it test set} simulations have been generated using the AENUS-ALCAR code~\citep{Just:2015}
+which combines special relativistic (magneto-)hydrodynamics, a modified Newtonian gravitational potential approximating the effects of general relativity~\citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
+two-moment neutrino transport solver~\citep{Just:2015}.
 All siimulations include the relevant reactions between matter and neutrinos of all
@@
 
-For the {\it model set}, we use the \pcd{$18$ }spherically symmetric (1D) simulations of \citep{Torres:2019a}
-including progenitors with zero-age main sequence (ZAMS) masses in the range 
-$M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 75]{$\Msol$}. The set contains simulations using  
-six different EOS. Details can be found in
- \citep{Torres:2019a}. The reason to use 1D simulations for the model set
- is that their computational cost is significantly smaller than that of multidimensional
- simulations which allows us to accumulate the statistics necessary to build a good model for $r(f)$.
- {For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
- described in \cite{Torres:2018,Torres:2019a,Torres:2019b}. }
+For the {\it model set}, we use the $18$ spherically symmetric (1D) simulations of~\citep{Torres:2019a} including progenitors with zero-age main sequence (ZAMS) masses in the range $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 75]{$\Msol$}.
+The set contains simulations using six different EOS.
+Details can be found in~\citep{Torres:2019a}.
+The reason to use 1D simulations for the model set is that their computational cost is significantly smaller than that of multidimensional simulations which allows us to accumulate the statistics necessary to build a good model for $r(f)$.
+For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis described in \cite{Torres:2018,Torres:2019a,Torres:2019b}.
 
@@
 All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 40]{$\Msol$}
- evolved through the hydrostatic phases by
-\cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
-We performed one simulation of each stellar model using the EOS of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
-an incompressibility of $K=$ \unit[220]{MeV} (LS220) and added
-comparison simulations with the SFHo EOS
-\cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
-and the GShen EOS 
-\cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
-for the {progenitor} with $M_{\mathrm{ZAMS}}=$ \unit[15]{$\Msol$}. To this set of
-simulations we add the waveform of a 2D model used in
-\cite{Torres:2019a}, denoted \texttt{s20S}. It corresponds to a star
-with the same initial mass, $M_{\mathrm{ZAMS}}=$ \unit[20]{$\Msol$}, as for
-one of the other seven axisymmetric simulations, but was taken from a
-newer set of stellar-evolution models
-\cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
+ evolved through the hydrostatic phases by~\citep{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
+We performed one simulation of each stellar model using the EOS of~\citep{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with an incompressibility of $K=$ \unit[220]{MeV} (LS220) and added comparison simulations with the SFHo EOS~\citep{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations} and the GShen EOS~\citep{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations} for the {progenitor} with $M_{\mathrm{ZAMS}}=$ \unit[15]{$\Msol$}.
+To this set of simulations we add the waveform of a 2D model used in~\citep{Torres:2019a}, denoted \texttt{s20S}.
+It corresponds to a star with the same initial mass, $M_{\mathrm{ZAMS}}=$ \unit[20]{$\Msol$}, as for one of the other seven axisymmetric simulations, but was taken from a newer set of stellar-evolution models~\citep{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
 It was evolved with the SFHo EOS.

### 2020-11-23 introduction.tex
Some editing in the intro.

@@
 This feature has been interpreted as a continuously excited gravity mode (g-mode, see~\citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS~\citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015wra, Kuroda:2016, Andresen:2017}.
-In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
-%promising {source} for 
-ground-based GW interferometers.
+In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.
+The {typical} frequencies of {these} modes make them interesting targets for ground-based GW interferometers.
 
- The {properties of} g-modes in hot {PNSs} have been studied since the 1990s
- %end of last century 
- {by means of linear perturbation analysis of background PNS models}.
+ The {properties of} g-modes in hot {PNSs} have been studied since the 1990s {by means of linear perturbation analysis of background PNS models}.
 The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.}~\citep{McDermott:1983}.
@@
 shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included~\citep{Torres:2019a}).
-%\tf{This last sentence might be improved.} \mab{MAB: What about: Computing the mode spectrum of the region within the shock (including the PNS and the post-shock region) consists in solving an eigenvalue problem using background simulations.
-Using results from 2D CCSN numerical simulations \citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}).
-} \pcd{[PCD: I am not convinced about the first sentence of mab's proposal.
-Maybe if we just delete "with respect to which to solve an eigenvalue problem" the sentence becomes more readable]}
 
@@
 
-Many previous data analysis efforts focused on the reconstruction of the gravitational wave strain amplitude
-without assuming a particular signal model~\citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As numerical
-simulations became more available other methods using catalogs of waveforms have been proposed to identify
+Previous data analysis efforts have focused on the reconstruction of the GW strain amplitude
+without assuming a particular signal model~\citep{Summerscales:2008,Klimenko:2015ypf,CornLitt}. As the amount of numerical
+simulations increased other methods using waveform catalogs have been proposed to identify
 the supernova explosion mechanism.
-Among them, principal component analysis helps at reducing the complexity of CCSN waveform catalogs to fewer
+Among them, principal component analysis helps at reducing the complexity of CCSN waveforms to fewer
 parameters~\citep{Heng:2009,roever:09,Edwards:2014,powell:2016,powell:2017,Roma:2019kcd}.

### 2021-01-20 referee_answers.txt
Referee comments

We thank the refereee for his/her supportive comments and provide
answers below.

----------------------------------------------------------------------
Report of the Referee -- DY12631/Bizouard
----------------------------------------------------------------------

1. As far as I can tell, the model assumes that the GW signal is
dominated by g-modes originating from the PNS, which is consistent
with results reported from recent CCSN simulations from some groups.
However, other groups, using three-dimensional simulations, attribute
the dominant component to f- and g-modes (Radice et al. (2019), ApJL,
876, L9) or p- and g-modes (Mezzacappa et al. (2020), Phys. Rev. D,
102, 023027). Can the authors comment on this. In particular, discuss
the possible impact on model predictions if the signal is hybrid in
nature.

2. Most analysis of GW signals is based on two-dimensional models, and
the test set used in the present study is also comprised of 2D
simulations. Recent comparisons of signals from 2D and 3D simulations
show that the signal amplitude from 2D simulations is considerably
higher than in 3D simulations (e.g., Andresen et al. (2017), MNRAS,
468, 2032; see also Mezzacappa et al. 2020). I would like the authors
to comment on this, and its potential impact on the results presented
in the present paper. Would this just lower the upper limit of the
detectability distance? Are there other crucial differences between
signals from 2D and 3D simulations (e.g., the low frequency signal
from the spiral SASI mode) that could compromise the accuracy of the
model derived here? Please discuss this.

3. Are the 2D simulations new (i.e., computed for the present paper),
or have they been published elsewhere? If they have been published
elsewhere, please provide citation. Were any of the 2D simulations
constrained to spherical symmetry in the inner core to alleviate the
CFL condition? If so, could this restriction potentially suppress
modes that would otherwise have been excited by instabilities in the
PNS?

4. How do you define R_{PNS} in the 1D and 2D simulations? Is it the
radius for which the density achieves a specific value? Please clarify
this.

5. Can you elaborate on the motivation for the degrees of the
polynomials in Eqs. (1) and (2)? Why is r_i of degree 3 and log
sigma_i of degree 2?

A: We have tested several combinations using s20 waveform (up to degree
3 for both r_i and sigma_i) and have selected the best fitting model.

6. What do the red circles in Figure 1 represent? Is it a scatter plot
of the results from all 1D simulations taken together? Specifically,
since there obviously is more than 18 circles, how often is data from
each simulation sampled to generate the circles in Figure 1? Please
clarify this. Perhaps using a unique color for each model would help
clarify (but I do not demand this).

A: Each of the 18 1D simulations is providing data points r_i that are
plotted as function of f_i. Each simulation has between 100 and 300
points. All of them are plotted on Figure 1.

Should we make the plot with a color different for the 18 simulations?

### 2021-01-22 referee_answers.txt
model update
- Updating Figure 1 and caption
- Description of model selection
- Answer to Q5

@@
 
-A: We have tested several combinations using s20 waveform (up to degree
-3 for both r_i and sigma_i) and have selected the best fitting model.
+A: Figure 1 gives an idea of the relationship between frequencies and 
+ratios, which is apparently quadratic.  However, we tested a bunch of 
+models, namely, linear, quadratic, cubic, and some other combinations.  
+We assessed their performance through the AIC criterion, getting the 
+cubic as the best model. This is explained in our article.
+
+About the model for log-sigma, we tested only 2 models: a linear and a 
+quadratic. We found that the best model is the latter according to the AIC 
+criterion.  Higher order models cause that the iterative fitting procedure 
+does not converge in a satisfactory manner, which is warned by the authors 
+of the lmvar package.
+
+Figure 1 shows that overall the model (for r and log sigma) fits 
+quite well the data set.
+
+We have rewritten that paragraph to clarify this.
 

### 2021-01-22 methods.tex
model update
- Updating Figure 1 and caption
- Description of model selection
- Answer to Q5

@@
  \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set.
-The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.}
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
+\textcolor{red}{We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.} }
  \label{fig:LMVAR}
@@
 
-The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to
-the Akaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{eq:model1}.
-The data and fit of the model including 95\% confidence bands are displayed in
-Figure~\ref{fig:LMVAR}.
+\textcolor{red}{For the mean of the ratios $r_i$, the cubic model has the best fitting amongst polynomials of lower degrees, and for the mean of the log standard deviation $\log\sigma_i$, the quadratic model has the best fit compared to a polynomial of degree one.
+The models were compared according to the Akaike information criterion.
+The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model including 95\% confidence bands are displayed in
+Figure~\ref{fig:LMVAR}.}
+
+%The best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the Akaike information criterion with coefficients given in Table \ref{tab:model}, which is actually the model defined in \eqref{eq:model1}.
+The data and fit of the model including 95\% confidence bands are displayed in Figure~\ref{fig:LMVAR}.
 

### 2021-01-29 referee_answers.txt
Q5
add a plot to answer to refere
update
Update referee_answers.txt

@@
 
+A. We believe that a discussion of the nature of the modes is out of 
+the scope of this paper. As the referee points out correctly, our model 
+assumes that the GW signal is dominated by a g-mode, whose time-frequency 
+evolution is related to the evolution of the ratio M/R^2. In Torres-Forne 
+et al. Phys. Rev. Lett. 123, 051102 (2019) fits to the various modes (f, p, 
+and g) that appear in the spectrograms of a number of CCSN simulations were 
+discussed, and universal relations for the frequency evolution were proposed.
+The best fit for the dominant feature found in all simulations was that 
+corresponding to what was identified as a g-mode. We agree with the referee 
+that uncertainties remain on the actual nature of the various modes identified 
+in simulations performed by various groups. Further work is certainly necessary 
+to clarify this issue, not only egarding three-dimensional simulations but also 
+regarding the development of improved perturbative models for mode identification 
+that incorporate 3D effects and rotation in a more accurate way that existing 
+approaches. The current uncertainty in the identification of the modes at work 
+during the evolution of a PNS, while must definitely be removed, is of no 
+importance for the message that this paper conveys.
+
 2. Most analysis of GW signals is based on two-dimensional models, and
@@
 
+A. The referee is correct. A reduction in the amplitude of the GW signal
+would reduce the detectability distance. Regarding the model we put forward
+in this paper, it is indeed possible that the physical effects that only 
+appear in 3D CCSN simulations could alter the relationship between the 
+mode frequencies and the PNS properties that we have used in this paper. 
+This work, thus, is far from providing the final word on the detectability
+prospects of GW signals from PNS evolution, but it certainly serves the 
+purpose of offering a promising proof-of-concept and a framework upon which 
+to build a better model. We plan to use data from 3D simulations to improve 
+our model and to study the differences in the relationships among 1D, 2D and 
+3D simulations.
+
 3. Are the 2D simulations new (i.e., computed for the present paper),
@@
 
+A. The simulations are new, except the s20 which was used in Torres-Forne 
+et al. Phys. Rev. Lett. 123, 051102 (2019). We avoid the severe time step 
+restriction near the centre by applying a coarsening scheme to the hydrodynamic 
+and neutrino data.  Where the lateral grid width is smaller than the radial 
+one (in the innermost ~ 10 km), we combine in theta-direction several grid cells 
+of the original grid to larger cells.  The resulting effective time step criterion 
+is identical to the radial one, i.e., corresponding to several 100 m.  That way, 
+we save up to two orders of magnitude in computing time.  This method does not 
+imply a reduction to spherical symmetry close to the origin, but maintains the 
+two-dimensional nature and allows for angular flows such as the modes generating the GW signal.
+
 4. How do you define R_{PNS} in the 1D and 2D simulations? Is it the
@@
 this.
+
+A. R_{PNS} is defined as the radius where the density becomes lower than
+10^12 gr/cm^3.
 
@@
 We assessed their performance through the AIC criterion, getting the 
-cubic as the best model. This is explained in our article.
+cubic as the best model. This is explained in the article.
 
@@
 
-We have rewritten that paragraph to clarify this.
+We have rewritten the corresponding paragraph to clarify this.
 
@@
 plotted as function of f_i. Each simulation has between 100 and 300
-points. All of them are plotted on Figure 1.
-
-Should we make the plot with a color different for the 18 simulations?
+points. All of them are plotted on Figure 1. We are included in this report
+the figure where each EOS is singled out with color (ratio_allmass.png).

### 2021-01-29 referee_answers.txt
Update referee_answers.txt

@@
 
-About the model for log-sigma, we tested only 2 models: a linear and a 
+Regarding the model for log-sigma, we tested only 2 models: a linear and a 
 quadratic. We found that the best model is the latter according to the AIC 
-criterion.  Higher order models cause that the iterative fitting procedure 
-does not converge in a satisfactory manner, which is warned by the authors 
+criterion.  For higher order models, the iterative fitting procedure 
+does not converge which is also mentioned in a warning by the authors 
 of the lmvar package.
@@
 plotted as function of f_i. Each simulation has between 100 and 300
-points. All of them are plotted on Figure 1. We are included in this report
+points. All of them are plotted on Figure 1. We now included 
 the figure where each EOS is singled out with color (ratio_allmass.png).

### 2021-01-31 conclusion.tex
add text to answer referee's comments

@@
 In the present study we have assumed that the sources are optimally oriented. The reported distances at which
-we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits.
-Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
+we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits. Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
+\textcolor{red}{Moreover, we have used 2D CCSN simulations that are known to produce higher amplitude GW signals than 3D CCSN simulations~\citep{Radice:2018usf, Mezzacappa:2020lsn}.
+This fact and the low frequency standing accretion shock instability that appears only in 3D simulations may reduce further the performance of the method.}
 Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network. 

### 2021-01-31 introduction.tex
add text to answer referee's comments

@@
 This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by~\citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.} Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
+Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{12}$]{$\rm g/cm^3$}}.
+%  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
+Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
 

### 2021-01-31 referee_answers.txt
Update referee_answers.txt
add text to answer referee's comments

--- 2021-01-29 referee_answers.txt
@@
 our model and to study the differences in the relationships among 1D, 2D and 
-3D simulations.
+3D simulations. We have added a comment on this important caveat in the conclusion.
+
+"Moreover, we have used 2D CCSN simulations
+that are known to produce higher amplitude GW signals
+than 3D CCSN simulations [60, 61]. This fact and the
+low frequency standing accretion shock instability that
+appears only in 3D simulations may reduce further the
+performance of the method."
 
@@
 A. R_{PNS} is defined as the radius where the density becomes lower than
-10^12 gr/cm^3.
+10^12 gr/cm^3. This definition has been added in the text.
 
@@
 
+"For the mean of the ratios $r_i$ , the cubic model has the
+best fitting amongst polynomials of lower degrees, and
+for the mean of the log standard deviation $\rm log \sigma_i$ , the
+quadratic model has the best fit compared to a polynomial
+of degree one. The models were compared according
+to the Akaike information criterion. The estimated coefficients
+are given in Table II and the data and fit of the
+final model including 95\% confidence bands are displayed
+in Figure 1."
+
 6. What do the red circles in Figure 1 represent? Is it a scatter plot

### 2021-01-31 referee_answers.txt

@@
 
-About the model for log-sigma, we tested only 2 models: a linear and a 
+Regarding the model for log-sigma, we tested only 2 models: a linear and a 
 quadratic. We found that the best model is the latter according to the AIC 
-criterion.  Higher order models cause that the iterative fitting procedure 
-does not converge in a satisfactory manner, which is warned by the authors 
+criterion.  For higher order models, the iterative fitting procedure 
+does not converge which is also mentioned in a warning by the authors 
 of the lmvar package.
 
-Figure 1 shows that overall the model (for r and log sigma) fits 
-quite well the data set.
+Figure 1 shows that overall the model (for r and log sigma) fits the data set quite well.
 

### 2021-02-01 referee_answers.txt
Working in the reply

@@
 A. We believe that a discussion of the nature of the modes is out of 
-the scope of this paper. As the referee points out correctly, our model 
-assumes that the GW signal is dominated by a g-mode, whose time-frequency 
-evolution is related to the evolution of the ratio M/R^2. In Torres-Forne 
+the scope of this paper. The nature of the mode visible in the time-frequency
+evolution do not affect our analysis. The only assumption is that there is 
+an excited mode (whatever its nature) whose frequency scales with the properties 
+of the proto-neutron star following a universal relation. It is true that
+the ratio used (M/R^2) is inspired by the presence of a g-mode and we and this
+interpretation is discussed in our previous work. In Torres-Forne 
 et al. Phys. Rev. Lett. 123, 051102 (2019) fits to the various modes (f, p, 
@@
 in simulations performed by various groups. Further work is certainly necessary 
-to clarify this issue, not only egarding three-dimensional simulations but also 
+to clarify this issue, not only regarding three-dimensional simulations but also 
 regarding the development of improved perturbative models for mode identification 
@@
 importance for the message that this paper conveys.
+
+We have modified slightly the text in page 2 to clarify this:
+
+"(different authors may have slightly different naming convention)"
+->
+"This same mode has been interpreted in some works as a f or a p- mode [e.g. 29, 48, 49]. 
+However, this does not affects our analysis, which only uses the relation between the 
+frequency of the main mode emitting in GWs (regardless of its nature) and the properties 
+of the PNS."
 
@@
 mode frequencies and the PNS properties that we have used in this paper. 
-This work, thus, is far from providing the final word on the detectability
-prospects of GW signals from PNS evolution, but it certainly serves the 
-purpose of offering a promising proof-of-concept and a framework upon which 
-to build a better model. We plan to use data from 3D simulations to improve 
-our model and to study the differences in the relationships among 1D, 2D and 
-3D simulations. We have added a comment on this important caveat in the conclusion.
+Sotani & Takiwaki 2020a,b, compared simulations with different dimensionality 
+(1D, 2D and 3D) and found no differences for a limited set of models (this
+is mentioned in our work in the introduction). This work, thus, is far from 
+providing the final word on the detectability prospects of GW signals from PNS 
+evolution, but it certainly serves the purpose of offering a promising 
+proof-of-concept and a framework upon which to build a better model. We plan 
+to use data from 3D simulations to improve our model and to study the differences 
+in the relationships among 1D, 2D and 3D simulations. We have added a comment on 
+this important caveat in the conclusion.
 
@@
 imply a reduction to spherical symmetry close to the origin, but maintains the 
-two-dimensional nature and allows for angular flows such as the modes generating the GW signal.
+two-dimensional nature and allows for angular flows such as the modes generating 
+the GW signal.
 
@@
 A. R_{PNS} is defined as the radius where the density becomes lower than
-10^12 gr/cm^3. This definition has been added in the text.
+10^11 gr/cm^3. This definition has been added in the text.
 

### 2021-02-01 introduction.tex
Working in the reply

@@
   extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
-This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{12}$]{$\rm g/cm^3$}}.
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b}
+ %(different authors may have slightly different naming convention)
+Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{11}$]{$\rm g/cm^3$}}.
 %  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
+\textcolor{red}{This same mode has been interpreted in some works as a f or a p-mode
+\citep[e.g.][]{Sotani:2016,Radice:2018usf,Mezzacappa:2020lsn}. 
+However, this does not affects our analysis, which 
+only uses the relation between the frequency of the main mode emitting in GWs (regardless
+of its nature) and the properties of the PNS.}    
 Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2021-02-02 referee_answers.txt
Small modification in 6
minor revision

@@
 approaches. The current uncertainty in the identification of the modes at work 
-during the evolution of a PNS, while must definitely be removed, is of no 
-importance for the message that this paper conveys.
+during the evolution of a PNS is of no importance for the message that this
+paper conveys.
 
@@
 that are known to produce higher amplitude GW signals
-than 3D CCSN simulations [60, 61]. This fact and the
+than 3D CCSN simulations [48, 49]. This fact and the
 low frequency standing accretion shock instability that
@@
 
-Figure 1 shows that overall the model (for r and log sigma) fits the data set quite well.
+Figure 1 shows that overall the model (for r and log sigma) fits the
+data set quite well.
 
@@
 plotted as function of f_i. Each simulation has between 100 and 300
-points. All of them are plotted on Figure 1. We now included 
-the figure where each EOS is singled out with color (ratio_allmass.png).
+points. All of them are plotted on Figure 1. We share with the referee 
+a figure where each EOS is singled out with color (ratio_allmass.png).

### 2021-02-02 methods.tex
minor revision

@@
 The models were compared according to the Akaike information criterion.
-The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model including 95\% confidence bands are displayed in
-Figure~\ref{fig:LMVAR}.}
+The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model, including 95\% confidence bands, are displayed in Figure~\ref{fig:LMVAR}.}
 

### 2021-02-02 appendixA.tex
Comments on the rest
add NLC comments
G-mode reconstruction update
Moving average added in the text
minor revision

--- 2020-06-30 appendixA.tex
@@
-\section{G-mode reconstruction}
-
-Given the spectrogram and an specified time interval for the g-mode reconstruction, our proposal method works as follows.
+\section{g-mode reconstruction}
+\label{app:gmode}
+Given the spectrogram and a prescribed time interval for the $^2g_2$-mode reconstruction, our proposed method works as follows.
 The starting point must be specified.
 It can be either at the beginning or at the end of the signal.
-Then, in one of these extremes, the maximum energy value is identified, registering its frequency.
+Then, in one of these two extremes, the maximum energy value is identified, registering its frequency.
 This is done independently for a number of consecutive time intervals.
-Then we calculate the median of these frequency values, providing a robust starting value for the g-mode reconstruction.
+Then, we calculate the median of these frequency values, providing a robust starting value for the $^2g_2$-mode reconstruction.
 
-The starting frequency value is the first g-mode estimate for the first or the last time interval, depending on the specified starting location.
-If the reconstruction is set to start at the beginning of the signal, the reconstruction will be done progressively over the time intervals, where each maximum frequency value will be calculated within a frequency range specified by the previous g-mode estimate.
-Given the non-decreasing behaviour of the true g-mode values, the g-mode estimates will be forced to be greater or equal than the one estimated for its previous time interval, and lower than a specified upper limit.
-As a result, the g-modes estimates will be a non-decreasing sequence of frequency values.
+The starting frequency value is the first $^2g_2$-mode estimate for the first or the last time interval, depending on the starting location we choose.
+If the reconstruction is set to start at the beginning of the signal, the reconstruction will be done progressively over the time intervals, where each maximum frequency value will be calculated within a frequency range specified by the previous $^2g_2$-mode estimate.
+Given the non-decreasing behaviour of the true $^2g_2$-mode values, the mode estimates will be forced to be greater or equal than the one estimated for its previous time interval, and lower than a specified upper limit.
+As a result, the $^2g_2$-mode estimates will be a non-decreasing sequence of frequency values.
+Then, the moving average is applied for smoothing the estimates.
 
-If the reconstruction is set to start at the end of the signal, the g-modes will be estimated backward in time.
-Each maximum frequency is calculated within a range determined by its successor (in time) g-mode estimate.
+If the reconstruction is set to start at the end of the signal, the $^2g_2$-mode will be estimated backward in time.
+Each maximum frequency is calculated within a range determined by its successor (in time) mode estimate.
 These estimates are forced to be lower or equal than its successor (in time) estimate, but greater than a specified lower limit.
-Thus, a non-decreasing sequence of g-mode estimates is guaranteed.
+Thus, a non-decreasing sequence of $^2g_2$-mode estimates is guaranteed.
+Then, the moving average is applied for smoothing the estimates.
+This $^2g_2$-mode reconstruction method works if and only if the signal is strong enough to provide information about the mode, which is reflected in the spectrogram.
 
-This g-mode reconstruction method works if and only if the signal is strong enough to provide information about the g-mode, which is reflected in the spectrogram.
-
-Given the sequence of g-mode estimates, the confidence band will be calculated by using the model defined in \eqref{eq:universal}.
-The g-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios.
-Since the g-mode estimates are indexed by time, the confidence intervals for the ratios are too.
-Thus, we generate the confidence band by interpolating the lower and upper limits of the collection of consecutive confidence intervals, which will be valid for the time range of the g-mode estimates.
-This confidence band is used to estimate the coverage probabilities in our simulation studies presented below.
+Given the sequence of $^2g_2$-mode estimates, the confidence band will be calculated by using the model defined in Eq.~\eqref{eq:model1}.
+The $^2g_2$-mode estimates are frequency values which we use as predictors in the model in order to generate confidence intervals for the ratios.
+Since the mode estimates are indexed by time, the confidence intervals for the ratios are too.
+Thus, we generate the confidence band by interpolating the lower and upper limits of the collection of consecutive confidence intervals, which will be valid for the time range of the $^2g_2$-mode estimates.
+This confidence band is used to estimate the coverage probabilities in our simulation studies presented in the main text.

### 2021-02-02 conclusion.tex
version corresponding to PRD resubmission

--- 2020-11-20 conclusion.tex
@@
 In the present study we have assumed that the sources are optimally oriented. The reported distances at which
-we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits.
-Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
+we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits. Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
+Moreover, we have used 2D CCSN simulations that are known to produce higher amplitude GW signals than 3D CCSN simulations~\citep{Radice:2018usf, Mezzacappa:2020lsn}.
+This fact and the low frequency standing accretion shock instability that appears only in 3D simulations may reduce further the performance of the method.
 Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network. 

### 2021-02-02 introduction.tex
version corresponding to PRD resubmission

--- 2021-01-31 introduction.tex
@@
   extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN.
-This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention).
-Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{12}$]{$\rm g/cm^3$}}.
+This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b}
+ %(different authors may have slightly different naming convention)
+Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{11}$]{$\rm g/cm^3$}.
 %  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
+This same mode has been interpreted in some works as a f or a p-mode
+\citep[e.g.][]{Sotani:2016,Radice:2018usf,Mezzacappa:2020lsn}. 
+However, this does not affects our analysis, which 
+only uses the relation between the frequency of the main mode emitting in GWs (regardless
+of its nature) and the properties of the PNS.  
 Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2021-02-02 methods.tex
version corresponding to PRD resubmission

--- 2021-01-22 methods.tex
@@
 The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
-\textcolor{red}{We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.} }
+We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.}
  \label{fig:LMVAR}
@@
 
-\textcolor{red}{For the mean of the ratios $r_i$, the cubic model has the best fitting amongst polynomials of lower degrees, and for the mean of the log standard deviation $\log\sigma_i$, the quadratic model has the best fit compared to a polynomial of degree one.
+For the mean of the ratios $r_i$, the cubic model has the best fitting amongst polynomials of lower degrees, and for the mean of the log standard deviation $\log\sigma_i$, the quadratic model has the best fit compared to a polynomial of degree one.
 The models were compared according to the Akaike information criterion.
-The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model including 95\% confidence bands are displayed in
-Figure~\ref{fig:LMVAR}.}
+The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model, including 95\% confidence bands, are displayed in Figure~\ref{fig:LMVAR}.
 

### 2021-11-02 erratum/ccsn_erratum.tex
Erratum

--- 2020-11-20 ccsn.tex
@@
 
-\title{Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae}
+\title{Erratum: Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae [Phys.~Rev.~D 105, 063006 (2021)]}
 
 \author{Marie-Anne~Bizouard}
-\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire de la C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
 
@@
 \author{Nelson~Christensen}
-\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
+\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire de la C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}
 \affiliation{Physics and Astronomy, Carleton College, Northfield, MN 55057, USA}
@@
 
-\begin{abstract}
-The eventual detection of gravitational waves from core-collapse supernovae (CCSN) will help improve our current understanding of the explosion mechanism of massive stars.
-The stochastic nature of the late post-bounce gravitational wave signal due to the non-linear dynamics of the matter involved and the large number of degrees of freedom of the phenomenon make the source parameter inference problem very challenging.
-In this paper we take a step towards that goal and present a parameter estimation approach which is based on the gravitational waves associated with oscillations of proto-neutron stars (PNS).
-Numerical simulations of CCSN have shown that buoyancy-driven g-modes are responsible for a significant fraction of the gravitational wave signal and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations, as demonstrated in~\citep{Torres:2019b}.
-We use a set of 1D CCSN simulations to build a model that relates the  evolution of the PNS properties with the frequency of the dominant g-mode, which is extracted from the gravitational-wave data using a new algorithm we have developed for our study.
-The model is used to infer the time evolution of a combination of the mass and the radius of the PNS.
-The performance of the method is estimated employing simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.
-Considering signals embedded in Gaussian gravitational wave detector noise, we show that it is possible to infer PNS properties for a galactic source using Advanced LIGO and Advanced Virgo data at design sensitivities.
-Third generation detectors such as Einstein Telescope and Cosmic Explorer will allow to test distances of ${\cal O}(100\, {\rm kpc})$.
-%\tf{Caveats?
-Gaussian noise?}
-% 
-%  Core-collapse supernovae are very important phenomena in the Universe and their expected gravitational
-%  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
-%  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
-%  makes the source parameter inference problem very challenging. In this paper, we have considered the
-%  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
-%  whose frequency time evolution is linked to the system properties through universal relations as
-%  demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract from the
-%  gravitational wave data the frequency evolution of the gravity mode. A set of 1D core collapse
-%  supernova simulations is used to build a model of evolution of the proto neutron star properties
-%  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
-%  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
-%  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
-%  11 and 40 solar mass and different equations of state. We have shown that for Advanced LIGO and Advanced
-%  Virgo detectors at design sensitivity it will be possible to infer proton neutron star properties for
-%  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
-%  to test distances of several hundreds of kiloparsec.
-
-\end{abstract}
-
 \maketitle
 
-\input{introduction}\bigskip
+There is a factor $2$ missing in the calculation of the ratio $r\equiv M_{\rm PNS}/R^2_{\rm PNS}$ in \cite{Bizouard2021}.  This only affects the scale in  
+Figs 1 and 3 and the numerical values in table II. Those should be replaced by Fig.~\ref{fig:LMVAR} and \ref{fig:ratio} and 
+Table~\ref{tab:model}. The conclusions of the paper remain unaffected.
 
-\input{simulations}\bigskip
+\begin{figure}[t]
+ \centering
+ \includegraphics[width=0.45\textwidth]{plots/model}
+ \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set.
+The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
+We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.
+ \pcd{[RATIO SHOULD BE MULTIPLIED BY A FACTOR 2 IN THE PLOT]}
+ }
+ \label{fig:LMVAR}
+\end{figure}
 
-\input{methods}\bigskip
+\begin{table}[h]
+ \begin{tabular}{crr}
+    \hline
+    Coefficient & \multicolumn{1}{c}{Estimate} & Standard error \\
+    \hline
+   $\beta_1$  &  $ 1.00 \times 10^{-06}$ & $2.12 \times 10^{-08}$ \\   
+   $\beta_2$  &  $-8.22 \times 10^{-10}$ & $5.00 \times 10^{-11}$ \\
+   $\beta_3$  &  $ 1.01 \times 10^{-12}$ & $2.70 \times 10^{-14}$ \\
+   $\alpha_0$ &  $-1.02 \times 10^{+01}$ & $6.80 \times 10^{-02}$ \\
+   $\alpha_1$ &  $ 7.24 \times 10^{-04}$ & $1.56 \times 10^{-04}$ \\
+   $\alpha_2$ &  $ 6.23 \times 10^{-07}$ & $8.15 \times 10^{-08}$ \\   
+    \hline
+  \end{tabular}
+\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.\pcd{[CORRECT NUMBERS IN THE TABLE.
+JUST A FACTOR 2 iIN ALL COEFFICIENTS?]}}\label{tab:model}
+\end{table}
 
-\input{results}\bigskip
+\begin{figure}
+ \centering
+ \includegraphics[width=0.5\textwidth]{plots/ratio}
+ \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles).
+The size of the triangles are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.
+\pcd{[RATIO SHOULD BE MULTIPLIED BY A FACTOR 2 IN THE PLOT]}}
+ \label{fig:ratio}
+\end{figure}
 
-\input{conclusion}\bigskip
-
-\bigskip\noindent\textit{Acknowledgments.--} Work supported by the Spanish Agencia Estatal de Investigaci\'on (PGC2018-095984-B-I00), by the Generalitat Valenciana (PROMETEO/2019/071) and by the European Union’s Horizon 2020 research and innovation (RISE) programme (H2020-MSCA-RISE-2017 GrantNo.
-FunFiCO-777740).
-N.C. acknowledges support from National Science Foundation grant PHY-1806990.
-R.M.
-gratefully acknowledges support by the James Cook Fellowship from Government funding, administered by the Royal Society Te  Ap\={a}rangi.
-R.M.
-and P.M.-R.
-also acknowledge funding from  DFG Grant KI 1443/3-2 and thank the Centre for eResearch at the University of Auckland for their technical support.
-P.C.-D. acknowledges the support of the Spanish {\it Ramon y Cajal} programme (RYC-2015-19074) supporting his research.
-\begin{appendices}
-\input{appendixA}
-\end{appendices}
-%\bibliographystyle{unsrt}
+\bibliographystyle{unsrt}
 \bibliography{biblio}

### 2021-11-02 erratum/ccsn_erratum.tex
Updating table

@@
     \hline
-   $\beta_1$  &  $ 1.00 \times 10^{-06}$ & $2.12 \times 10^{-08}$ \\   
-   $\beta_2$  &  $-8.22 \times 10^{-10}$ & $5.00 \times 10^{-11}$ \\
-   $\beta_3$  &  $ 1.01 \times 10^{-12}$ & $2.70 \times 10^{-14}$ \\
-   $\alpha_0$ &  $-1.02 \times 10^{+01}$ & $6.80 \times 10^{-02}$ \\
+   $\beta_1$  &  $ 2.00 \times 10^{-06}$ & $4.24 \times 10^{-08}$ \\   
+   $\beta_2$  &  $-1.64 \times 10^{-9}$ & $1.00 \times 10^{-10}$ \\
+   $\beta_3$  &  $ 2.02 \times 10^{-12}$ & $5.40 \times 10^{-14}$ \\
+   $\alpha_0$ &  $-9.51 \times 10^{+00}$ & $6.80 \times 10^{-02}$ \\
    $\alpha_1$ &  $ 7.24 \times 10^{-04}$ & $1.56 \times 10^{-04}$ \\
@@
   \end{tabular}
-\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.\pcd{[CORRECT NUMBERS IN THE TABLE.
-JUST A FACTOR 2 iIN ALL COEFFICIENTS?]}}\label{tab:model}
+\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.\pcd{[CHECK IF NUMBERS ARE CORRECT]}}\label{tab:model}
 \end{table}

### 2021-11-08 erratum/ccsn_erratum.tex
Fig1 & Fig3 redone + Table 2 updated

@@
 
-There is a factor $2$ missing in the calculation of the ratio $r\equiv M_{\rm PNS}/R^2_{\rm PNS}$ in \cite{Bizouard2021}.  This only affects the scale in  
-Figs 1 and 3 and the numerical values in table II. Those should be replaced by Fig.~\ref{fig:LMVAR} and \ref{fig:ratio} and 
-Table~\ref{tab:model}. The conclusions of the paper remain unaffected.
+There is a factor $2$ missing in the calculation of the ratio $r\equiv M_{\rm PNS}/R^2_{\rm PNS}$ in \cite{Bizouard2021}.
+This only affects the scale in Figs 1 and 3 and the numerical values in table II.
+Those should be replaced by Fig.~\ref{fig:LMVAR} and \ref{fig:ratio} and Table~\ref{tab:model}.
+The conclusions of the paper remain unaffected.
 
@@
 The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.
-We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.
- \pcd{[RATIO SHOULD BE MULTIPLIED BY A FACTOR 2 IN THE PLOT]}
- }
+We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.}
+
  \label{fig:LMVAR}
@@
     \hline
-   $\beta_1$  &  $ 2.00 \times 10^{-06}$ & $4.24 \times 10^{-08}$ \\   
-   $\beta_2$  &  $-1.64 \times 10^{-9}$ & $1.00 \times 10^{-10}$ \\
-   $\beta_3$  &  $ 2.02 \times 10^{-12}$ & $5.40 \times 10^{-14}$ \\
-   $\alpha_0$ &  $-9.51 \times 10^{+00}$ & $6.80 \times 10^{-02}$ \\
+   $\beta_1$  &  $ 2.00 \times 10^{-06}$ & $4.23 \times 10^{-08}$ \\   
+   $\beta_2$  &  $-1.64 \times 10^{-9}$ & $9.99 \times 10^{-11}$ \\
+   $\beta_3$  &  $ 2.03 \times 10^{-12}$ & $5.41 \times 10^{-14}$ \\
+   $\alpha_0$ &  $-9.54 \times 10^{+00}$ & $6.80 \times 10^{-02}$ \\
    $\alpha_1$ &  $ 7.24 \times 10^{-04}$ & $1.56 \times 10^{-04}$ \\
@@
   \end{tabular}
-\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.\pcd{[CHECK IF NUMBERS ARE CORRECT]}}\label{tab:model}
+\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $\mbox{}^2g_2$ mode.}\label{tab:model}
 \end{table}
@@
 The size of the triangles are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.
-\pcd{[RATIO SHOULD BE MULTIPLIED BY A FACTOR 2 IN THE PLOT]}}
+}
  \label{fig:ratio}
